---
title: "Diachronic collaboration in Kaggle competitions"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
---

This experiment investigates the impact of inheritance on the problem solving ability of individuals. The simplest form of inheritance is where one person passes on their solution to a problem to someone else to serve as a starting point for continued work. How can we measure the impact of this inheritance on the problem solving ability of individuals? If inheritance makes people better problem solvers than they would have been given the same amount of time on their own, then it suggests that the tools and abilities that allow and facilitate collaboration based on inheritance or **diachronic collaboration** may have served to bootstrap the problem solving abilities of humans.

Diachronic collaboration is not the only form of collaboration that may serve to bootstrap the problem solving abilities of humans. In general the ability to communicate and organize in teams allows groups and cultures to do things that individuals cannot. In future experiments, I will compare the productivity of diachronic teams to synchronic teams who are allowed to work all at the same time on a solution. The purpose of the present experiment is to first establish that diachronic teamwork is indeed a more effective way of problem solving than individuals spending the same number of hours working alone.

Why should diachronic collaboration be more effective than individuals working alone? One reason might be that diachronic collaboration encourages better exploration of possible solutions, which in some cases may lead to greater payoffs in solution success. Formal analyses of problem solving strategies necessitate a tradeoff between exploration and optimization in honing in on the best possible solution. These analyses have revealed that there are some problems for which exploration and diversity are more important than optimization, the reason being that for some problems, optimization might lead to premature convergence on a locally minimal solution. As applied to groups of people attempting to solve a problem together, this tradeoff between exploration and optimization is what makes small world networks sometimes (but not always) more effective than fully connected networks, and also why diverse groups of problem solvers in some cases may be more effective than homogenous groups, even if the homogenous groups are all smart.

As applied to diachronic collaboration, inheriting a solution from someone else who likely approached the problem very differently than what would have been done working alone is a form of exploration in the space of possible solutions. The individual who inherits a solution rather than creating one from scratch is exposed to possible solutions that likely would not have been viewed on one's own. This boost in exploration may, in some cases, result in better solutions based on inheritance than would be reached alone.

```{r config, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  fig.width = 3,
  fig.height = 3,
  fig.path = "figs/",
  cache.path = ".cache/"
)

library(evoteams)
library(crotchet)
library(DiagrammeR)
library(magrittr)

teams_per_condition <- 10
diachronic_team_size <- 2
n_participants <- (teams_per_condition) + (teams_per_condition * diachronic_team_size)

```

# Methods

## Participants

For this pilot experiment, participants for `r teams_per_condition` teams in each condition will be recruited, for a total of `r n_participants` participants. These participants will be recruited and screened to determine eligibility. Eligible participants are assigned conditions and contacted to schedule a time to complete the experiment.

## Materials

Participants are recruited to compete in a Kaggle competition. Kaggle competitions are data-oriented challenges in which teams and individuals use statistics and machine learning to make accurate predictions about unlabeled data. For example, a standard Kaggle competition is to predict the survivors of the Titanic. Given a passenger’s ticket information, how accurately can teams predict whether or not the passenger survived? This example is illustrative of the fact that perfect performance is rarely possible in a Kaggle competition, but people can do far better than chance. The way teams improve their predictions is by using statistics and machine learning to generate more and more accurate predictions. Teams can use any method in statistics and machine learning to generate the predictions, which means there are many possible solutions to any Kaggle competition.

## Procedure

Participants are randomly assigned either solo or diachronic conditions. Solo condition participants have 4 hours to come up with a solution. Each participant in the diachronic condition has 2 hours. Participants in the diachronic condition are randomly assigned either the first or the second generation. Second generation participants start with someone in the first generation's final solution. Second generation participants are not constrained by the solution they inherit and may start over if necessary. However, teamwork between diachronic collaborators is incentivized by awarding prizes to the best team solutions in each condition. This is expected but not guaranteed to encourage continuation of inherited solutions, a manipulation that can be checked by investigating changes to solution files.

```{r conditions}
find_graphviz("diachronic-v-solo", package = "evoteams") %>% grViz
```
