---
title: "Proposal"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    fig_caption: true
---

```{r config, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4
)
```

```{r setup}
library(evoteams)
read_evoteams_chunks()
```

Our species is great at working together in teams. Teamwork is typically studied as a network, e.g. the problem solving capabilities of groups of people arranged in different network structures. But a different and pervasive form of teamwork is collaboration over time, where the products of labor (artifacts) are passed from one generation to the next. In this research I will investigate the extent to which team structure determines the type and quality of solutions reached through teamwork. I will compare the effectiveness of teams working either synchronously, at the same time, or diachronically, over evolutionary time to see which form of collaboration results in the greatest progress on a particular problem.

# Hypotheses

# Problem space

# Kaggle competitions

> In this competition, Kagglers are challenged to create a classification algorithm that accurately identifies which customers have the most potential business value for Red Hat based on their characteristics and activities.

# Team structures

```{r team-structure, engine = "dot", dev = "svg", fig.cap = "Two ways to allocate the same number of labor hours. In synchronic teams, team members work on a project at the same time. In diachronic teams, the products of the collaboration are passed from one team member to the next."}
```

# Monotonic growth

# Multiple projects

```{r multiple-projects, engine = "dot", dev = "svg", fig.cap = "Three ways of allocating the same number of calendar hours toward multiple projects. In synchronic teams, team members work on each project from beginning to completion. In divide and conquer teams, team members are assigned independent projects. In diachronic teams, team members work on different projects at the same time, and collaborate over time."}
```