---
title: "Diachronic collaboration yields adaptive problem solving"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    fig_caption: true
---

Teamwork is considered effective when the team accomplishes something together that the members of the team could not have alone. One of the most basic forms of teamwork involves breaking big problems into smaller problems that can be divided among the team members. I call this **synchronic collaboration** because all the members of the team are working, though independently, at the same time. In contrast, **diachronic collaboration** involves team members working one at a time, each building upon the previous solution. Wikipedia article writing is an example of diachronic collaboration. Most real world collaborations have both synchronic and diachronic elements, which makes it difficult to assess the relative effectiveness of each type of collaboration. The experiments outlined below are designed to measure the problem solving capacities of teams organized either diachronically or synchronically. My hypothesis is that organizing teamwork diachronically will make teams more efficient at utilizing feedback to solve problems.

```{r config, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4
)

library(tidyr)
library(pander)
```

```{r setup}
library(evoteams)
read_evoteams_chunks()
read_chunk("chunks/problem-spaces.R")
read_chunk("chunks/hypotheses.R")
```

```{r ivs, results = "asis"}
ivs <- list(
  `Team structure` = c("diachronic", "synchronic"),
  `Feedback` = c("yes", "no", "enforce"),
  `Problem` = c("adaptive", "insight")
)
pandoc.table(ivs, caption = "Independent variables", justify = "left")

exps <- data.frame(
  row.names = c("Team structure", "Problem", "Feedback"),
  `Exp 1` = c("diachronic/synchronic", "adaptive", "yes"),
  `Exp 2` = c("diachronic/synchronic", "adaptive", "no/enforce"),
  `Exp 3` = c("diachronic/synchronic", "insight", NA),
  `Exp 4` = c("diachronic/synchronic", "multiple adaptive", "yes")
) %>% t
pandoc.table(exps, caption = "Overview of experiments and conditions. All manipulations are between team.", justify = "left", missing = "")
```

# Team structures

```{r team-structure, engine = "dot", dev = "svg", fig.cap = "Team structures. Teams of four are allotted the same number of labor hours. In synchronic teams, teams work on a problem all at the same time. In diachronic teams, team members work one at a time, each inheriting a previous solution and improving upon it."}
```

Kaggle competitions are data-oriented challenges where teams build machine learning algorithms and use them to predict some outcome variable. For example, in the Titanic competition, teams are given data about the passengers on the Titanic, including their age, gender, and cabin, and tasked with predicting whether or not the passenger survived the Titanic crash.

Kaggle competitions are examples of adaptive problems because there are many possible solutions to a classification problem (e.g., which algorithm to us, which features to train on, etc.) and they all vary in success. It's not always clear at the start of a classification problem what level of success is obtainable. What teams do know is how their performance compares to others performance, including their past performance. In fact, iteration appears to be key in predicting Kaggle competition winners. The winners in Kaggle competitions appear to be the ones that submit the most entries.

```{r team-structure-results}
```

# Exp 2: Feedback

```{r feedback, engine = "dot", dev = "svg", fig.width = 6, fig.cap = "Three ways of sharing individual contributions with a team of collaborators. When feedback is provided (**left**), the impact of an individual contribution on the performance of the team as a whole is made available on every contribution. When feedback is not provided (**middle**), individual contributions are shared directly with the members of the team without revealing its impact on team performance. When feedback is enforced (**right**), contributions that do not improve team performance are rejected, and cannot be shared with the members of the team."}
```

```{r feedback-results}
```

# Exp. 3: Insight problems

```{r problem-spaces, fig.width = 6}
```

```{r problem-spaces-results}
```

# Exp. 4: Multiple projects

```{r multiple-projects, engine = "dot", dev = "svg", fig.cap = "Two ways of allocating the same number of calendar hours toward multiple projects. In synchronic teams, team members work on each project from beginning to completion in lockstep. In diachronic teams, team members work on different projects at the same time, and collaborate over time."}
```

```{r multiple-projects-results}
```
