---
title: "Diachronic collaboration yields adaptive problem solving"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    fig_caption: true
    css: style.css
---

Teamwork is considered effective when the team accomplishes something together that the members of the team could not have on their own. One of the most common forms of teamwork involves breaking big problems into smaller problems that can be divided among the members of the team and worked on independently. Though the team members are working independently, they are all working at the same time, and this type of teamwork can be called **synchronic collaboration**. In contrast, **diachronic collaboration** involves team members working one at a time on the same problem, each building upon the previous solution. Wikipedia article writing is an example of diachronic collaboration. Most real world collaborations have both synchronic and diachronic elements, which makes it difficult to assess the relative effectiveness of each type of collaboration. The experiments outlined below are designed to measure the problem solving capacities of teams organized either diachronically or synchronically. My hypothesis is that organizing teamwork diachronically will make teams more effective at utilizing feedback to solve problems.

Understanding the relative strengths of diachronic and synchronic collaboration is relevant to the history of human culture and its evolution, in particular the evolution of technology. As a culture, we uphold certain inventors and inventing teams as heroes, working in isolation and creating artifacts well ahead of their time. On this view, inventions would primarily be the products of synchronic collaboration because the inventions are created without precedent. The historical evidence favors an alternative view. Technological progress does not advance by leaps and bounds, but by many small steps and incremental improvements. One conclusion to draw from this evidence is that it is easier to adapt and incrementally improve an existing artifact than to create one from scratch. However, this may only be true when looking across years and generations of human history. This research for the first time compares diachronic and synchronic collaboration as techniques for solving adaptive problems over the same time frame. If, as predicted, diachronic organization is more effective than synchronic organization at utilizing feedback to solve adaptive problems, it suggests that the human abilities that allow for diachronic collaboration--e.g., having a way to inherit and modify past solutions--are more important factors in explaining the evolution of technology than accounting for the ingenuity of individuals.

```{r config, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4,
  fig.align = "center",
  cache = FALSE  # warning! cache is bad with graphviz * svg * read_chunk
)

library(tidyr)
library(pander)
```

```{r setup}
library(evoteams)
read_evoteams_chunks()
read_chunk("chunks/problem-spaces.R")
read_chunk("chunks/hypotheses.R")
read_chunk("chunks/time.R")
```

```{r hypotheses}
```

```{r time}
```

# Design

```{r ivs, results = "asis"}
ivs <- list(
  `Team structure` = c("diachronic", "synchronic"),
  `Feedback` = c("yes", "no", "enforce"),
  `Problem` = c("adaptive", "insight")
)
pandoc.table(ivs, caption = "Independent variables.", justify = "left")

exps <- data.frame(
    row.names = c("Experiment", "Team structure", "Problem", "Feedback"),
    `1` = c("Team structure", "diachronic, synchronic", "adaptive", "yes"),
    `2` = c("Feedback", "diachronic, synchronic", "adaptive", "no, enforce"),
    `3` = c("Insight problem", "diachronic, synchronic", "insight", NA),
    `4` = c("Multiple problems", "diachronic, synchronic", "multiple adaptive", "yes")
  ) %>%
  t %>%
  as.data.frame %>%
  mutate(
    `#` = 1:n()
  ) %>%
  select(`#`, everything())
pandoc.table(exps, justify = "left", missing = "",
             caption = "Overview of experiments and conditions. All manipulations are between team.")
```

Experiment 1 is designed to illustrate the predicted effect of team structure on performance where diachronic teams are expected to outperform synchronic teams on adaptive problems. To investigate whether this difference is due to responsiveness to feedback, in Experiment 2 the feedback made available to the teams is manipulated, either removing it completely or enforcing it such that team performance is only allowed to improve. Removing feedback is expected to impair diachronic more than synchronic performance, and enforcing feedback should improve synchronic more than diachronic performances. Experiment 3 tests the impact of team structure on solving insight problems. Unlike for adaptive problems, for insight problems feedback is unhelpful, as there is no gradation between partial and whole solutions. In solving insight problems, diachronic collaboration is no longer predicted to be more effective than synchronic collaboration, because diachronic teams no longer have an advantage in being able to adapt in response to feedback. The final Experiment 4 is an extended version of Experiment 1 that addresses the discrepancy between calendar hours and labor hours in the team structure manipulation by having both teams complete 4 problems in the same number of calendar hours. These experiments are designed to demonstrate the conditions under which diachronic collaboration is a more effective strategy than synchronic collaboration.

# 1 Team structure

```{r team-structure, engine = "dot", dev = "svg", fig.cap = "**Team structures.** Teams are allotted the same number of labor hours to solve a problem. In synchronic teams, teams work on a problem all at the same time. In diachronic teams, team members work one at a time, each inheriting the team's solution and improving it."}
```

In solving a problem where there are many possible solutions, all varying in degree of success, a team organized diachronically is expected to use labor hours more efficiently than a team organized synchronically. In Experiment 1, teams of four are tasked with solving a classification problem. Classification problems involve generating predicted classes for unlabeled data using statistics and machine learning to improve classification accuracy. For example, a typical classification problem is to predict whether or not a group of passengers on the Titanic survived based on their age, gender, ticket price, cabin, and other features. Teams working diachronically or synchronically are tasked with building a computer program that, when run, generates predicted classes for test data that was withheld from the training set. The programs developed by diachronic teams are predicted to have higher classification accuracy on this test data at the end of the allotted labor hours than the programs developed by synchronic teams.

Classification problems have many possible solutions. Teams can use decision trees, linear regression, neural networks, or any other method to generate the predicted classes. Within each type of solution, there exist innumerable parameters that can be tweaked, including transformations of the input and how to weigh different variables. As varied as the different solutions are, they can all be compared in terms of classification accuracy of the test data. Given this clear measure of success, classification problems lend themselves easily to competitions. For example, Kaggle competitions offer cash rewards to teams with the highest classification accuracy on a particular classification problem.

In Experiment 1, participants with experience solving classification problems are recruited and randomly assigned into teams. Synchronic teams are given two hours with all four team members in the same room to develop a classifier. Each team member is provided a personal work station and network access to the team's program files, but there are no constraints on how the members of the synchronic teams interact or distribute labor over the course of the two hours[^synchronic]. Diachronic teams work one at a time, each spending two hours on the program, and thus are never allowed to interact with the whole team at once[^diachronic]. For diachronic teams, only a single workstation is needed, as each team member simply picks up where the last one left off.

[^synchronic]: To illustrate two extremes, synchronic team members could huddle around a single work station and work on a single solution for two hours, or they could all work independent of one another and at the end pick whoever came up with the best solution on their own.

[^diachronic]: Diachronic team members are allowed to communicate with future collaborators via comments in their code and other project files, although no such communication is required.

Teams can receive feedback on their performance by submitting their program for evaluation against the test data at any time. Teams receive their classification accuracy as feedback, but are not given information about the particular items that were misclassified. After eight labor hours have been spent by each team, the final version of each team's classifier is submitted and tested. The primary hypothesis tested in Experiment 1 is that programs created by diachronic teams will have higher classification accuracy than programs created by synchronic teams.

```{r team-structure-results, fig.cap = "**Adaptive problems**. Diachronic teams are predicted to be more effective at solving adaptive problems than synchronic teams."}
```

The reason diachronic teams are more effective at solving classification problems is hypothesized to be because they are better able to iterate on solutions in response to feedback. If iteration is key to the success of diachronic teams, we would expect to see that diachronic teams should submit more solutions than synchronic teams.

```{r submissions-per-team, fig.cap = "**Submissions per team**. Diachronic teams are predicted to submit more versions of their program for evaluation than synchronic teams."}
```

In understanding the difference between the number of diachronic and synchronic submissions, it's important to distinguish between calendar hours and labor hours. Calendar hours are hours in the traditional sense of hours in a day. Labor hours are calendar hours multiplied by how many people are working. Both diachronic and synchronic teams are allotted 8 labor hours. While synchronic teams spend all their labor hours in the first 2 calendar hours, diachronic teams spread the labor out over 8 calendar hours.

```{r hours-per-team, fig.cap = "**Hours per team**. Diachronic teams work one at a time, and expend one labor hour per calendar hour. Synchronic teams work all at the same time, and expend four labor hours per calendar hour."}
```

The type of collaboration required for this task centers on a single product: the computer program that implements a classifier. Given that the purpose of the teamwork is to collaborate on this one, single product, the number of submissions a team can provide is expected to be constrained by calendar hours more than by labor hours. The rate at which submissions are made by diachronic teams and synchronic teams is expected to be equal in terms of calendar hours. But if the rate at which submissions are made is calculated instead in terms of labor hours, the rate of submissions per labor hour for synchronic teams is expected to be much lower than for diachronic teams.

```{r submissions-per-hour, fig.cap = "**Submissions per hour**. The number of submissions a team makes is expected to depend on calendar time more than labor time."}
```

It seems likely that the rate of submissions per labor hour will be greater for diachronic teams given that the submission rate is constrained more by calendar hours than by labor hours. If diachronic teams outperform synchronic teams, one of the reasons may be because they are better able to adapt a program by incorporating feedback that comes with each additional submission. The purpose of the next experiment is to manipulate the feedback provided to each team.

# 2 Feedback

```{r feedback, engine = "dot", dev = "svg", fig.width = 6, fig.cap = "Three ways of sharing individual contributions with a team of collaborators. When feedback is provided (**left**), the impact of an individual contribution on the performance of the team as a whole is made available on every contribution. When feedback is not provided (**middle**), individual contributions are shared directly with the members of the team without revealing its impact on team performance. When feedback is enforced (**right**), contributions that do not improve team performance are rejected, and cannot be shared with the members of the team."}
```

The hypothesis being tested in Experiment 2 is that diachronic teams outperform synchronic teams on adaptive problem solving because they are better able to utilize feedback to incrementally improve solutions. The type of feedback used in Experiment 1 was passive because it was given on request and in no way affected the team's collaborations. Teams in Experiment 2 either receive no feedback or feedback is enforced such that each contribution cannot be shared with the members of the team unless the contribution improves team performance.

```{r feedback-results, fig.cap = "Impact of feedback types on team performance. Taking feedback away is predicted to impair diachronic team performance more than synchronic team performance. Enforcing feedback such that every change to the code base must improve program performance is predicted to aid synchronic teams more than diachronic teams."}
```

# 3 Insight problem

On an insight problem, you can still get feedback on whether or not your solution is correct, but if you are wrong you don't know if you've gotten any closer to the correct solution.

```{r problem-spaces, fig.width = 6}
```

```{r insight-results}
```

# 4 Multiple problems

```{r multiple-projects, engine = "dot", dev = "svg", fig.cap = "Two ways of allocating the same number of calendar hours toward multiple projects. In synchronic teams, team members work on each project from beginning to completion in lockstep. In diachronic teams, team members work on different projects at the same time, and collaborate over time."}
```

```{r multiple-projects-results}
```

# Summary
