---
title: "Diachronic collaboration yields adaptive problem solving"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
    css: style.css
    fig_caption: true
---

Teamwork is considered effective when the team accomplishes something together that the members of the team could not have on their own. Teams can organize labor in a variety of different ways, but what ways are most effective for solving certain kinds of problems? A common way to organize teamwork is to break big problems into smaller problems that can be divided among the members of the team and worked on independently. Though the team members are working independently, they are all working at the same time, and this type of teamwork can be called **synchronic collaboration**. In contrast, **diachronic collaboration** involves team members working one at a time on the same problem, each building upon the previous solution. Wikipedia article writing is an example of diachronic collaboration. Most real world collaborations have both synchronic and diachronic elements, which makes it difficult to assess the relative effectiveness of each type of collaboration. The experiments outlined below are designed to measure the problem solving capacities of teams organized either diachronically or synchronically. My hypothesis is that organizing teamwork diachronically will make teams more effective at utilizing feedback to solve problems.

Utilizing feedback to solve problems is called adaptive problem solving. Adaptive problem solving involves changing or adapting solutions in response to feedback. Adaptive problems are pervasive in the history of human culture and its evolution, in particular the evolution of technology. Culturally we uphold certain inventors as heroes, having created artifacts and ideas without precedent or precursor, when in fact the historical evidence favors an alternative view. Technological progress does not advance by leaps and bounds, but by many small steps and incremental improvements. One conclusion that can be drawn from this evidence is that it is easier to adapt and incrementally improve an existing artifact through diachronic collaboration than to create one from scratch via synchronic collaboration. However, this continuity may only be true when looking across years and generations of human history. This research for the first time compares diachronic and synchronic collaboration as techniques for solving adaptive problems over the same time frame. If teams organized diachronically are more effective at solving adaptive problems, it suggests that the tools and abilities that allow for diachronic collaboration--e.g., having a way to inherit and modify past solutions--are more important factors in explaining the evolution of technology than accounting for the ingenuity of individuals.

```{r config, echo = FALSE, message = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4,
  fig.align = "left",
  cache = FALSE  # warning! cache is bad with graphviz * svg * read_chunk
)

library(evoteams)
read_evoteams_chunks()
read_chunk("chunks/hypotheses.R")
```

# Design

```{r hypotheses}
```

```{r experiment-table, results = "asis"}
```

These experiments are designed to reveal the conditions under which diachronic collaboration is more effective than synchronic collaboration. Participants are randomly assigned into teams of four and given a problem to solve. The purpose of the Pilot Experiment is to calibrate the difficulty of the problems used in the main experiments to be suitable for four person teams. Experiment 1 investigates the effect of team structure on solving adaptive problems, with the prediction that diachronic collaboration will be more effective than synchronic collaboration. To investigate whether this difference might be due to responsiveness to feedback, in Experiment 2 the feedback made available to the teams is manipulated, either removing it completely or enforcing it such that team performance is only allowed to improve. Removing feedback is predicted to impair diachronic more than synchronic performance, and enforcing feedback should improve synchronic more than diachronic performance. Experiment 3 tests the impact of team structure on solving insight problems. Unlike adaptive problems, for insight problems feedback is generally unhelpful, as there is no gradation between partial and whole solutions. In solving insight problems, diachronic collaboration is no longer predicted to be more effective than synchronic collaboration, because informative feedback is no longer available. The final Experiment 4 is an extension of Experiment 1 that addresses the discrepancy between calendar hours and labor hours in the team structure manipulation by having teams complete four problems in the same number of calendar hours.

# 1 Team structure

Experiment 1 is designed to test the hypothesis that diachronic collaboration is more effective than synchronic collaboration at solving adaptive problems. The type of adaptive problem used in Experiment 1 is a classification problem. Classification problems involve generating predicted labels for unlabeled data using statistics and machine learning to improve the accuracy of the predictions. For example, a typical classification problem is to predict whether or not a group of passengers on the Titanic survived based on their age, gender, ticket price, cabin location, and other features available in the training data. Classification problems are good examples of adaptive problems because they require responsiveness to feedback. Given that there are so many possible solutions to a classification problem (different models to use and parameters to tweak), a useful strategy is to iterate and try many different strategies, each time incrementally improving team performance. Although both diachronic and synchronic teams are able to iteratively develop solutions to classification problems in response to feedback, diachronic collaboration is hypothesized to be able to adapt in response to this feedback more effectively, and thus result in overall better classification accuracy.

```{r team-structure, engine = "dot", dev = "svg", fig.cap = "**Team structures.** Teams of four are allotted the same number of labor hours to solve a problem. Synchronic teams work at the same time on a single solution. Diachronic teams work one at a time, each inheriting the team's solution and improving it."}
```

Synchronic teams are given two hours with all four team members in the same room to write a program that accepts unlabeled data and produces predicted labels. Each team member is provided a personal work station and network access to the team's program files, but there are no constraints on how the members of the synchronic teams interact or delegate labor over the course of the two hours. Diachronic teams work one at a time, each spending two hours on the program, and are never allowed to interact with their whole team all at once. For diachronic teams, only a single workstation is needed. Team members can at any time submit their program for evaluation against the test data, and receive feedback in the form of a classification accuracy score. After eight labor hours have been spent by each team, the final version of each team's program is submitted and evaluated. The primary hypothesis tested in Experiment 1 is that programs created by diachronic teams will have higher classification accuracy than programs created by synchronic teams.

## Is teamwork necessary for this task?

```{r teamwork}
```

## Is diachronic collaboration more effective at adaptive problem solving?

```{r team-structure-results, fig.cap = "**Adaptive problems**. Diachronic teams are predicted to be more effective at solving adaptive problems than synchronic teams."}
```

## Do diachronic teams make more submissions?

```{r submissions-per-team, fig.width = 7, fig.cap = "**Rate of submission depends on calendar hours**. Calendar hours are hours in the traditional sense of hours in a day. Labor hours are calendar hours multiplied by how many people are working. Both diachronic and synchronic teams are allotted 8 labor hours. While synchronic teams spend all labor hours in the first 2 calendar hours, diachronic teams spread the labor out over 8 calendar hours (**left**). If rate of submission depends on calendar hours more than labor hours, then diachronic teams will have more opportunities to respond to feedback than synchronic teams."}
```

Diachronic teams are predicted to outperform synchronic teams, and also to make more submissions than synchronic teams. The purpose of the next experiment is to manipulate the feedback provided to each team.

# 2 Feedback

```{r feedback, engine = "dot", dev = "svg", fig.width = 6, fig.cap = "Three ways of sharing individual contributions with a team of collaborators. When feedback is provided (**left**), the impact of an individual contribution on the performance of the team as a whole is made available on every contribution. When feedback is not provided (**middle**), individual contributions are shared directly with the members of the team without revealing its impact on team performance. When feedback is enforced (**right**), contributions that do not improve team performance are rejected, and cannot be shared with the members of the team."}
```

## Does removing feedback impair diachronic performance?

```{r removing-feedback-results}
```

## Does enforcing feedback improve synchronic performance?

```{r enforcing-feedback-results}
```

# 3 Insight problem

On an insight problem, you can still get feedback on whether or not your solution is correct, but if you are wrong you don't know if you've gotten any closer to the correct solution.

## Do synchronic teams excel at insight problems?

```{r insight-results}
```

# 4 Multiple problems

```{r multiple-projects, engine = "dot", dev = "svg", fig.cap = "Two ways of allocating the same number of calendar hours toward multiple projects. In synchronic teams, team members work on each project from beginning to completion in lockstep. In diachronic teams, team members work on different projects at the same time, and collaborate over time."}
```

## Do the results extend to multiple projects?

```{r multiple-projects-results}
```

## Do teams get more efficient?

```{r team-efficiency}
```

# Summary

These experiments are designed to reveal the problem solving capacities of teams organized either diachronically or synchronically. The purpose of this investigation is not to prescribe certain team structures, since it seems likely different team structures are better equipped to solve different problems, but to outline the conditions for which diachronic collaboration is most effective. Diachronic collaboration is hypothesized to be most effective in solving problems that can be adapted in response to feedback.
