---
title: "Diachronic collaboration yields adaptive problem solving"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
    css: style.css
    fig_caption: true
---

```{r config, echo = FALSE, message = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4,
  fig.align = "left",
  cache = FALSE  # warning! cache is bad with graphviz * svg * read_chunk
)

library(evoteams)
read_evoteams_chunks()
read_chunk("chunks/hypotheses.R")
```

Teamwork is considered effective when the team accomplishes something together that the members of the team could not have on their own. One of the most common forms of teamwork involves breaking big problems into smaller problems that can be divided among the members of the team and worked on independently. Though the team members are working independently, they are all working at the same time, and this type of teamwork can be called **synchronic collaboration**. In contrast, **diachronic collaboration** involves team members working one after the other on the same problem, each building upon the previous solution. Wikipedia article writing is an example of diachronic collaboration. Most real world collaborations have both synchronic and diachronic elements, which makes it difficult to assess the relative effectiveness of each type of collaboration. The experiments outlined below are designed to measure the problem solving capacities of teams organized either diachronically or synchronically. My hypothesis is that organizing teamwork diachronically will make teams more effective at utilizing feedback to solve problems.

Examining the relative strengths of diachronic and synchronic collaboration is relevant to the history of human culture and its evolution, in particular the evolution of technology. Culturally we uphold specific inventors as heroes, having created artifacts and ideas without precedent or precursor, when in fact the historical evidence favors an alternative view. Technological progress does not advance by leaps and bounds, but by many small steps and incremental improvements. One conclusion to draw from this evidence is that it is easier to inherit and incrementally improve an existing artifact than to create one from scratch. However, this continuity may only exist when looking across years and generations of human history. This research for the first time compares diachronic and synchronic collaboration as techniques for solving problems over the same time frame. If teams organized diachronically are more effective at solving adaptive problems, it suggests that the abilities that allow for diachronic collaboration (e.g., having reliable ways to inherit and modify past solutions), may be more important factors in accounting for the evolution of technology than the ingenuity of individuals.

# Design

```{r hypotheses}
```

```{r experiment-table, results = "asis"}
```

These experiments are designed to reveal the conditions under which diachronic collaboration is more effective than synchronic collaboration. Participants are randomly assigned into teams of four and given a problem to solve. The purpose of the Pilot Experiment is to calibrate the difficulty of the problems used in the main experiments. Experiment 1 investigates the effect of team structure on solving adaptive problems, with the prediction that diachronic collaboration will be more effective than synchronic collaboration. To investigate whether this difference might be due to responsiveness to feedback, in Experiment 2 the feedback made available to the teams is manipulated, either removing it completely or enforcing it such that team performance is only allowed to improve. Removing feedback should impair diachronic more than synchronic performance, and enforcing feedback should improve synchronic more than diachronic performance. Experiment 3 tests the impact of team structure on solving insight problems. Unlike adaptive problems, for insight problems feedback is generally unhelpful, as there is no gradation between partial and whole solutions. In solving insight problems, diachronic collaboration is no longer predicted to be more effective than synchronic collaboration, because informative feedback is no longer available. The final Experiment 4 is an extension of Experiment 1 that addresses the discrepancy between calendar hours and labor hours in the team structure manipulation by having teams complete four problems in the same number of calendar hours.

Classification problems involve generating predicted labels for unlabeled data using statistics and machine learning to improve the accuracy of the predictions. For example, a typical classification problem is to predict whether or not a group of passengers on the Titanic survived based on their age, gender, ticket price, cabin location, and other features availabe in the training data. Regardless of method or parameters, all solutions can be compared in terms of classification accuracy, which makes classification problems lend themselves to competition. For example, Kaggle competitions offer cash rewards to teams with the highest classification accuracy on a particular classification problem. Experiment 1 is modeled after a Kaggle competition for the purposes of recruiting skilled and motivated participants.

Synchronic teams are given two hours with all four team members in the same room to write a program that accepts unlabeled data and produces predicted labels. Each team member is provided a personal work station and network access to the team's program files, but there are no constraints on how the members of the synchronic teams interact or delegate labor over the course of the two hours. (For example, all four synchronic team members could work on their own programs and the team could submit the best one at the end, although this would most certainly be an inefficient use of the team's available labor hours.) Diachronic teams work one at a time, each spending two hours on the program, and are never allowed to interact with their whole team all at once. For diachronic teams, only a single workstation is needed. As in a typical Kaggle competition, team members can at any time submit their program for evaluation against the test data, and receive feedback in the form of a classification accuracy score. After eight labor hours have been spent by each team, the final version of each team's program is submitted and evaluated. The primary hypothesis tested in Experiment 1 is that programs created by diachronic teams will have higher classification accuracy than programs created by synchronic teams.

# 1 Team structure

```{r team-structure, engine = "dot", dev = "svg", fig.cap = "**Team structures.** Teams are allotted the same number of labor hours to solve a problem. Synchronic teams work on the same or part of the same solution. Diachronic teams work one at a time, each inheriting the team's solution and improving it."}
```

## Is teamwork necessary for this task?

```{r teamwork}
```

## Is diachronic collaboration more effective at adaptive problem solving?

```{r team-structure-results, fig.cap = "**Adaptive problems**. Diachronic teams are predicted to be more effective at solving adaptive problems than synchronic teams."}
```

## Do diachronic teams make more submissions?

```{r submissions-per-team, fig.width = 7, fig.cap = "**Rate of submission depends on calendar hours**. Calendar hours are hours in the traditional sense of hours in a day. Labor hours are calendar hours multiplied by how many people are working. Both diachronic and synchronic teams are allotted 8 labor hours. While synchronic teams spend all labor hours in the first 2 calendar hours, diachronic teams spread the labor out over 8 calendar hours (**left**). If rate of submission depends on calendar hours more than labor hours, then diachronic teams will have more opportunities to respond to feedback than synchronic teams."}
```

Diachronic teams are predicted to outperform synchronic teams, and also to make more submissions than synchronic teams. The purpose of the next experiment is to manipulate the feedback provided to each team.

# 2 Feedback

```{r feedback, engine = "dot", dev = "svg", fig.width = 6, fig.cap = "Three ways of sharing individual contributions with a team of collaborators. When feedback is provided (**left**), the impact of an individual contribution on the performance of the team as a whole is made available on every contribution. When feedback is not provided (**middle**), individual contributions are shared directly with the members of the team without revealing its impact on team performance. When feedback is enforced (**right**), contributions that do not improve team performance are rejected, and cannot be shared with the members of the team."}
```

## Does removing feedback impair diachronic performance?

```{r removing-feedback-results}
```

## Does enforcing feedback improve synchronic performance?

```{r enforcing-feedback-results}
```

# 3 Insight problem

On an insight problem, you can still get feedback on whether or not your solution is correct, but if you are wrong you don't know if you've gotten any closer to the correct solution.

## Do synchronic teams excel at insight problems?

```{r insight-results}
```

# 4 Multiple problems

```{r multiple-projects, engine = "dot", dev = "svg", fig.cap = "Two ways of allocating the same number of calendar hours toward multiple projects. In synchronic teams, team members work on each project from beginning to completion in lockstep. In diachronic teams, team members work on different projects at the same time, and collaborate over time."}
```

## Do the results extend to multiple projects?

```{r multiple-projects-results}
```

## Do teams get more efficient?

```{r team-efficiency}
```

# Summary

These experiments are designed to reveal the problem solving capacities of teams organized either diachronically or synchronically. The purpose of this investigation is not to prescribe certain team structures, since it seems likely different team structures are better equipped to solve different problems, but to outline the conditions for which diachronic collaboration is most effective. Diachronic collaboration is hypothesized to be most effective in solving problems that can be adapted in response to feedback. 
