---
title: "What is the role of inheritance in the problem solving abilities of individuals?"
author: "Pierce Edmiston"
bibliography: ../refs/diachronic-collaboration.bib
csl: ../refs/apa.csl
output:
  html_document:
    theme: flatly
    css: style.css
    fig_caption: true
---

```{r config, echo = FALSE, message = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev = "svg",
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4,
  fig.align = "center",
  cache = FALSE  # warning! cache is bad with graphviz * svg * read_chunk
)

library(evoteams)
read_evoteams_chunks()
read_chunk("chunks/hypotheses.R")
read_chunk("chunks/problem-spaces.R")
```

```{r hypotheses}
```

What individuals can do and learn to do today critically depends on what others were able to do yesterday. Humans are smart, but we are also hoarders; human cultures accumulate artifacts, ideas, and information far more than any other animal[^uniqueness]. Part of what makes this accumulation helpful is that people never have to start from scratch when solving a problem. Inventors, for example, inherit and adapt existing solutions to problems so pervasively as to make the history of technology one of near continuous change, with each new invention depending heavily on its precursors [@Basalla:1989wq]. But what this historical evidence does not provide is a way to understand more directly the role inheritance may play in the problem solving abilities of individuals. _Does inheriting a partial solution actually make individuals more effective problem solvers or does it simply give people a head start?_ The experiments below are designed to test and measure the impact of inheritance on the problem solving abilities of individuals by studying their productivity in teams. In the control condition, three person teams are given two hours in the same room to solve a problem. Because the members of the team are all working at the same time, these are called **synchronic teams**. In the experimental condition, teamwork is not synchronous and is instead organized around inheritance over time. These **diachronic teams** work one at a time, each (other than the first) inheriting as a starting point the previous team member's solution. My hypothesis is that organizing teamwork diachronically will make for more effective problem solving controlling for the total number of labor hours spent on a solution. These results would suggest that the tools and abilities that facilitate inheritance and diachronic collaboration may serve to bootstrap the problem solving abilities of humans.

[^uniqueness]: Some elements of cumulative culture appear in other animals. See @Hunt:2003dj for an investigation of inherited variation in tools among New Caledonian crows.

```{r team-structure, engine = "dot", fig.cap = "Team structures. Teams of three are allotted the same number of labor hours to solve a problem. Synchronic teams work all at the same time on a single solution. Diachronic teams work one at a time, each (other than the first) inheriting the previous solution and improving it. Additional control groups (not pictured) will be used to verify the difficulty of the problems being solved. Comparing the solution success achieved by diachronic and synchronic teams is proposed as a test of the role of inheritance in the problem solving abilities of individuals."}
```

The conclusions that can be drawn from these experiments depend heavily on the type of problem being solved. Obviously there are many problems that cannot be accomplished by three-person teams each working for two hours, and many more that are not difficult enough to warrant teamwork. In order to be an informative experiment, the problems must be hard enough to benefit from three-person teamwork but not so hard as to prevent contributions from all team members. These conditions will be verified with additional control groups. In order to draw meaningful conclusions from these experiments, problem difficulty must be calibrated so that teams consistently outperform individuals given the same amount of time, and that all individuals can contribute to solutions. In addition to being the right level of difficulty, the problems being solved must also allow for inheritance of solutions. Teams must have a way to share solutions with one another in order to measure the impact of inheritance on problem solving ability. The final requirement is that there are objective and continuous measures of solution success available. This requirement enables a fair comparison of different team structures in terms of problem solving ability by looking at solution success.

Why might inheritance lead to more effective problem solving? One possible explanation is that inheriting an existing solution promotes adaptation and responsiveness to feedback more effectively than in synchronic collaboration. If this is true, then altering the feedback available to teams should differentially impact diachronic and synchronic teams. What this means for understanding cultural evolution is that inheritance may have the strongest impact on problem solving ability for problems where there are informative sources of feedback.

# Design

```{r experiment-table, results = "asis"}
```

The experiments are designed to reveal the conditions under which diachronic collaboration based on inheritance is more effective than synchronic collaboration controlling for labor hours. Participants are randomly assigned into teams and given a problem to solve in a fixed amount of time. The purpose of the Pilot Experiment is to calibrate the difficulty of the problems used in the main experiments to be difficult enough to warrant teamwork, but not impossible for the average participant to progress given enough time. Experiment 1 investigates the effect of team structure on solving classification problems, with the prediction that diachronic collaboration will be more effective than synchronic collaboration. To investigate whether this difference might be due to responsiveness to feedback, in Experiment 2 the feedback made available to the teams is manipulated, either removing it completely or enforcing it such that team performance is only allowed to improve. Removing feedback is predicted to impair diachronic more than synchronic performance, and enforcing feedback should improve synchronic more than diachronic performance. Experiment 3 tests the impact of team structure on solving insight problems. Unlike classification problems, for insight problems feedback is generally unhelpful, as there is no gradation between partial and whole solutions. In solving insight problems, diachronic collaboration is no longer predicted to be more effective than synchronic collaboration, because informative feedback is not available. The final Experiment 4 is an extension of Experiment 1 that addresses the discrepancy between calendar hours and labor hours in the team structure manipulation by having teams complete four problems in the same number of calendar hours.

# 1. Team structure

Experiment 1 is designed to test the hypothesis that diachronic collaboration is more effective than synchronic collaboration at solving classification problems. The solution to a classification problem involves generating predicted labels or "classes" for unlabeled data using statistics and machine learning to improve the accuracy of the predictions. For example, a typical classification problem is to predict whether or not passengers on the Titanic survived based on their age, gender, ticket price, cabin location, and other features of the passenger. Classification problems are good candidates for comparing the effectiveness of different team structures because there are many possible solutions, all varying in degree of success. For example, teams can use decision trees, linear regression, neural network models, and combinations of these, each varying in which features they are trained on and the parameters of the models. Given the vast space of possible solutions, a useful strategy for solving classification problems is to iteratively try out different solutions and incrementally improve classification accuracy[^kaggle]. Although both diachronic and synchronic teams are able to iteratively develop solutions to classification problems, diachronic teams are hypothesized to be more effective than synchronic teams at utilizing this feedback to iteratively improve classification accuracy.

[^kaggle]: For some preliminary evidence of the role of iteration in completing real world classification problems, see my analysis of [classification competition leaderboards](http://rpubs.com/pedmiston/kaggle).

```{r all-teams, engine = "dot", fig.width = 6, fig.cap = "All team structures. **Synchronic teams** are the primary control group, where team members work all at the same time and can discuss and divide labor without constraint. **Diachronic teams** are the experimental condition organized around inheritance where team members work one at a time, each (after the first) starting with the previous team member's solution. To verify that teamwork is necessary to solve the problem, **solo \"teams\"** are recruited and work alone for 6 hours in 2 hour increments. **Expensive teams** are given 3 times as many labor hours as diachronic and synchronic teams and are used to estimate ceiling performance. The effect of inheritance on the problem solving ability of teams is measured as the difference in solution success between diachronic and synchronic teams."}
```

**Procedure**. Skilled and motivated participants are randomly assigned team and condition. Each team is given training data and instructed to write a program that accepts unlabeled test data and generates predicted labels. Synchronic teams are given two hours with all four team members in the same room. Each team member is provided a personal work station and network access to the team's program files, but there are no constraints on how the members of the synchronic teams interact or delegate labor over the course of the two hours. Diachronic teams work one at a time, each spending two hours on the program, and are not allowed to interact with their whole team all at once. Diachronic teamwork is incentivized by explaining diachronic team structures and rewarding participants based on team performance, not on individual performance. For diachronic teams, only a single workstation is needed. In either team, members can at any time submit their program for evaluation against the test data, and receive feedback in the form of a classification accuracy score. After eight labor hours have been spent by each team, the final version of each team's program is submitted and evaluated. The primary hypothesis tested in Experiment 1 is that diachronic teams will produce more accurate predictions than synchronic teams.

## Is it worth it to work in teams on this problem?

In order to interpret the impact of team structure on team performance, it's important to confirm that the task being performed is one that benefits from teamwork. This analysis is done by comparing overall team performance to performance on the same problem for the same amount of time by an individual with equal skill and motivation working alone (**solo teams**). Teams are predicted to outperform individuals on classification problems.

## Are diachronic teams more effective adaptive problem solvers?

The primary hypothesis tested in Experiment 1 is that diachronic collaboration will be more effective than synchronic collaborations at solving classification problems. Problem solving effectiveness is measured in terms of the classification accuracy of the team's final submission. Participants are equally skilled and motivated, in which case a difference in team performance can be attributed to the structure of the team's interactions.

## Do diachronic teams make more submissions?

The primary reason diachronic teams are expected to outperform synchronic teams is because they are better able to utilize feedback and iteratively improve solutions. If this is true, we should expect to see diachronic teams make more submissions (and thus receive more feedback) than synchronic teams.

```{r hours, fig.cap = "Calendar and labor hours. Teams are allotted 8 total labor hours. Synchronic teams of four use all 8 labor hours in the first 2 calendar hours. Diachronic teams spread out their labor hours over 8 calendar hours."}
```

Part of the reason diachronic teams may make more submissions reflects the fact that a team's submission rate may be more dependent on calendar hours than on labor hours. If controlling for submission rate by calendar hours still reveals a difference in submissions for diachronic and synchronic teams, then part of the reason diachronic teams are better able to utilize feedback may be because they able to engage in a faster development cycle.

# 2. Feedback

In Experiment 2, the feedback made available to teams about their performance is altered in order to test the hypothesis that the effectiveness of diachronic collaboration in solving problems depends on reliable sources of feedback. In the first condition, feedback is removed completely, and teams do not know for certain their performance until the end of the experiment. Removing feedback is expected to impair diachronic more than synchronic team performance. In the second condition, feedback is enforced such that a team members contributions can only be shared with the rest of the team if that contribution measurably improves team performance. If diachronic teams are already using feedback to incrementally improve team performance, enforcing feedback should help synchronic teams more by encouraging them to focus on adaptive and incremental improvements in team performance.

```{r feedback, engine = "dot", fig.width = 6, fig.cap = "Feedback mechanisms. When feedback is provided (**left**), changes in team performance are made available to the team after submitting the program for evaluation. This is the feedback condition used in Experiment 1. Experiment 2 includes two other feedback conditions. Feedback is either removed (**middle**) or enforced (**right**) such that contributions that do not improve team performance are rejected and cannot be shared with the members of the team."}
```

**Procedure**. Teams in Experiment 2 are assigned to either the no feedback condition or the enforced feedback position. In the no feedback condition, diachronic and synchronic teams are no longer able to submit their programs for evaluation prior to the end of the experiment. In the enforced feedback condition, every time a team member submits a change to the team's program files, the program is evaluated against the test data. If the contribution does not improve past team performance, it is rejected, and returned for further editing. Otherwise, the contribution can be distributed to the other members of the team.

## Do diachronic teams depend on feedback?

If the adaptive problem solving abilities of diachronic teams are due to increased responsiveness to feedback, then removing feedback should impair diachronic team performance relative to synchronic team performance.

## Does enforcing feedback improve synchronic performance?

If the reason synchronic teams are outperformed by diachronic teams is that they aren't utilizing feedback as effectively, enforcing feedback so that team performance must improve through all changes to the team's solution should improve synchronic team performance more than diachronic team performance.

# 3. Insight problem

Not all problems have informative sources of feedback. Where classification problems have many possible solutions, all varying in degree of success, insight problems have only one solution; they are either solved or not. Experiment 3 investigates the effect of team structure on solving insight problems. Because feedback is unhelpful, diachronic teams do not have an advantage in being able to incrementally improve solutions. For teamwork to be helpful in solving an insight problem, everyone in a team should search for the solution in parallel. This type of problem solving is hypothesized to be better equipped for synchronic collaboration.

```{r problem-spaces, fig.width = 6}
```

The insight problems in Experiment 3 are similarly data-oriented to facilitate comparison across experiments. Participants with the same skills and motivation compete in a "Draw my data" contest. In this contest, teams are given artificial data and asked to identify the object who's image is contained hidden within the data. Solving this problem requires searching through the many variables of the data (most of them irrelevant) and visualizing various combinations until the correct view of the data is discovered. Discovering the object that was "drawn" in the data is the moment of insight that solves the problem.

**Procedure**. Teams complete as many "Draw my data" problems as they can in 8 labor hours, organized synchronically or diachronically. Solutions are submitted by naming the object in the data. If correct, teams receive another dataset. The code used to generate the visualizations and the plots themselves are shared among the members of the team.

## Do insight problems benefit from teamwork?

Before the effect of team structure on performance can be interpreted, the effectiveness of teamwork in solving insight problems first needs to be demonstrated. Teams are expected to complete insight problems at a faster rate than individuals working alone.

## Do synchronic teams excel at insight problems?

Without feedback to guide iteration and change, diachronic teams no longer have an advantage over synchronic teams when solving insight problems. Synchronic teams are expected to outperform diachronic teams because they are better able to coordinate search for the solution in parallel.

# 4. Multiple problems

The purpose of Experiment 4 is to replicate Experiment 1 and extend it to include teams solving multiple adaptive problems over the same number of calendar hours. In Experiment 1, synchronic teams completed their task in 2 calendar hours, where diachronic teams required 8. Even if diachronic teams outperformed synchronic teams, synchronic collaboration may still be seen as effective because they completed the task sooner. In Experiment 4, the conclusions of Experiment 1 are tested in conditions where teams of four complete four different problems in either diachronic or synchronic teams over the same 8 calendar hour time frame.

```{r multiple-projects, engine = "dot", fig.cap = "Two ways of allocating the same number of calendar hours toward multiple problems. Different people are different colors, and different projects are different shapes. In synchronic teams, team members work on each problem from beginning to end in lockstep. In diachronic teams, team members work on different problems at the same time, and collaborate over time."}
```

**Procedure**. Experiment 4 is conducted like Experiment 1, except both teams now compete for 8 hours. Diachronic teams are now placed in the same room with three workstations so they can work all at the same time, though they are prevented from working on unassigned problems.

## Do the results extend to multiple problems?

The predicted results for Experiment 4 are that diachronic teams will outperform synchronic teams across multiple problems.

## Do teams get more efficient?

Participants in Experiment 4 work on four different problems in a row. Do they get better at solving adaptive problems?

# Summary

These experiments are designed to reveal the problem solving capacities of teams organized either diachronically or synchronically. The purpose of this investigation is not to prescribe certain team structures over others, since it seems likely that the most effective teams have elements of both team structures, but to outline the conditions for which diachronic collaboration is most effective. Diachronic collaboration is hypothesized to be most effective in solving problems that can be adapted in response to feedback. Since adaptive problem solving forms the basis for technological evolution, these results suggest that the methods and mechanisms that allow for diachronic collaboration might be critical to explaining the evolution of our species.

# References
