---
title: "Diachronic collaboration yields adaptive problem solving"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    fig_caption: true
---

Teamwork is considered effective when the team accomplishes something together that the members of the team could not have alone. One of the most common forms of teamwork involves breaking big problems into smaller problems that can be divided among the members of the team. I call this **synchronic collaboration** because all the members of the team are working, though independently, at the same time. In contrast, **diachronic collaboration** involves team members working one at a time on the same problem, each building upon the previous solution. Wikipedia article writing is an example of diachronic collaboration. Most real world collaborations have both synchronic and diachronic elements, which makes it difficult to assess the relative effectiveness of each type of collaboration. The experiments outlined below are designed to measure the problem solving capacities of teams organized either diachronically or synchronically. My hypothesis is that organizing teamwork diachronically will make teams more effective at utilizing feedback to solve problems.

```{r config, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/",
  fig.height = 4,
  fig.width = 4,
  cache = FALSE  # warning! cache is bad with graphviz * svg * read_chunk
)

library(tidyr)
library(pander)
```

```{r setup}
library(evoteams)
read_evoteams_chunks()
read_chunk("chunks/problem-spaces.R")
read_chunk("chunks/hypotheses.R")
read_chunk("chunks/time.R")
```

```{r hypotheses}
```

```{r time}
```

# Design

```{r ivs, results = "asis"}
ivs <- list(
  `Team structure` = c("diachronic", "synchronic"),
  `Feedback` = c("yes", "no", "enforce"),
  `Problem` = c("adaptive", "insight")
)
pandoc.table(ivs, caption = "Independent variables.", justify = "left")

exps <- data.frame(
  row.names = c("Team structure", "Problem", "Feedback"),
  `Exp 1` = c("diachronic/synchronic", "adaptive", "yes"),
  `Exp 2` = c("diachronic/synchronic", "adaptive", "no/enforce"),
  `Exp 3` = c("diachronic/synchronic", "insight", NA),
  `Exp 4` = c("diachronic/synchronic", "multiple adaptive", "yes")
) %>% t
pandoc.table(exps, justify = "left", missing = "",
             caption = "Overview of experiments and conditions. All manipulations are between team.")
```

Experiment 1 is designed to illustrate the predicted effect of team structure on team performance, where diachronic teams outperform synchronic teams. To investigate whether this difference is due to responsiveness to feedback, in Experiment 2 the feedback made available to the teams is manipulated, either removing it completely or enforcing monotonic improvements in team performance. The predictions are that removing feedback should impair diachronic performance more than synchronic performance, and enforcing feedback should improve synchronic performance more than diachronic performances. In Experiment 3, the impact of team structure on solving insight problems is tested. Unlike adaptive problems, for insight problems feedback is usually unhelpful, as there is not a gradation between a partial solution and a whole solution. In solving these problems, synchronic collaboration is hypothesized to be more effective than diachronic collaboration. Finally, Experiment 4 is a larger scale version of Experiment 1 that addresses the discrepancy between calendar hours and labor hours in the team structure manipulation by having both teams complete 4 problems in the same timeframe. These results as predicted stand to demonstrate the conditions under which diachronic collaboration is a more effective strategy than synchronic collaboration.

The conclusions that can be drawn from these results shed light on the relative roles of diachronic and synchronic collaboration in human cultural evolution, especially in the evolution of technology. Although our culture upholds inventors as visionaries working all alone and creating artifacts well ahead of their time, the reality is that technological progress is much more gradual than punctuated by the rare genius. One conclusion that can be drawn from this evidence of continuous improvements is that humans are especially adept at solving tinkering problems. However, this may only be true at the time scale of human generations. This research for the first time compares diachronic and synchronic collaboration as techniques for tinkering over the same time frame. If diachronic organization is more affective, it speaks to the role of diachronic collaboration in the history of human cultural evolution.

# 1 Team structure

```{r team-structure, engine = "dot", dev = "svg", fig.cap = "**Team structures.** Teams are allotted the same number of labor hours to solve a problem. In synchronic teams, teams work on a problem all at the same time. In diachronic teams, team members work one at a time, each inheriting the team's solution and improving it."}
```

In solving a problem where there are many possible solutions, all varying in degree of success, a team organized diachronically is expected to use labor hours more efficiently than a team organized synchronically. In Experiment 1, teams of four are tasked with solving a classification problem. Classification problems involve generating predicted classes for unlabeled data using statistics and machine learning to improve classification accuracy. For example, a typical classification problem is to predict whether or not a passenger on the Titanic survived based on their age, gender, ticket price, cabin, and other features available in the training data. Teams working diachronically or synchronically are tasked with building a computer program that, when run, generates predicted classes for test data that was withheld from the training set. The programs developed by diachronic teams are predicted to have higher classification accuracy on this test data at the end of the allotted labor hours than the programs developed by synchronic teams.

Classification problems have many possible solutions. Teams can use decision trees, linear regression, neural networks, or any other method to generate the predicted classes. Within each type of solution, there exist innumerable parameters that can be tweaked based on the available training data, including transformations of the input and how to weigh different variables. As varied as the different solutions are, they can all be compared in classification accuracy of the same test data. Given this clear measure of success, classification problems lend themselves easily to competitions. For example, Kaggle competitions offer cash rewards to teams with the highest classification accuracy on a particular classification problem.

In Experiment 1, participants with experience solving classification problems are recruited and randomly assigned into teams. Synchronic teams are given two hours with all four team members in the same room to develop a classifier. Each team member is provided a personal work station and network access to the team's program files, but there are no constraints on how the members of the synchronic teams interact or distribute labor over the course of the two hours. Diachronic teams work one at a time, each spending two hours on the program, and thus are never allowed to interact with the whole team at once. For diachronic teams, only a single workstation is needed, as each team member simply picks up where the last one left off.

Teams can choose to test the success of their program at any time against the test data. As feedback, teams receive their classification accuracy, but are given no information about the particular items that were misclassified. After eight labor hours have been spent by each team, the final version of each team's classifier is submitted. The primary hypothesis tested in Experiment 1 is that programs created by diachronic teams will have higher classification accuracy than programs created by synchronic teams.

```{r team-structure-results, fig.cap = "Classification accuracy by team structure. Diachronic teams are predicted to be more effective at solving adaptive problems than synchronic teams."}
```

The reason diachronic teams are more effective at solving classification problems is hypothesized to be because they are better able to iterate on solutions in response to feedback. If true, this would mean that diachronic teams should submit more solutions than synchronic teams.

```{r submissions-per-team}
```

In understanding the difference between the number of diachronic and synchronic submissions, it's important to distinguish between calendar hours and labor hours. Calendar hours are hours in the traditional sense of hours in a day. Labor hours are calendar hours multiplied by how many people are working. Both diachronic and synchronic teams are allotted 8 labor hours. While synchronic teams spend all their labor hours in the first 2 calendar hours, diachronic teams spread the labor out over 8 calendar hours.

```{r hours-per-team}
```

The type of collaboration required for this task centers on a single product: the computer program that implements a classifier. Given that the purpose of the teamwork is to collaborate on this one, single product, the number of submissions a team can provide is expected to be constrained by calendar hours more than by labor hours. The rate at which submissions are made by diachronic teams and synchronic teams is expected to be equal in terms of calendar hours. But if the rate at which submissions are made is calculated instead in terms of labor hours, the rate of submissions per labor hour for synchronic teams is expected to be much lower than for diachronic teams.

```{r submissions-per-hour}
```

It seems likely that the rate of submissions per labor hour will be greater for diachronic teams given that the submission rate is constrained more by calendar hours than by labor hours. If diachronic teams outperform sychronic teams, one of the reasons may be because they are better able to adapt a program by incorporating feedback that comes with each additional submission. The purpose of the next experiment is to manipulate the feedback provided to each team.

# 2 Feedback

```{r feedback, engine = "dot", dev = "svg", fig.width = 6, fig.cap = "Three ways of sharing individual contributions with a team of collaborators. When feedback is provided (**left**), the impact of an individual contribution on the performance of the team as a whole is made available on every contribution. When feedback is not provided (**middle**), individual contributions are shared directly with the members of the team without revealing its impact on team performance. When feedback is enforced (**right**), contributions that do not improve team performance are rejected, and cannot be shared with the members of the team."}
```

The hypothesis being tested in Experiment 2 is that diachronic teams outperform sychronic teams on adaptive problem solving because they are better able to utilize feedback to incrementally improve solutions. The type of feedback used in Experiment 1 was passive because it was given on request and in no way affected the team's collaborations. Teams in Experiment 2 either receive no feedback or feedback is enforced such that each contribution cannot be shared with the members of the team unless the contribution improves team performance.

```{r feedback-results, fig.cap = "Impact of feedback types on team performance. Taking feedback away is predicted to impair diachronic team performance more than synchronic team performance. Enforcing feedback such that every change to the code base must improve program performance is predicted to aid synchronic teams more than diachronic teams."}
```

# 3 Insight problem

On an insight problem, you can still get feedback on whether or not your solution is correct, but if you are wrong you don't know if you've gotten any closer to the correct solution.

```{r problem-spaces, fig.width = 6}
```

```{r insight-results}
```

# 4 Multiple projects

```{r multiple-projects, engine = "dot", dev = "svg", fig.cap = "Two ways of allocating the same number of calendar hours toward multiple projects. In synchronic teams, team members work on each project from beginning to completion in lockstep. In diachronic teams, team members work on different projects at the same time, and collaborate over time."}
```

```{r multiple-projects-results}
```
