---
title: "Diachronic collaboration in Kaggle competitions"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
---

```{r config, echo = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  fig.width = 3,
  fig.height = 3,
  fig.path = "figs/",
  cache.path = ".cache/"
)

library(evoteams)
library(crotchet)
library(DiagrammeR)
library(magrittr)

teams_per_condition <- 10
diachronic_team_size <- 2
n_participants <- (teams_per_condition) + (teams_per_condition * diachronic_team_size)

```

# Methods

## Participants

For this pilot experiment, participants for `r teams_per_condition` teams in each condition will be recruited, for a total of `r n_participants` participants. These participants will be recruited and screened to determine eligibility. Eligible participants are assigned conditions and contacted to schedule a time to complete the experiment.

## Procedure

Participants are randomly assigned either solo or diachronic conditions. Participants in the solo condition have 4 hours to come up with a solution. Each participant in the diachronic condition has 2 hours. Participants in the diachronic condition are randomly assigned either the first or the second generation. Second generation participants start with someone in the first generation's final solution. Second generation participants are not constrained by the solution they inherit and may start over if necessary. However, teamwork between diachronic collaborators is incentivized by awarding prizes to the best team solutions in each condition.

```{r conditions}
find_graphviz("diachronic-v-solo", package = "evoteams") %>% grViz
```

## Materials

Participants are recruited to compete in a Kaggle competition. Kaggle competitions are data-oriented challenges in which teams and individuals use statistics and machine learning to make accurate predictions about unlabeled data. For example, a standard Kaggle competition is to predict the survivors of the Titanic. Given a passenger’s ticket information, how accurately can teams predict whether or not the passenger survived? This example is illustrative of the fact that perfect performance is rarely possible in a Kaggle competition, but people can do far better than chance. The way teams improve their predictions is by using statistics and machine learning to generate more and more accurate predictions. Teams can use any method in statistics and machine learning to generate the predictions, which means there are many possible solutions to any Kaggle competition.
