---
title: "Understanding difficulty in the totems landscape"
author: "Pierce Edmiston"
output:
  html_document:
    theme: flatly
---

```{r config, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
read_chunk("../R/0-setup.R")
```

```{r setup}
# source("docs/R/0-setup.R")
```

Difficulty is a factor of inventory size--the number of items available for
recombination--and the number of adjacent items that could be discovered. If
difficulty were only a factor of inventory size, the game would get
exponentially more difficult as you went on. However, that doesn't seem to be
the case. What matters more than the total number of combinations that could be
made is the number of adjacent items that could be discovered.

```{r}
data("Combinatorics")

Combinatorics %<>%
  tidyr::gather(Type, Count, -InventorySize)

ggplot(Combinatorics) +
  aes(InventorySize, Count, color = Type) +
  geom_point()
```

```{r}
data("InventoryInfo")

Difficulties <- InventoryInfo %>%
  select(InventorySize, GuessDifficulty, CombnDifficulty) %>%
  tidyr::gather(Type, Difficulty, -InventorySize)

ggplot(Difficulties) +
  aes(InventorySize, Difficulty, color = Type) +
  geom_point(alpha = 0.5) +
  facet_wrap("Type")
```
