---
title: "The impact of inheritance on problem solving"
output: pdf_document
abstract: |
  Humans are able to accumulate tools and other solutions to problems like no
  other species of animal. A consequence of our system of cultural inheritance
  is that future generations rarely have to start from scratch when solving new
  problems, as demonstrated by the prevalence of reuse in the history of human
  technology. We examined how inheriting solutions from a previous
  generation impacted problem solving by comparing the performance of teams who
  share solutions across generations to alternative strategies for solving the
  same set of problems but without inheritance. We found that inheritance was
  effective in that participants were reliably able to solve more problems than
  their ancestors, and that the benefits of inheritance were similar to
  individuals given the same total amount of time and teams working in
  parallel. We also found that inheritance caused participants to engage in
  a more efficient guessing strategy, and as a consequence were less likely to
  become fixated on incorrect solutions. These results suggest that human
  cultural inheritance does more than give future generations a head start on
  problem solving, and may actually change the way new problems are solved.
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  dpi = 300,
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "../R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)

# Read all the diagrams made with graphviz in the totems package
chunks <- c(
  "team-structures",
  "team-structures-all",
  "team-structures-exp1",
  "team-structures-exp2",
  "team-structures-exp3",
  "team-structures-exp4"
)
sapply(chunks, read_graphviz_chunk, package = "totems")
```

# Experiment 1

```{r team-structures-exp1, engine="dot", out.width="60%"}
```

- Participants are likely to exceed their ancestors.
- Some evidence that the likelihood of exceeding decreases for later generations and with larger inheritances.
- However, due to the properties of the landscape, a better test is to compare to alternative strategies (Exp. 2-4).

# Experiment 2

```{r team-structures-exp2, engine="dot", out.width="40%"}
```

- Inheritance is not less efficient than Isolated (50min) or Synchronic (2-person).
- Some evidence that inheritance results in a less redundant guessing strategy.

# Experiment 3

```{r team-structures-exp3, engine="dot", out.width="60%"}
```

- Inheriting from oneself is generally better than inheriting from someone
  else, but this is due to faster time in recreating inherited items, not
  necessarily in improvements to guessing efficiency.
- Isolated individuals tend to get fixated more than diachronic individuals.

# Experiment 4

```{r team-structures-exp4, engine="dot", out.width="60%"}
```

- As synchronic teams add individuals, the team gets less efficient,
  and this effect is less pronounced in diachronic individuals.