---
title: "Guessing strategy"
output: bookdown::html_document2
---

```{r config, include = FALSE}
library(knitr)
library(printr)
library(magrittr)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 4,
  fig.height = 3,
  cache = TRUE
)

list.files("../R", "*.R", full.names = TRUE) %>% sapply(read_chunk)
```

```{r setup, include=FALSE}
```

```{r methods, include=FALSE}
```

```{r exp1, include=FALSE}
```

```{r exp1-extra, include=FALSE}
```

Each guess can be classified in one of four ways (Table \@ref(tab:types-of-guesses)). Guesses that had never been made by the participant before were unique item guesses if they created a new item, or simply a unique guess if the guess did not create any item. Guesses that had been made before either could repeat a previously created item, or they could be redundant guesses. We refer to the distribution of guesses
in these four categories as the guessing strategy.

(ref:types-of-guesses) Four kinds of guesses. Redundant guesses were incorrect guesses that had already been made. Repeat item guesses recreated a previously discovered item. Unique guesses had not been tried before, but did not result in a new item. Unique item guesses were unique guesses that resulted in the creation of a new item.

```{r types-of-guesses, results="asis"}
recode_guess_type() %>%
  select(GuessTypeLabel, UniqueGuess, CreatedItem, UniqueResult) %>%
  kable(caption = "(ref:types-of-guesses)")
```

To investigate the impact of inheritance on problem solving ability, we compared the
guessing strategy of problem solvers without inheritance (generation 1 problem solvers)
to the guessing strategy of problem solvers with inheritance. We found that
diachronic inheritance resulted in fewer redundant guesses than problem solving
without inheritance, `r report_lm_mod(redundant_guesses_mod, "Diachronic_v_NoInheritance")`.

Problem solving with inheritance did not result in an increase in the proportion of
unique guesses, `r report_lm_mod(redundant_guesses_mod, "Diachronic_v_NoInheritance")`.

```{r guess-types-prop, fig.width=6}
guess_types_prop_plot
```