---
title: "Trajectories through the totems landscape"
output: bookdown::html_document2
---

```{r config, include = FALSE}
library(knitr)
library(printr)
library(magrittr)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 4,
  fig.height = 3,
  cache = TRUE
)

list.files("../R", "*.R", full.names = TRUE) %>% sapply(read_chunk)
```

```{r setup, include=FALSE}
```

```{r methods, include=FALSE}
```

```{r exp1, include=FALSE}
```

```{r exp1-extra, include=FALSE}
```

(ref:diachronic-player-trajectories-plot) Rate of discovery in learning and
playing stages. Each line is a player. In the learning phase (left of midline), problem
solvers are recreating the innovations created by their ancestor. After having
recreated the inventions of their ancestors, the problem solvers enter the
playing stage (right of midline), where they are tasked with discovering new solutions not
discovered by their ancestor.

```{r diachronic-player-trajectories-plot, fig.width=6, fig.cap="(ref:diachronic-player-trajectories-plot)"}
diachronic_player_trajectories_plot
```
