---
title: Cumulative improvements in iterated problem solving
author:
  - name: Pierce Edmiston
    email: pedmiston@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - name: Maxime Derex
    email: maxime.derex@gmail.com
    department: Biosciences Department
    university: University of Exeter
  - name: Gary Lupyan
    email: lupyan@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
    last: yes
bibliography: references.bib
abstract: |
  As compared to other animals, humans are particularly skilled at using and
  improving tools and other solutions to problems that were first discovered by
  other people. Although the human capacity for cumulative cultural evolution
  is well-known, the effectiveness of inheritance as a form of problem solving
  is an area in need of further research. We report an experiment designed to
  understand how effectively problem solving knowledge accumulates over
  generations of problem solving. We found that problem solvers were
  consistently able to solve more problems in a 25 minute session than their
  ancestors. In investigating whether the number of inherited solutions had an
  impact on future problem solving performance, we found that larger
  inheritances were harder to exceed, but they also allowed individuals to
  solve more complex problems than they would have solved on their own. We
  discuss the limitations of this work, and motivate future directions.
keywords: "cultural evolution; transmission chain; iterated learning"
output:
  pdf_document:
    template: templates/cogsci.tex
    citation_package: natbib
    keep_tex: yes
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  fig.crop = F,
  fig.pos = "tb",
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)
```

```{r setup, include=FALSE, cache=FALSE}
```

```{r methods, include=FALSE}
```

```{r exp1, include=FALSE}
```

# Introduction

Humans are effective problem solvers, having solved a wide range of problems
related to foraging, hunting, and preparing food, while surviving predators,
each other, and a large range of terrestrial environments [@Boyd:2017diff;
@Fernandez:2001civ]. What has enabled our success in being able to solve such
a diverse set of problems? Some have suggested that the answer lies more in our
ability to inherit knowledge from others than our ability to solve problems as
individuals [@Richerson:2008not; @Henrich:2015srt; @Boyd:2017diff]. Cultural
inheritance enables problem solving to extend over generations, far longer than
any single lifetime, allowing future generations of problem solvers to benefit
from solutions to problems that they could not have discovered on their own.
However, what is unclear is how effectively individual problem solvers are
able to benefit from inherited solutions at shorter timescales. To
better understand the extent to which future generations of problem solvers are
able to benefit from inherited problem solving knowledge, we investigated the
accumulation of problem solving knowledge over generations of problem solving
(Fig. 1).

```{r fig1, fig.pos="b", fig.align="center", fig.width=3.5, fig.cap="Iterated problem solving paradigm. Participants were assigned to generations within chains. Each participant completed the same problem solving task for 25 minutes. Participants in generations after the first began the problem solving task with the solutions that were discovered by the previous generation."}
draw_graphviz("iterated-problem-solving", package = "totems")
```

The accumulation of problem solving knowledge is thought to be enabled by
a number of social learning abilities that facilitate transmission of problem solving knowledge with high fidelity across individuals. These include
teaching through verbal instruction, prosociality, and advanced imitation
abilities [@Dean:2012gv]. These skills for high fidelity social learning are
argued to benefit problem solvers by providing a reliable shortcut to
individual (asocial) learning [@Boyd:1985cu]. Rather than having individuals
struggle to solve a problem on their own through trial-and-error learning,
social learners can inherit solutions that have already been discovered by
others. Over generations, these shortcuts accumulate, giving populations
of social learners to get a head start on future problem solving, and allowing
them to solve problems that no individual learner would be able to solve on
their own and in a single lifetime.

However, the ability to learn socially from others is not sufficient to explain
cumulative cultural evolution. Although social learning was once thought to be
rare in the animal kingdom [e.g., @Thorndike:1898], it has now been documented
in a range of species from chimps [@Whiten:1999cic] to fish [@Laland:1997ts]
and even bees [@Alem:2016fu]. If cumulative cultural evolution depended simply
on social learning, we might expect these species to likewise show evidence of
cumulative cultural evolution, yet such evidence is notably lacking
[@Dean:2012gv; @Tennie:2009gaa]. There are a few exceptions: New Caledonian
crows have been argued to have improved the grub-skewers they craft from
pandanus leaves [@Hunt:2003dj], and at least one population of chimpanzees has
evolved a more advanced termite probe [@Sanz:2009ja], but for the most part
tool use in animals has remained largely unchanged [cf. @Mercader:2007fd].

Humans, in contrast, have demonstrated a remarkable ability to reproduce,
adapt, and improve the innovations of others, so much so that the history of
human technology has been argued to be better understood as a process of
continual refinement and repurposing rather than punctuated advances brought by
the discoveries of rare geniuses [@Basalla:1989wq; @Sole:2013cm]. Rapid
refinement of inherited innovations has not always been the case over the
course of human history, as demonstrated by the long periods in the
archaeological record of slow or stagnant growth in stone tool complexity
[@Torre:2011jl; @Lycett:2008di], raising the question of what it was that
allowed our ancestors to begin to develop more complex tools. An answer that
has been proposed is that humans evolved more robust ways of transmitting
problem solving knowledge to the next generation, allowing future generations
of problem solvers to more quickly learn the skills honed by their ancestors,
and thus giving them more time to make improvements to those technologies. This
raises the question of how effectively human problem solvers are able to
accumulate problem solving knowledge over generations of problem solving. Is
cumulative cultural evolution only visible at the levels of populations and
human lifetimes, or are human problem solvers able to accumulate problem
solving knowledge effectively even at shorter timescales?

Previous investigations of the accumulation of problem solving knowledge via
vertical transmission have found that problem solvers benefit the most from
previous generations when high fidelity transmission mechanisms such as verbal
instruction are available [@Morgan:2015hu]. However, although verbal
instruction improves the fidelity of the transmission process, vertical
transmission alone is not guaranteed to result in the cumulative improvement in
problem solving performance. What are the minimal requirements for observing
reliable improvements in problem solving knowledge over generations?

Here we investigated the accumulation of problem solving knowledge with high
fidelity transmission in place, such that the solutions generated by one
generation are precisely transmitted to the next generation. Our experiment was
designed to test the hypothesis that high fidelity vertical transmission is
sufficient to allow for the accumulation of problem solving knowledge. More
importantly than simply demonstrating that problem solving knowledge _can_
accumulate, we also sought to determine how effectively future generations can
capitalize on the advancements of the previous generation. Finally, we also
sought to understand whether inheriting solutions from a previous generation
had any influence on future problem solving ability. Does the benefit to
inheriting solutions diminish as the size of the inheritance increases,
or does inheritance provide a reliable improvement in problem solving
performance regardless of size?

# Methods

To understand how problem solving knowledge accumulates through vertical
transmission, we used a transmission chain paradigm where participants were
assigned to a single generation within a four-generation chain. Each
participant attempted the same problem solving task for 25 minutes. The
solutions that each participant had discovered by the end of the session were
passed on to be inherited by a participant in the next generation of the
chain. Thus, participants assigned to generations after the first began the
problem solving task with solutions inherited from the previous generation.

Participants played the "Totems" game adapted from @Derex:2015cbb. Their task
was to discover how to build tools with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. 2). More advanced tools produce larger and more
intricate totems, resulting in higher performance scores.

```{r fig2, fig.height=6, fig.cap="A sample of the solution landscape. The top row of 6 items were available to problem solvers at the start of the game. New items could be produced through the combination of different items (more than one arrow points to the item) or through the refinement of a single item (a single arrow points to the item). The axe is required to construct the first totem pole."}
draw_image("landscape-sample.gv", package = "totems")
```

Participants discovered new tools by combining existing items. Participants
could combine up to four items at a time (with replacement), meaning the
initial six items could form a total of `r exp1$n_unique_guesses_6`
combinations. Of all possible combinations, very few resulted in new items. For
example, of all the guesses that could be formed from the initial items, only
three (`r exp1$n_unique_guesses_6_pct`%) yielded new tools (Fig. 2). As
participants accumulate solutions, the combinatorial complexity of the problem
space increases exponentially such that the discovery of more complex tools is
less likely to happen by chance.

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record. Participants could review their past innovations and see the
recipes for their previous discoveries. **Participants assigned to generations
after the first inherited the innovation record of the previous generation
participant.** From the beginning of the experiment, these participants could
review the recipes for all the innovations that had been discovered by their
ancestor. Note that the participants inherited the recipes, but not the tools
themselves. In order use these tools in futher combinations, the tools and all
of their constituent parts first had to be recreated.

## Participants

Participants were recruited from the UW-Madison student body and received
course credit in exchange for participation. Each participant was assigned to
a single generation of a four-generation chain. Data was collected
for a total of `r exp1$n_teams` teams (N=`r exp1$n_participants` participants).

# Results

Participants who inherited the tools discovered by the previous generation
took on average `r exp1$mean_learning_time_min` minutes of the 25 minute
session (`r exp1$proportion_learning_time`%) to recreate the inherited tools---
a portion of the experiment we refer to as the learning period. The length of
the learning period correlated positively with the number of inherited tools,
`r report_lm_mod(learning_rates_mod, "InheritanceSize")` (Fig. 3). A small
group of participants (N=`r exp1$n_outliers`) took disproportionately long to
recreate the tools they inherited perhaps because they were confused by the game
interface. These participants are identified by X's in all figures and are not
included in any statistical analyses.

```{r fig3, fig.cap="Learning rates. Correlation between the number of tools inherited and the time it took to recreate the inherited items. Outliers who were unwilling or unable to recreate the inherited items are shown as X's, and excluded from analysis."}
learning_rates_plot
```

After recreating all inherited tools, participants attempted to
discover new tools not discovered by their ancestor. Most participants
(`r exp1$pct_did_exceed`%) were able to exceed the number of tools that they
inherited `r exp1$logodds_of_exceeding` (Fig. 4). Using Page's trend test (a
repeated measures test for monotonicity) we found that iterated problem solving
resulted in cumulative improvements in the number of tools that could be
discovered in 25 minutes along each transmission chain, `r exp1$page_test`
(Fig. 5).

```{r fig4, fig.cap="Number of innovations discovered by each generation. Each of the thin green lines is a chain. The thick blue line is the model predictions with ±1 standard error."}
innovations_by_generation_plot
```

We next tested whether the benefit to inheritance changed over
generations. We fit a hierarchical regression model to the number of
tools discovered in each generation with these effects nested by chain. On average, second generation participants were able to
discover `r exp1$gen0_slope` more tools than first generation participants,
`r exp1$gen0_slope_stats`. This effect decreased by `r exp1$gen0sqr_slope`
each generation for third and fourth generation participants,
`r exp1$gen0sqr_slope_stats`---a diminishing return. In line with
this result, we also found that an increase in the
number of inherited tools was associated with a decrease in the
number of new tools discovered by future generations of problem solvers,
`r exp1$inheritance_size_slope_stats` (Fig. 3B). Together, these findings
suggest that inheritance has a measurable influence on problem solving beyond
providing a shortcut to individual learning. In this case, the impact
appears to be negative, such that inheriting more solutions appears to
_decrease_ problem solving ability.

```{r fig5, fig.height=6, fig.cap="A. Number of tools created relative to those inherited. The dotted line is a reference with slope=1 such that points above the line indicate future generations exceeding their ancestors. B. Number of new tools relative to those inherited. The same reference line is now shown horizontally. The error range specifies the model predictions ±1 standard error. The thin gray line shows the regression line when outliers, shown as X's, are included."}
grid.arrange(
  innovations_created_and_inherited_plot + ggtitle("A"),
  new_innovations_plot + ggtitle("B"),
  ncol = 1
)
```

However, an alternative explanation, not having to do with inheritance
negatively affecting problem solving ability, is that as more tools are
accumulated, additional tools are less likely to be discovered by chance,
meaning that problem solving might slow down regardless of what was
inherited. To control for combinatorial complexity, we created an alternative
outcome measure that, rather than counting all tools equally, instead
weighted each tool relative to the number of guesses that could be
made out of the items present at the time at which the tool was discovered.
Fig. 5 shows the change in accumulated complexity score based on the number of
tools that were inherited. In contrast to the analysis that treated all
tools equally, when tools are weighted by combinatorial complexity,
inheriting more tools was associated with an ever-larger improvement in
total complexity score than inheriting fewer tools,
`r exp1$delta_diff_slope_stats`.

```{r fig6, fig.cap="Change in accumulated complexity score by inheritance. Accumulated complexity scores are sums of discovered tools weighted by the combinatorial complexity of the possible choices available to the participant when the tool was discovered. The outcome measure is the difference between the accumulated complexity scores of subsequent generations. The line shows the predictions of a hierarchical regression model with ±1 standard error."}
delta_difficulty_plot
```

Participants who inherited more tools also took longer to recreate
those tools (Fig. 2), leaving less time to discover new tools,
To determine whether inheriting more tools had an impact on
the rate at which new tools were discovered, we first report
overall rates of new tool discovery, and then test whether
the discovery rate is influenced by the number of inherited tools.

Fig. 6 shows problem solving rates based on the
length of the discovery period. The overall rate of problem solving 
was `r exp1$minutes_per_new_innovation` minutes per new
innovation (`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based on the
number of inherited tools, as revealed by a model comparison comparing a model
predicting unique innovations from playing time alone to one predicting unique
innovations from the interaction between playing time and inheritance size,
`r exp1$playing_time_modcomp`. This suggests that the number of
inherited tools did not meaningly influence the rate of future discovery.

```{r fig7, fig.env="figure", fig.cap="Problem solving rate in the discovery stage. Discovery time is the amount of time out of a 25 minute session dedicated to discovering new innovations that were not discovered by an ancestor. The line shows the predictions of the hierarchical regression model with ±1 standard error. The slope of this line did not significantly vary based on the size of the inheritance."}
playing_time_plot
```

# Discussion

The ability to accumulate a body of socially learned problem solving knowledge
is visible in all human cultures [@Boyd:2017diff; @Fernandez:2001civ]. Although
the capacity for cumulative improvement is believed to be integral to human
cultural evolution, we do not fully understand how effectively problem solving
knowledge can be utilized by future generations of problem solvers. How
effective are human problem solvers at ratcheting up the tools and other
solutions to problems that were discovered by previous generations? To answer
this question, we investigated the ability of problem solvers to accumulate
problem solving knowledge over four generations of iterated problem solving.

In contrast to previous research that has investigated how social learning
strategies such as verbal instruction play a role in high fidelity transmission
of problem solving knowledge, here we examined the likelihood that problem
solving knowledge would accumulate given that problem solving knowledge could
be effectively transferred to the next generation. Participants in our
study inherited the solutions that had been discovered by
their ancestor, and we measured their ability to recreate and extend
this body of problem solving knowledge.

We found that problem solvers were consistently able to solve more problems in
a single 25 minute session than their ancestors, and thus were able to
cumulatively improve upon the solutions they inherited. All participants
were expected to be able to recreate the tools they inherited, but whether
they could discover new tools, beyond those inherited, was unknown.
Given the combinatorial complexity of the solution landscape, participants
were highly unlikely to strike upon a beneficial combination by chance
alone. Because of this, some participants were unable to discover any new tools, but most did discover new tools, even
when inheriting an already large number of previously discovered tools. 

We also explored the impact of inherited solutions on future problem solving
performance. Here our results were mixed. We found that problem solvers who
inherited more tools tended to discover fewer new tools than their ancestors,
suggesting that inheritance has a diminishing return on future problem solving
performance. However, if the combinatorial complexity of the solutions is taken
into account, we found that these problem solvers were able to accumulate more
complex tools than they would have been likely to discover on their own. In
addition, we found that controlling for the amount of time each participant had
to discover new tools (as opposed to recreating inherited tools) failed to
reveal an effect of inheritance size on the rate of problem solving.

Our conclusions are limited by the nature of the solution landscape in the
Totems game, and the restriction in our methods to a single problem solving
strategy. The sparsity of the solution landscape, where many thousands of
possible combinations can be made but far less than 1% yield new tools,
indicates that participants must use tacit knowledge to help form combinations
that are most likely to yield new tools. For example, once an axe was
discovered, its combination with the tree to create a totem was very likely.
This challenges the notion that the difficulty of a particular tool is directly
related to its combinatorial complexity. In addition, we believe the
accumulation of problem solving knowledge through vertical transmission must be
compared with the accumulation of problem solving knowledge through other forms
of problem solving that do not involve vertical transmission. We are 
conducting additional experiments on the Totems game with isolated individuals
who play the game for the same total time as multi-generation chains, as well
as groups of problem solvers who play the Totems game in parallel, and thus
share information through horizontal transmission. Results indicate that
problem solving knowledge accumulated over generations is as effective
as assembling problem solving knowledge using these alternative forms
of collaboration.

More than any other animal, humans are particularly skilled at inheriting and
improving tools and other solutions to problems, but whether the ability
to inherit from others has effects on problem solving beyond giving a head
start to individual learning is not known. Although much work is still
needed to fully understand the human propensity for cumulative cultural
evolution, we believe our research is a valuable contribution to ongoing
efforts to understand how and why human culture is so integrally cumulative.

\setlength{\parindent}{-0.1in}
\setlength{\leftskip}{0.125in}
\noindent
