---
title: Cumulative improvements in iterated problem solving
author:
  - name: Pierce Edmiston
    email: pedmiston@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - name: Maxime Derex
    email: maxime.derex@gmail.com
    department: Biosciences Department
    university: University of Exeter
  - name: Gary Lupyan
    email: lupyan@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
    last: yes
bibliography: references.bib
abstract: |
  Humans are able to use and improve solutions to problems including tools and
  other technology that were inherited from others like no other species of
  animal. What is the impact of inheritance---starting off a problem solving
  task with partial solutions inherited from someone else---on future problem
  solving ability? Does inheritance provide a reliable shortcut to individual
  learning without affecting underlying problem solving ability, or does
  inheritance change the way future problems are solved? To begin to answer
  these questions, we measured changes in problem solving ability over four
  generations of iterated problem solving, where the solutions discovered by
  one generation are inherited to be used and improved by the next generation.
  We found that through four generations of iterated problem solving, problem
  solvers were consistently able to solve more problems in a single 25 minute
  session than their predecessors, even as problems became more difficult to
  solve. We investigated whether the size of the inheritance had an impact on
  future problem solving ability, and found mixed results: larger inheritances
  were harder to exceed, but they also allowed individuals to solve more
  difficult problems than they would have on their own. We also found no
  evidence that a larger inheritance had a negative impact on the rate at which
  new problems were solved. We discuss the limitations of this work, and
  motivate future directions.
keywords: "cultural evolution; transmission chain; iterated learning"
output:
  pdf_document:
    template: templates/cogsci2016.tex
    citation_package: natbib
    keep_tex: yes
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  fig.crop = F,
  fig.pos = "tb",
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)
```

```{r setup, include=FALSE, cache=FALSE}
```

```{r methods, include=FALSE}
```

```{r exp1, include=FALSE}
```

# Introduction

Humans are effective problem solvers, having solved a wide range of problems
related to foraging, hunting, and preparing food, while surviving predators,
each other, and the harsh environment in nearly every terrestrial environment
on the planet [@Boyd:2017diff; @Fernandez:2001civ]. What has enabled our
success in being able to solve such a diverse set of problems as posed by
environments ranging from deserts to rainforests, from high altitudes to low?
One answer to this question is that our system of cultural inheritance, more
than our ability to solve problems as individuals, explains how humans are able
to solve the problems necessary for survival in diverse habitats
[@Richerson:2008not; @Henrich:2015srt; @Boyd:2017diff]. As problem solvers,
humans are particularly adept at using, reproducing, and, in some cases, improving
cultural products including tools and other solutions to problems that were
first created by other people, a phenomena generally known as the ratchet
effect [@Tomasello:1993vh; @Tennie:2009gaa]. I refer to the type of
collaboration that enables ratcheting as **diachronic collaboration** (Fig. 1).
In diachronic collaboration, the products of labor, such as a tool designed to
solve some problem, are shared with or inherited by future generations of
problem solvers. The purpose of this research is understand why
diachronic collaboration may be crucial to human cultural evolution by
assessing its effectiveness as a problem solving strategy.

```{r fig1, fig.pos="b", fig.align="center", fig.width=3.5, fig.cap="Iterated problem solving. Participants were assigned to four-person teams. Each participant completed the same problem solving task for 25 minutes. Participants in generations after the first began the problem solving task with the solutions that were discovered by the previous generation."}
draw_graphviz("iterated-problem-solving", package = "totems")
```

The inheritance of tools and other solutions to problems occurs through social
learning, or what @Thorndike:1898 called "from an act witnessed, learn to do an
act" (p. 50). Thorndike's failed experiments on social learning in chicks, cats,
and dogs made him and many others skeptical that social learning was a
widespread phenomenon among animals, but since then, a growing number of species
have been argued to show evidence of the ability to inherit skilled behaviors
from the observation of others. This includes ethnographic evidence of cultural
traditions sustained through social learning in primates, monkeys, cetaceans,
and birds, as well as experimental evidence of social learning in some of the
same species and others, including rats, fish, and even some insects. Given that
social learning is present in such diverse taxa, it is surprising that only a
few species (namely crows [@Hunt:2003dj], chimpanzees [@Sanz:2009ja], and
perhaps pigeons [@Sasaki:2017ha]) have demonstrated a capacity for cumulative
improvements in inherited innovations. In contrast, the ability to accumulate
beneficial variants in solutions to problems is endemic to human culture and
specifically the history of human technology [@Basalla:1989wq]. As a result,
many researchers have concluded that the difference between human social
learning and social learning exhibited in other animals is not qualitative, but
the gap is quantitatively vast.

Cultural inheritance via social learning mechanisms is believed to benefit
learners by providing a reliable shortcut to individual learning.
Computational models of the cultural evolutionary process show how social
learning can be evolutionarily advantageous as compared to investing in more
advanced individual learning capacities when the environment is stable and
individual learning is costly [@Boyd:1985cu]. Under these conditions, the
ability to copy the behaviors of others saves future generations the time
already invested by the previous generations, allowing cultures to discover
innovations that no individual would be likely to discover on their own
[@Derex:2015cbb]. Additional modeling work has demonstrated the importance of
being able to filter among possible variants in order to reliably accumulate
beneficial variants [@Enquist:2010gw]. However, this view presupposes
that behaviors learned socially are functionally equivalent to behaviors
learned through individual learning. Successful transmission either
happens and future generations pick up directly where the previous generation
left off, or it does not and future generations are unable to benefit from the
progress of the previous generation. But inheritance is not believed
to impact underlying problem solving ability, only the starting point
for future problem solving.

<!--
Justify why inheritance may affect problem solving ability.
* Cultural transmission is a process of transformation, not replication [@Sperber:1996exc].
* Mentoring and creating a cultural niche for problem solving [@Sterelny:2012ea].

Explain the goals of the experiment.
* Use an iterated learning paradigm to measure the effectiveness of problem solving
  organized around inheritance.
* Investigate whether the solutions that were inherited impact future problem
  solving against the null hypothesis that problem solving is constant,
  and unaffected by inheritance.
-->

# Methods

We used a iterated learning paradigm where participants are assigned to
generations within chains [@Caldwell:2008gf]. To emphasize that participants in
each chain are working together cooperatively on a problem solving task rather
than engaging in unguided repetition [cf. @Bartlett:1933remembering;
@Kirby:2008kja; @Edmiston:2017jx] we refer to the transmission chains instead
as diachronic teams. We tested the problem solving ability of diachronic teams
over four generations of problem solving, where each generation attempted the
same problem solving task for 25 minutes. Diachronic problem solvers after the
first generation began the problem solving task with the solutions created by
the previous generation.

The problem solving task that participants completed was adapted from
@Derex:2015cbb. In this computer game, called the Totems game, participants are
required to build tools with the ultimate goal of creating "a sacred totem to
appease the gods." To build a totem, participants first needed to construct an
axe out of three independently discovered tools: a refined stick used as
a handle, a sharpened rock for the blade, and a string wound from bark fibers
for binding (Fig. 2B). In the game, more advanced tools produce larger and more
intricate totems, resulting in higher performance scores. Diachronic problem
solvers after the first generation began the problem solving task with access
to the solutions discovered by the previous generation.

## Participants

Participants were recruited from the UW-Madison student body and received
course credit in exchange for participation. Each participant was assigned to
a single generation of a four-person team (Table 1). Data was collected for
a total of `r exp1$n_teams` teams (N=`r exp1$n_participants` participants).

```{r tab1, results="asis"}
Exp1Participants %>%
  kable(caption = "Participants.", format = "latex")
```

## Procedure

Participants played the Totems game for 25 minutes. To play the game,
participants generated guesses (combinations of items) by dragging items into
the Workshop panel and selecting "Try" (Fig. 2A). Guesses were between one and
four items in length with replacement, meaning the initial six resources could be
combined for a total of `r exp1$n_unique_guesses_6`  unique guesses. Of the
possible combinations, very few resulted in new items. For example, of all the
guesses that could be formed from the initial items, only three
(`r exp1$n_unique_guesses_6_pct`%) yielded new tools (Fig. 2B). As problem solvers
accumulate solutions, the combinatorial complexity of the problem space increases
exponentially such that the discovery of more complex tools is less likely to
happen by chance.

Once a tool was discovered, the recipe for its creation was recorded in the
History panel (Fig. 2A). Participants could review their history and see the
recipes for their previous discoveries, even items they had discarded. The
History panel was also how diachronic inheritance was implemented. Diachronic
problem solvers after the first generation began the problem solving task with
an additional tab in the History panel that, when selected, listed all of the
recipes for the items created by the previous generation.

```{r fig2, fig.height=6, fig.cap="A sample of the solution landscape. The axe is required to construct the first totem pole."}
draw_image("landscape-sample.gv", package = "totems")
```

# Results

Participants who inherited the solutions discovered by the previous generation
took on average `r exp1$mean_learning_time_min` minutes of the 25 minute
session (`r exp1$proportion_learning_time`%) to recreate the inherited
tools, a portion of the experiment we refer to as the learning stage. The
length of the learning stage correlated with the size of the inheritance,
`r report_lm_mod(learning_rates_mod, "InheritanceSize")` (Fig. 3), but there
were some exceptions: some participants (N=`r exp1$n_outliers`) took
disproportionately long to recreate the tools they inherited, suggesting they
were unwilling to participate in the experiment or that they did not understand
how to use the game interface. These participants are identified by X's in all
figures, but are not included in any statistical analyses.

```{r fig3, fig.cap="Learning rates. Correlation between the number of innovations inherited and the time it took to recreate the inherited items. Outliers who were unwilling or unable to recreate the inherited items are shown as X's, and excluded from analysis."}
learning_rates_plot
```

After completing the learning stage, participants then entered the
discovery stage, where new solutions not discovered by any
predecessor were attempted. Most participants (`r exp1$pct_did_exceed`%)
were able to exceed the number of innovations that they inherited
`r exp1$logodds_of_exceeding` (Fig. 4). Using Page's trend test,
which is a repeated measures test for monotonicity, we found
that iterated problem solving resulted in cumulative improvements
in the number of problems that could be solved in 25 minutes
`r exp1$page_test` (Fig. 5).

```{r fig4, fig.cap="Number of innovations created relative to those inherited. The dotted line is a reference with slope=1 such that points above the line indicate future generations exceeding their ancestors."}
innovations_created_and_inherited_plot
```

```{r fig5, fig.cap="Number of innovations discovered by each generation. Each of the thin green lines is a team. Means in each generation are shown as green X's. The thick blue line is the model predictions with ±1 standard error."}
innovations_by_generation_plot
```

We next measured the linear increase in innovations over generations
within each team by fitting a hierarchical regression model to the total
innovations achieved in each generation with nested effects (intercepts
and slopes) for teams. On average, second generation participants were
able to discover `r exp1$gen0_slope` more innovations than first generation
participants, `r exp1$gen0_slope_stats`. This linear effect decreased by
`r exp1$gen0sqr_slope` each generation for third and fourth generation
participants, `r exp1$gen0sqr_slope_stats`. A model comparison of hierarchical
regression models fit with linear and quadratic components revealed
that a quadratic fit was significantly better than a linear fit alone,
`r exp1$quad_modcomp`. This result suggests that accumulated inheritance had a
diminishing return on future problem solving.

We next consider explanations for the diminishing return of diachronic
inheritance on problem solving. First, we tested whether larger inheritances had
a detrimental effect on future problem solving. We found that an increase in the
number of inherited innovations is negatively associated with a decrease in the
number of new innovations discovered by future generations of problem solvers,
`r exp1$inheritance_size_slope_stats` (Fig. 4). This finding preliminarily
suggests that inheritance has a measurable influence on problem solving beyond
simply providing a shortcut to individual learning. In this case, the impact
appears to be negative, such that inheriting more solutions appears to
_decrease_ problem solving ability.

However, an alternative explanation, not having to do with inheritance
negatively affecting problem solving ability, is that as more items are
accumulated, further items are less likely to be discovered by chance,
meaning we would expect problem solving slow down regardless of what was
inherited. To control for combinatorial complexity, we created an alternative
outcome measure that, rather than counting all innovations equally, instead
weighted each innovation relative to the number of guesses that could be
made from the inventory size at the time at which the innovation was discovered.

Fig. 6 shows the change in accumulated difficulty score based on the number of innovations
that were inherited. In contrast to the analysis that treated all
innovations equally, when innovations are weighted by combinatorial complexity,
the effect of inheritance does not decrease over time, such that
larger inheritances result in higher accumulated difficulty scores,
`r exp1$delta_diff_slope_stats`.

```{r fig6, fig.cap="Change in accumulated difficulty score by inheritance. Accumulated difficulty scores are sums of discovered innovations weighted by the combinatorial complexity of the possible choices available to the participant when the innovation was discovered. The outcome measure is the difference between the accumulated difficulty scores of subsequent generations. The line shows the predictions of a hierarchical regression model with ±1 standard error."}
delta_difficulty_plot
```

The final explanation we consider regarding whether inheritance affected problem
solving ability is that larger inheritances take more time to recreate, thus
leaving less time for future problem solving. Even if inheritance did not affect
problem solving ability, participants who inherit more innovations may not
discover as many future innovations simply because they used up most of their
time recreating the items that were inherited. To investigate, we looked at
problem solving rates controlling for the amount of time spent recreating the
inherited items.

Problem solvers who spend more time in the learning stage conversely have less
time in the **playing stage**, where new innovations are tried and discovered.
Fig. 7 shows the problem solving
rates for diachronic problem solvers based on the length of the playing stage.
The overall rate of problem solving in the playing stage was
`r exp1$minutes_per_new_innovation` minutes per new innovation
(`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based
on the size of the inheritance, as revealed by a model comparison comparing
a model predicting unique innovations from playing time alone to one predicting
unique innovations from the interaction between playing time and inheritance size,
`r exp1$playing_time_modcomp`. This result suggests that inheriting more
items did not have an effect on the rate at which new problems are solved
when playing time is controlled.

```{r fig7, fig.env="figure", fig.cap="Problem solving rate in the playing stage. Playing time is the amount of time out of a 25 minute session dedicated to discovering new innovations that were not discovered by an ancestor. The line shows the predictions of the hierarchical regression model with ±1 standard error. The slope of this line did not significantly vary based on the size of the inheritance."}
playing_time_plot
```

# Discussion

We measured the problem solving ability of four-person diachronic teams, and
found that diachronic problem solvers were consistently able to cumulatively
improve upon the solutions they inherited. Although this finding is not
surprising, given that participants basically started of a problem solving task
with some of the problems already solved, it is exactly this ability that is
believed to be rare among non-human animals, and fundamental to human cultural
evolution.

In our experiment, problem solvers who benefited from inheritance first had to
recreate all of the inherited solutions, which took time and effort. Once
recreated, some participants were unable to solve any new problems. Others were
unable even to recreate all of their inherited solutions in the 25-minute
session---although this behavior is likely attributable to non-compliant
behavior or a misunderstanding of the task. Nonetheless, the presence of these
outliers highlights that in our experiment, performance is not guaranteed to
improve cumulatively, which is why it is important to establish that it does.

We also attempted to determine whether inheritance had a measurable impact
on future problem solving ability. We primarily tested this question in
this experiment by investigating the relationship between inheritance size
and the number of unique innovations contributed by each generation. We
found multiple, conflicting results. The impact of inheritance on problem
solving appeared to diminish such that later generation problem solvers
were less able to exceed the accomplishments of their ancestors, but if
the combinatorial complexity of the solutions was taken into account, the
effect of inheritance was not found to decrease. If the amount of time
problem solvers have to solve new problems not solved by their ancestors
is taken into account, future generations of problem solvers progress
at the same rate regardless of the number of items inherited.

In the face of conflicting results, we propose better experiments, some
of which are already underway. We argue that the only way to truly understand
whether inheritance impacts future problem solving is to compare diachronic
problem solving based around inheritance to alternative strategies for
solving the same set of problems, but without diachronic inheritance.

Although more work on this topic is clearly required, we believe extending
the iterated learning paradigm to answer questions about the impact of
inheritance on problem solving is a valuable contribution.

\setlength{\parindent}{-0.1in}
\setlength{\leftskip}{0.125in}
\noindent
