---
title: Cumulative improvements in iterated problem solving
author:
  - name: Pierce Edmiston
    email: pedmiston@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - name: Maxime Derex
    email: maxime.derex@gmail.com
    department: Biosciences Department
    university: University of Exeter
  - name: Gary Lupyan
    email: lupyan@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
    last: yes
bibliography: references.bib
abstract: |
  As compared to other animals, humans are particularly skilled at using and
  improving tools and other solutions to problems that were first discovered by
  other people. Although the human capacity for cumulative cultural evolution
  is well-known, the effectiveness of inheritance as a form of problem solving
  is an area in need of further research. We report an experiment designed to
  understand how effectively problem solving knowledge accumulates over
  generations of problem solving. We found that problem solvers were
  consistently able to solve more problems in a 25 minute session than their
  ancestors. In investigating whether the number of inherited solutions had an
  impact on future problem solving ability, we found that larger inheritances
  were harder to exceed, but they also allowed individuals to solve more
  complex problems than they would have solved on their own. We discuss
  the limitations of this work, and motivate future directions.
keywords: "cultural evolution; transmission chain; iterated learning"
output:
  pdf_document:
    template: templates/cogsci.tex
    citation_package: natbib
    keep_tex: yes
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  fig.crop = F,
  fig.pos = "tb",
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)
```

```{r setup, include=FALSE, cache=FALSE}
```

```{r methods, include=FALSE}
```

```{r exp1, include=FALSE}
```

# Introduction

Humans are effective problem solvers, having solved a wide range of problems
related to foraging, hunting, and preparing food, while surviving predators,
each other, and a large range of terrestrial environments [@Boyd:2017diff;
@Fernandez:2001civ]. What has enabled our success in being able to solve such
a diverse set of problems? Some have suggested that the answer lies more in our
ability to inherit knowledge from others than our ability to solve problems as
individuals [@Richerson:2008not; @Henrich:2015srt; @Boyd:2017diff]. Cultural
inheritance enables problem solving to extend over generations, far longer than
any single lifetime, allowing future generations of problem solvers to benefit
from solutions to problems that they could not have invented on their own.
However, what is unclear is how effectively individual problem solvers are
able to benefit from inherited solutions at much shorter timescales. To
better understand the extent to which future generations of problem solvers are
able to benefit from inherited problem solving knowledge, we investigated the
accumulation of problem solving knowledge over generations of problem solving.

The accumulation of problem solving knowledge is thought to be enabled by
a number of social learning abilities that facilitate transmission of problem solving knowledge with high fidelity across individuals. These include
teaching through verbal instruction, prosociality, and advanced imitation
abilities [@Dean:2012gv]. These skills for high fidelity social learning are
argued to benefit problem solvers by providing a reliable shortcut to
individual (asocial) learning [@Boyd:1985cu]. Rather than having individuals
struggle to solve a problem on their own through trial-and-error learning,
social learners can inherit solutions that have already been discovered by
others. Over generations, these shortcuts accumulate, giving populations
of social learners to get a head start on future problem solving, and allowing
them to solve problems that no individual learner would be able to solve on
their own and in a single lifetime.

However, the ability to learn socially from others is not sufficient to explain
cumulative cultural evolution. Although social learning was once thought to be
rare in the animal kingdom [e.g., @Thorndike:1898], it has now been documented
in a range of species from chimps [@Whiten:1999cic] to fish [@Laland:1997ts]
and even bees [@Alem:2016fu]. If cumulative cultural evolution depended simply
on social learning, we might expect these species to likewise show evidence of
cumulative cultural evolution, yet such evidence is notably lacking
[@Dean:2012gv; @Tennie:2009gaa]. There are a few exceptions: New Caledonian
crows have been argued to have improved the grub-skewers they craft from
pandanus leaves [@Hunt:2003dj], and at least one population of chimpanzees has
evolved a more advanced termite probe [@Sanz:2009ja], but for the most part
tool use in animals has remained largely unchanged [cf. @Mercader:2007fd].

Humans, in contrast, have demonstrated a remarkable ability to reproduce,
adapt, and improve the innovations of others, so much so that the history of
human technology has been argued to be better understood as a process of
continual refinement and repurposing rather than periods of relative stasis
punctuated by the discoveries of rare geniuses [@Basalla:1989wq; @Sole:2013cm].
Rapid improvement of inherited innovations has not always been the case over
the course of human history, as demonstrated by long periods in the
archaeological record of relative stagnation in stone tool design, raising the
question of what it was that allowed our ancestors to begin to develop more
complex tools. One proposed answer is that humans evolved more
high-fidelity ways of transmitting problem solving knowledge to the next
generation, allowing future generations of problem solvers to more quickly
learn the skills honed by their ancestors, and thus giving them more time to
make improvements to those technologies.

Although this ability to accumulate problem solving knowledge over generations
of problem solvers is widely believed to be fundamental to the story of human
history, we do not fully understand how effectively problem solving knowledge
can be transferred and improved by future generations. To further our
understanding of the human propensity to engage in cumulative cultural
inheritance, we designed an experiment to measure the accumulation of problem
solving knowledge along transmission chains. In addition to establishing that
future generations of problem solvers are able to inherit and exceed their
ancestors, we also sought to understand whether inheriting solutions from
a previous generation had any influence on future problem solving ability. Does
inheritance affect the way in which future problems are solved, and if so, does
it increase or decrease problem solving abilities of future generations?

# Methods

To understand how problem solving knowledge accumulates through vertical
transmission, we used a transmission chain paradigm where participants were
assigned to a single generation within a four-generation chain. Each participant
attempted the same problem solving task for 25 minutes. The solutions that each
participant had discovered by the end of the session were passed on to be
inherited by the participant in the next generation of the chain. Thus,
participants assigned to generations after the first began the problem solving
task with the solutions inherited from the previous generation.

Participants played the "Totems" game adapted from @Derex:2015cbb. Their task
was to discover how to build tools with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. 2). More advanced tools produce larger and more
intricate totems, resulting in higher performance scores.

```{r fig2, fig.height=6, fig.cap="A sample of the solution landscape. New items could be produced through the combinations of different items (more than one arrow points to the item) or through the refinement of a single item (a single arrow points to the item).  The axe is required to construct the first totem pole."}
draw_image("landscape-sample.gv", package = "totems")
```

Participants discovered new tools by combining existing items. Participants
could combine up to four items at a time (with replacement), meaning the initial
six items could form a total of `r exp1$n_unique_guesses_6` combinations.
Of all possible combinations, very few resulted in new items. For example,
of all the guesses that could be formed from the initial items, only three
(`r exp1$n_unique_guesses_6_pct`%) yielded new tools (Fig. 2). As participants
accumulate solutions, the combinatorial complexity of the problem space
increases exponentially such that the discovery of more complex tools is less
likely to happen by chance.

Once a tool was discovered, the recipe for its creation was recorded in an
innovation record. Participants could review their past innovations and see the
recipes for their previous discoveries. Participants assigned to generations
after the first inherited the innovation record of the previous generation
participant. Thus, they could review the recipes for all the innovations that had
been discovered by their ancestor. In order use these tools in futher
combinations, the tools and all of their constituent parts first had to be recreated.

## Participants

Participants were recruited from the UW-Madison student body and received
course credit in exchange for participation. Each participant was assigned to
a single generation of a four-person team (Table 1). Data was collected for
a total of `r exp1$n_teams` teams (N=`r exp1$n_participants` participants).

```{r tab1, results="asis"}
Exp1Participants %>%
  kable(caption = "Participants.", format = "latex")
```

# Results

Participants who inherited the tools discovered by the previous generation
took on average `r exp1$mean_learning_time_min` minutes of the 25 minute
session (`r exp1$proportion_learning_time`%) to recreate the inherited tools---
a portion of the experiment we refer to as the learning period. The length of
the learning period correlated positively with the number of inherited tools,
`r report_lm_mod(learning_rates_mod, "InheritanceSize")` (Fig. 3). A small
group of participants (N=`r exp1$n_outliers`) took disproportionately long to
recreate the tools they inherited perhaps because they were confused by the game
interface. These participants are identified by X's in all figures and are not
included in any statistical analyses.

```{r fig3, fig.cap="Learning rates. Correlation between the number of tools inherited and the time it took to recreate the inherited items. Outliers who were unwilling or unable to recreate the inherited items are shown as X's, and excluded from analysis."}
learning_rates_plot
```

After recreating all inherited tools, participants attempted to
discover new tools not discovered by their ancestor. Most participants
(`r exp1$pct_did_exceed`%) were able to exceed the number of tools that they
inherited `r exp1$logodds_of_exceeding` (Fig. 4). Using Page's trend test (a
repeated measures test for monotonicity) we found that iterated problem solving
resulted in cumulative improvements in the number of tools that could be
discovered in 25 minutes `r exp1$page_test` (Fig. 5).

```{r fig4, fig.cap="Number of innovations created relative to those inherited. The dotted line is a reference with slope=1 such that points above the line indicate future generations exceeding their ancestors."}
innovations_created_and_inherited_plot
```

```{r fig5, fig.cap="Number of innovations discovered by each generation. Each of the thin green lines is a team. The thick blue line is the model predictions with ±1 standard error."}
innovations_by_generation_plot
```

We next tested whether the benefit to inheritance changed over
generations. We fit a hierarchical regression model to the number of
tools discovered in each generation with nested effects (intercepts and
slopes) for chains. On average, second generation participants were able to
discover `r exp1$gen0_slope` more tools than first generation participants,
`r exp1$gen0_slope_stats`. This effect decreased by `r exp1$gen0sqr_slope`
each generation for third and fourth generation participants,
`r exp1$gen0sqr_slope_stats`---a diminishing return. In line with
this result, we also found that an increase in the
number of inherited tools was associated with a decrease in the
number of new tools discovered by future generations of problem solvers,
`r exp1$inheritance_size_slope_stats` (Fig. 4). Together, these findings
suggest that inheritance has a measurable influence on problem solving beyond
providing a shortcut to individual learning. In this case, the impact
appears to be negative, such that inheriting more solutions appears to
_decrease_ problem solving ability.

However, an alternative explanation, not having to do with inheritance
negatively affecting problem solving ability, is that as more tools are
accumulated, additional tools are less likely to be discovered by chance,
meaning we would expect problem solving slow down regardless of what was
inherited. To control for combinatorial complexity, we created an alternative
outcome measure that, rather than counting all innovations equally, instead
weighted each innovation relative to the number of guesses that could be
made out of the items present at the time at which the innovation was discovered.
Fig. 6 shows the change in accumulated difficulty score based on the number of
innovations that were inherited. In contrast to the analysis that treated all
innovations equally, when innovations are weighted by combinatorial complexity,
the effect of inheritance does not decrease over time, such that larger
inheritances result in higher accumulated difficulty scores,
`r exp1$delta_diff_slope_stats`.

```{r fig6, fig.cap="Change in accumulated difficulty score by inheritance. Accumulated difficulty scores are sums of discovered innovations weighted by the combinatorial complexity of the possible choices available to the participant when the innovation was discovered. The outcome measure is the difference between the accumulated difficulty scores of subsequent generations. The line shows the predictions of a hierarchical regression model with ±1 standard error."}
delta_difficulty_plot
```

The final explanation we consider regarding whether inheritance affected problem
solving performance is that larger inheritances take more time to recreate, thus
leaving less time for future problem solving. Even if inheritance did not affect
problem solving performance, participants who inherit more innovations may not
discover as many future innovations simply because they used up most of their
time recreating the items that were inherited. To investigate, we looked at
problem solving rates controlling for the amount of time spent recreating the
inherited items.

Problem solvers who spend more time recreating their inherited items
conversely have less time to discover new tools, which we refer to as
the discovery period. Fig. 7 shows problem solving rates based on the
length of the discovery period. The overall rate of problem solving in the
discovery period was `r exp1$minutes_per_new_innovation` minutes per new
innovation (`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based on the
size of the inheritance, as revealed by a model comparison comparing a model
predicting unique innovations from playing time alone to one predicting unique
innovations from the interaction between playing time and inheritance size,
`r exp1$playing_time_modcomp`. This result suggests that inheriting more tools
did not have an effect on the rate at which new problems are solved.

```{r fig7, fig.env="figure", fig.cap="Problem solving rate in the discovery stage. Discovery time is the amount of time out of a 25 minute session dedicated to discovering new innovations that were not discovered by an ancestor. The line shows the predictions of the hierarchical regression model with ±1 standard error. The slope of this line did not significantly vary based on the size of the inheritance."}
playing_time_plot
```

# Discussion

We measured the accumulation of problem solving knowledge over four-generations
of iterated problem solving. Later generations were consistently able to
cumulatively improve upon the solutions they inherited. Although this finding is
perhaps not surprising, given that participants in effect started a problem
solving task with some of the problems already solved, it is exactly this
ability, so easy for humans to engage in, that is believed to be rare or
non-existent among non-human animals, and fundamental to human cultural
evolution.

In our experiment, problem solvers past the first generation first had to
recreate all of the inherited solutions, which took time and effort. Once
recreated, some participants were unable to solve any new problems. Others were
unable even to recreate all of their inherited solutions in the 25-minute
session-although this behavior is likely attributable to misunderstanding of the
task. Nonetheless, the presence of these outliers highlights that in our
experiment, performance is not guaranteed to improve cumulatively, which is why
we first establish that performance does reliably accumulate.

We also attempted to determine whether inheritance had a measurable impact on
future problem solving ability. We primarily tested this question in this
experiment by investigating the relationship between inheritance size and the
number of unique innovations contributed by each generation. We found multiple,
conflicting results. The impact of inheritance on problem solving appeared to
diminish such that later generation problem solvers were less able to exceed the
number of solutions discovered by their ancestors, but if the combinatorial
complexity of the solutions was taken into account, the effect of inheritance
was not found to decrease. If the amount of time problem solvers have to solve
new problems not solved by their ancestors is taken into account, future
generations of problem solvers progress at the same rate regardless of the
number of items inherited.

In the face of conflicting results, we propose better experiments, some of which
are already underway. We argue that the only way to truly understand whether
inheritance impacts future problem solving is to compare diachronic problem
solving based around inheritance to alternative strategies for solving the same
set of problems, but without diachronic inheritance. Although more work on this
topic is clearly required, we believe extending the iterated learning paradigm
to answer questions about the impact of inheritance on problem solving is a
valuable contribution.

\setlength{\parindent}{-0.1in}
\setlength{\leftskip}{0.125in}
\noindent
