---
title: "Cumulative improvements in iterated problem solving"
author: "Pierce Edmiston"
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  dpi = 96,
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)

# Read all the diagrams made with graphviz in the totems package
chunks <- c(
  "team-structures-exp1",
)
sapply(chunks, read_graphviz_chunk, package = "totems")
```

```{r setup, include=FALSE, cache=FALSE}
```

# Introduction

# Methods

(ref:team-structures-exp1) Diachronic problem solving examined in
Experiment 1. Participants were assigned to four-person diachronic teams.
Each participant completed the same problem solving task for 25 minutes.
Participants in generations 2-4 began the problem solving task with the
solutions that had been discovered by the previous generation.

```{r team-structures-exp1, engine="dot", fig.cap="(ref:team-structures-exp1)"}
```

## Materials

## Participants

## Procedure

# Results

```{r exp1, include=FALSE}
```

# Discussion

