---
title: Cumulative improvements in iterated problem solving
author:
  - name: Pierce Edmiston
    email: pedmiston@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - name: Maxime Derex
    email: maxime.derex@gmail.com
    department: Biosciences Department
    university: University of Exeter
  - name: Gary Lupyan
    email: lupyan@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
    last: yes
bibliography: references.bib
abstract: |
  As compared to other animals, humans are particularly skilled at using and
  improving tools and other solutions to problems that were first discovered by
  other people. Although the human capacity for cumulative cultural evolution
  is well-known, the effectiveness of inheritance as a form of problem solving
  is an area in need of further research. We report an experiment designed to
  understand how effectively solutions to problems accumulate over generations
  of problem solving. Using a tool-discovery game, we found that participants
  were consistently able to discover more tools in a 25 minute session than
  their ancestors. Participants who inherited more tools tended to discover
  fewer novel ones, but the ones they did discover were more complex. We also
  found that participants who inherited tools tended to generate less redundant
  guesses, suggesting a potential downstream consequence to iterated problem
  solving. We discuss the limitations of this work, and motivate future
  directions.
keywords: "cultural evolution; transmission chain; iterated learning"
output:
  bookdown::pdf_document2:
    template: templates/cogsci.tmpl
    citation_package: natbib
    keep_tex: yes
---
```{r iterated-problem-solving-config, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  fig.crop = F,
  fig.pos = "tb"
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "iterated-problem-solving.R"))
```

```{r iterated-problem-solving, include=FALSE}
```

# Introduction {-}

Humans are effective problem solvers, having solved a wide range of problems
related to foraging, hunting, and preparing food, while surviving predators,
each other, and a large range of terrestrial environments [@Boyd:2018diff;
@Fernandez:2001civ]. What has enabled our success in being able to solve such a
diverse set of problems? Some have suggested that the answer lies more in our
ability to inherit knowledge from others than our ability to make discoveries by
ourselves [@Richerson:2005not; @Henrich:2015srt; @Boyd:2018diff]. Humans possess
a number of advanced social learning abilities including teaching through verbal
instruction and imitation that provide reliable ways of transferring problem
solving knowledge across individuals [@Dean:2012gv]. If problem solving
knowledge can be acquired via social learning, then future generations can adapt
and improve it, allowing cultures to accumulate technological complexity over
generations.

However, the ability to learn socially from others is not sufficient to explain
cumulative cultural evolution. Although social learning was once thought to be
rare in the animal kingdom [e.g., @Thorndike:1898], it has now been documented
in a range of species from chimps [@Whiten:1999cic] to fish [@Laland:1997ts]
and even bees [@Alem:2016fu]. If cumulative cultural evolution depended simply
on social learning, we might expect these species to likewise show evidence of
cumulative cultural evolution, yet such evidence is notably lacking
[@Dean:2012gv; @Tennie:2009gaa]. There are a few exceptions: New Caledonian
crows have been argued to have improved the grub-skewers they craft from
pandanus leaves [@Hunt:2003dj], and at least one population of chimpanzees has
evolved a more advanced termite probe [@Sanz:2009ja], but for the most part
tool use in animals has remained largely unchanged [cf. @Mercader:2007fd].

Humans, in contrast, have demonstrated a remarkable ability to adapt
and improve the tools and other innovations discovered by others. The
history of human technology is argued to be better understood as a process
of gradual refinement and repurposing rather than punctuated advances brought by
the discoveries of rare geniuses [@Basalla:1989wq; @Sole:2013cm]. Rapid
refinement of inherited innovations has not always been the case over the course
of human history, as demonstrated by the long periods in the archaeological
record of slow or stagnant growth in stone tool complexity [@Torre:2011jl;
@Lycett:2008di]. As to what it was that allowed our ancestors to begin to
develop more complex tools, an answer that has been proposed is that humans
evolved more robust ways of transmitting cultural information to the next
generation. This allowed future generations to more quickly learn
the skills honed by their ancestors, and thus giving them more time to make
improvements to those technologies [@Sterelny:2012ea]. On this view,
human cultural evolution has been defined not by our ability to inherit the
skills of our ancestors, but by the ability to exceed and improve them.

```{r iterated-problem-solving-fig, fig.pos="b", fig.align="center", fig.width=3.5, fig.cap="Iterated problem solving paradigm. Participants were assigned to generations within chains. Each participant completed the same problem solving task for 25 minutes. Participants in generations after the first began the problem solving task with the solutions that were discovered by the previous generation."}
draw_graphviz("iterated-problem-solving", package = "totems")
```

We investigated the human propensity to exceed our ancestors in a problem
solving task using a transmission chain paradigm (Fig.
\@ref(fig:iterated-problem-solving-fig)). Previous research using this paradigm
has found that problem solving performance for simple tasks like folding paper
airplanes and building towers out of spaghetti can accumulate over generations
[@Caldwell:2008gf], although the mechanisms are unclear [@Caldwell:2009fl].
@Morgan:2015hu more carefully manipulated the mechanisms of transmission, and
found that teaching, and especially verbal instruction, was crucial for
successful transmission of skilled behaviors along chains. Here we controlled
for successful transmission by allowing future generations to inherit symbolic
information about how to recreate the tools that had been discovered by an
ancestor. Specifically, participants inherited the recipes for creating tools
from the combination of items. Once inherited, the tools had to be created
before they could be used in further combinations. We then measured the ability
of participants to recreate and exceed the tools they inherited.

In addition to asking whether participants were able to exceed the total number
of tools discovered by their ancestor, we also asked whether inheritance
influenced the way in which future problems were solved. To answer this
question, we analyzed whether participants who inherited more tools from their
ancestors were more or less effective at discovering new tools. We also
analyzed the guessing strategies used by participants who benefited from
inheritance as compared to first generation participants who did not inherit
from any ancestor. These analyses are used to address potential downstream
consequences to iteratively inheriting from a previous generation.

# Methods {-}

To understand how solutions to problems accumulate through vertical
transmission, we used a transmission chain paradigm where participants were
assigned to a single generation within a four-generation chain. Each
participant attempted the same tool discovery task for 25 minutes. The recipes
for how to create the tools that each participant had discovered by the end of
the session were passed on to be inherited by a participant in the next
generation of the chain. Thus, participants assigned to generations after the
first began the experiment with information about how to create the tools
inherited from the previous generation.

Participants played the "Totems" game adapted from @Derex:2015cbb. Their task
was to discover how to build tools with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample-1)). More advanced tools
produce larger and more intricate totems, resulting in higher performance
scores.

```{r landscape-sample-1, fig.height=6, fig.cap="A sample of the solution landscape. The top row of 6 items were available to participants at the start of the game. New items could be produced through the combination of different items (more than one arrow points to the item) or through the refinement of a single item (a single arrow points to the item). The axe is required to construct the first totem pole."}
draw_image("landscape-sample.gv", package = "totems")
```

Participants discovered new tools by combining existing items. Participants
could refine individual items, or combine up to four items at a time (with
replacement), meaning the initial six items could form a total of
`r methods$n_unique_guesses_6` combinations. Of all possible combinations, very
few resulted in new items. For example, of all the guesses that could be formed
from the initial items, only three (`r methods$n_unique_guesses_6_pct`%)
yielded new tools.

As tools are accumulated, the number of possible combinations that can be made
with those tools increases exponentially such that the discovery of later tools
was less likely to happen by chance alone. This increase in combinatorial
complexity suggests that later tools might be harder to discover than earlier
tools. However, combinatorial complexity is not a perfect measure of difficulty
in this task because of the way the solution space was designed, and because
participants do not guess entirely at random. Based on previous research using
this task, we know that participants are far more likely to make some guesses
than others, indicating they are using tacit knowledge about the tools to
generate combinations. For example, once discovering an axe, participants
quickly discover that they can use the axe to chop down a tree, regardless of
the other tools they may have at the time. At the same time, participants do
not find all tools equally intuitive, and the difference in combinatorial
complexity should not be ignored. In our results, we report performance based
on both measures.

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record. Participants could review their past innovations and see the
recipes for their previous discoveries. **Participants assigned to generations
after the first inherited the innovation record of the previous generation
participant.** From the beginning of the experiment, these participants could
review the recipes for all the innovations that had been discovered by their
ancestor. Note that the participants inherited the recipes, but not the tools
themselves. In order use these tools in further combinations, the tools and all
of their constituent parts first had to be recreated.

## Participants {-}

Participants were recruited from the UW-Madison student body and received
course credit in exchange for participation. Each participant was assigned to
a single generation of a four-generation chain. Data was collected
for a total of `r exp1$n_teams` complete chains (N=`r exp1$n_participants` participants).

# Results {-}

Our results are presented in three sections. First, we report performance over
generations along each chain to determine the extent to which each new
generation was able to exceed the innovations discovered by their ancestor. The
remaining analyses were carried out to determine whether participants' ability
to discover new tools was influenced by the tools they inherited. In the second
section, we report performance based on the number of inherited tools as
opposed to generation. In the final section, we compare the guessing strategies
of participants who inherited from an ancestor compared to first generation
participants who did not benefit from inheritance.

## Performance by generation {-}

We found that tools accumulated over generations such that participants in
later generations were able to discover more tools in the same amount of time
than their ancestors (Fig. \@ref(fig:innovations-by-generation)). We tested
this in multiple ways. First, we used Page's trend test (a repeated measures
test for monotonicity), and found evidence that the number of tools discovered
within each chain increased over generations, `r exp1$page_test`. Next, we
tested whether tools accumulated linearly or quadratically over generations. We
fit a hierarchical regression model with polynomial contrasts to the number of
tools discovered in each generation with random effects for chain. On average,
second generation participants were able to discover `r exp1$gen0_slope` more
tools than first generation participants, `r exp1$gen0_slope_stats`. This
effect decreased by `r exp1$gen0sqr_slope` each generation for third and fourth
generation participants, `r exp1$gen0sqr_slope_stats`. That the numbers
of new tools per generation levels off for later generations could suggest
that later generations have a harder time exceeding their ancestors, but it
also could be due to these later generations, having inherited more tools,
requiring more time in the experimental session to recreate the inherited
items. We return to both of these potential explanations below.

```{r innovations-by-generation, fig.cap="Performance by generation. Each of the thin green lines is a chain. The thick blue line is the model predictions with ±1 standard error."}
innovations_by_generation_plot
```

Although later generation participants made fewer new discoveries, the
discoveries they made were often more complex. We computed the complexity of a
tool as the number of combinations that could be made given the number of items
in the participants' inventories at the time the tool was discovered. A tool
that was discovered among 12 items was deemed more complex than a tool that was
discovered among 6 items. Using a hierarchical regression model as
reported above, we found that when discoveries were weighted by combinatorial
complexity, participants discovered more complex tools with each generation,
`r exp1$difficulty_by_generation_mod_slope_stats`, and this benefit did not
diminish for later generations, `r exp1$difficulty_by_generation_quad_mod_slope_stats`.

## Performance by inheritance size {-}

```{r inheritance-size, fig.height=8, fig.cap="Performance by inheritance size. A. Number of tools created relative to those inherited. The dotted line is a reference with slope=1 such that points above the line indicate future generations exceeding their ancestors. B. Number of new tools relative to those inherited. The same reference line is now shown horizontally. C. Change in accumulated difficulty score. Analogous to B, but with innovations weighted by complexity. In B and C, the error ranges specify the model predictions with ±1 standard error."}
grid.arrange(
  innovations_created_and_inherited_plot + ggtitle("A"),
  new_innovations_plot + ggtitle("B"),
  delta_difficulty_plot + ggtitle("C"),
  ncol = 1
)
```

Because there is no difference between a second generation participant who
inherits 10 tools and a fourth generation participant who inherits the same
tools, we also looked at problem solving performance relative to the number
of tools that were inherited regardless of generation (Fig.
\@ref(fig:inheritance-size)). As the number of inherited tools increased, the
number of new tools discovered decreased, `r exp1$inheritance_size_slope_stats`
(Fig. \@ref(fig:inheritance-size)B). That participants who inherited more items
had a harder time exceeding them is expected based on the analysis of
performance over generations reported above. However, when tools are weighted
by combinatorial complexity, the amount by which participants are able to
exceed their ancestors actually increases for larger inheritance sizes,
`r exp1$delta_diff_slope_stats` (Fig. \@ref(fig:inheritance-size)C), though we are
hesitant to draw strong conclusions from this finding as it may be explainable
by the characteristics of the solution landscape, a point we return to in the
Discussion.

Participants who inherited more tools also required more time to recreate those
tools. We next asked whether inheriting more tools had an impact on the rate of new
tool discovery, controlling for the length of time spent recreating inherited tools.
Participants took on average `r exp1$mean_learning_time_min` minutes of the 25
minute session (`r exp1$proportion_learning_time`%) to recreate the inherited
tools---a portion of the experiment we refer to as the learning period. The
length of the learning period correlated positively with the number of inherited
tools (Fig. \@ref(fig:learning-rates)).

```{r learning-rates, fig.cap="Learning rates. Correlation between the number of tools inherited and the time it took to recreate the inherited items. Outliers who were appear unwilling or unable to recreate the inherited items are shown as X's, but included in all analyses."}
learning_rates_plot
```

We then calculated discovery rates for each participant: the number of
new tools discovered relative to the amount of time out of the 25 minute
session available to discover new tools (Fig. \@ref(fig:playing-time)). The overall rate of
problem solving was `r exp1$minutes_per_new_innovation` minutes per new
innovation (`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based on
the number of inherited tools, as revealed by comparing a model predicting
novel tools from discovery time alone to one predicting novel tools
from the interaction between discovery time and inheritance size,
`r exp1$playing_time_modcomp`. This suggests that inheriting more tools did not
meaningfully affect the rate of future problem solving.

```{r playing-time, fig.cap="New tool discovery rates. Discovery time is the amount of time out of a 25 minute session dedicated to discovering new innovations that were not discovered by an ancestor. The line shows the predictions of the hierarchical regression model with ±1 standard error. The slope of this line did not significantly vary based on the number of inherited tools."}
playing_time_plot
```

## Performance by inheritance {-}

In this section, we report whether participants who inherited tools from an
ancestor guessed differently than participants in the first generation who had
to discover the first tools on their own. We found that participants who
benefited from inheritance on average made fewer guesses per tool than when the
same tools were attempted in the first generation,
`r exp1$guesses_per_item_by_inheritance`. As expected, this effect is especially
pronounced for the tools that were inherited. We did not find any evidence that
inheritance had an affect on the number of guesses per new tool,
`r exp1$guesses_per_new_item_by_inheritance_mod` (Fig.
\@ref(fig:guesses-per-innovation-1)). This finding suggests that although
inheritance benefits participants in recreating inherited solutions, it does
not confer any benefits to future problem solving.

```{r guesses-per-innovation-1, fig.height=6, fig.cap="Guesses per tool by inheritance condition. Each line is the average number of guesses it took to discover a particular tool in either inheritance condition. Error bars show ±1 standard of the model predictions. A. Overall guesses per tool, including those inherited. B. Guesses per novel tool."}
grid.arrange(
  guesses_per_item_by_inheritance_plot + ggtitle("A"),
  guesses_per_new_item_by_inheritance_plot + ggtitle("B"),
  ncol = 1
)
```

Finally, we analyzed the guessing strategy employed by participants who
inherited from an ancestor versus those who did not. Guesses were categorized
into one of four types: redundant guesses, repeat items, unique guesses, and
unique items. Redundant guesses were incorrect guesses made once before. Repeat
items were recreated tools. Unique guesses were incorrect guesses never made before.
Unique items were unique guesses that created a new tool.

The proportion of guesses made by participants based on their inheritance condition
is shown in Fig. \@ref(fig:guess-types-1). We found that participants who
inherited solutions at the start tended to make fewer redundant guesses than
first generation participants, `r exp1$num_redundant_guesses`, a difference of
12%, `r exp1$prop_redundant_guesses`. This result suggests that starting off
the problem solving task with solutions inherited from a previous generation
may have influenced the way in which future guesses are generated.

```{r guess-types-1, fig.width=3.8, fig.cap="Guessing strategies by inheritance condition."}
prop_guess_types_50min_plot
```

# Discussion {-}

We found that participants were consistently able to solve more problems in
a single 25 minute session than their ancestors, and thus were able to
cumulatively improve upon the solutions they inherited. All participants were
expected to be able to recreate the tools they inherited, but whether they
could discover new tools, beyond those inherited, was unknown. Given the
combinatorial complexity of the solution landscape, participants were unlikely
to strike upon beneficial combinations by guessing at random. Because of this,
some participants were unable to discover any new tools, but most did discover
new tools, even when inheriting an already large number of previously
discovered tools.

We also explored the impact of inherited solutions on future problem solving
performance. Here our results were mixed. We found that participants who
inherited more tools tended to discover fewer new tools than their ancestors,
suggesting that the benefit to inheritance leveled off for later generations.
However, if the combinatorial complexity of the solutions is taken into
account, we found that participants were able to accumulate more complex tools
than they would have been likely to discover on their own (i.e., as a first
generation participant). In addition, we found that controlling for the amount
of time each participant had to discover new tools (as opposed to recreating
inherited tools) failed to reveal an effect of inheritance size on the rate of
new tool discovery. Finally, we investigated whether starting off a problem
solving task with inherited solutions influenced the way in which tools were
discovered, and found that inheritance did not have an influence on the number
of guesses required for new items, but that inheritance might benefit future
generations by promoting a less redundant guessing strategy.

Our conclusions are limited by the design of the solution landscape in the
Totems game, and the restriction in our methods to a single problem solving
strategy. The sparsity of the solution landscape, where many combinations can
be made but very few yield new tools, indicates that in order to succeed
participants must use tacit knowledge to help form combinations that are most
likely to yield new tools. This challenges the notion that the difficulty of a
particular tool is directly related to its combinatorial complexity. In
addition, we believe the accumulation of problem solving knowledge through
vertical transmission must be compared with the accumulation of problem solving
knowledge through other forms of problem solving that do not involve vertical
transmission.

More than any other animal, humans are particularly skilled at inheriting and
improving tools and other solutions to problems, but whether the ability
to inherit from others has effects on problem solving beyond giving a head
start to individual learning is not known. Although much work is still
needed to fully understand the human propensity for cumulative cultural
evolution, we believe our research is a valuable contribution to ongoing
efforts to understand how and why human culture is so integrally cumulative.
