---
title: Cumulative improvements in iterated problem solving
author:
  - name: Pierce Edmiston
    email: pedmiston@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - name: Maxime Derex
    email: maxime.derex@gmail.com
    department: Biosciences Department
    university: University of Exeter
  - name: Gary Lupyan
    email: lupyan@wisc.edu
    department: Department of Psychology
    university: University of Wisconsin-Madison
    last: yes
bibliography: references.bib
abstract: |
  As compared to other animals, humans are particularly skilled at using and
  improving tools and other solutions to problems that were first discovered by
  other people. Although the human capacity for cumulative cultural evolution is
  well-known, the effectiveness of inheritance as a form of problem solving is
  an area in need of further research. We report an experiment designed to
  understand how effectively problem solving knowledge accumulates over
  generations of problem solving. We found that participants were consistently
  able to solve more problems in a 25 minute session than their ancestors.
  Participants who inherited more solutions tended to solve fewer novel ones,
  but the ones they did solve were more complex. We also found that participants
  who inherited solutions tended to generate less redundant guesses, suggesting
  a potential downstream consequence to iterated problem solving. We discuss the
  limitations of this work, and motivate future directions.
keywords: "cultural evolution; transmission chain; iterated learning"
output:
  bookdown::pdf_document2:
    template: templates/cogsci.tmpl
    citation_package: natbib
    keep_tex: yes
---
```{r iterated-problem-solving-config, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  fig.crop = F,
  fig.pos = "tb",
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "iterated-problem-solving.R"))
```

```{r iterated-problem-solving, include=FALSE}
```

# Introduction

Humans are effective problem solvers, having solved a wide range of problems
related to foraging, hunting, and preparing food, while surviving predators,
each other, and a large range of terrestrial environments [@Boyd:2017diff;
@Fernandez:2001civ]. What has enabled our success in being able to solve such a
diverse set of problems? Some have suggested that the answer lies more in our
ability to inherit knowledge from others than our ability to make discoveries by
ourselves [@Richerson:2008not; @Henrich:2015srt; @Boyd:2017diff]. Humans possess
a number of advanced social learning abilities including teaching through verbal
instruction and imitation that provide reliable ways of transferring problem
solving knowledge across individuals [@Dean:2012gv]. If problem solving
knowledge can be acquired via social learning, then future generations can adapt
and improve it, allowing cultures to accumulate technological complexity over
generations.

However, the ability to learn socially from others is not sufficient to explain
cumulative cultural evolution. Although social learning was once thought to be
rare in the animal kingdom [e.g., @Thorndike:1898], it has now been documented
in a range of species from chimps [@Whiten:1999cic] to fish [@Laland:1997ts]
and even bees [@Alem:2016fu]. If cumulative cultural evolution depended simply
on social learning, we might expect these species to likewise show evidence of
cumulative cultural evolution, yet such evidence is notably lacking
[@Dean:2012gv; @Tennie:2009gaa]. There are a few exceptions: New Caledonian
crows have been argued to have improved the grub-skewers they craft from
pandanus leaves [@Hunt:2003dj], and at least one population of chimpanzees has
evolved a more advanced termite probe [@Sanz:2009ja], but for the most part
tool use in animals has remained largely unchanged [cf. @Mercader:2007fd].

Humans, in contrast, have demonstrated a remarkable ability to adapt
and improve the tools and other innovations discovered by others. The
history of human technology is argued to be better understood as a process
of gradual refinement and repurposing rather than punctuated advances brought by
the discoveries of rare geniuses [@Basalla:1989wq; @Sole:2013cm]. Rapid
refinement of inherited innovations has not always been the case over the course
of human history, as demonstrated by the long periods in the archaeological
record of slow or stagnant growth in stone tool complexity [@Torre:2011jl;
@Lycett:2008di]. As to what it was that allowed our ancestors to begin to
develop more complex tools, an answer that has been proposed is that humans
evolved more robust ways of transmitting cultural information to the next
generation, allowing future generations of problem solvers to more quickly learn
the skills honed by their ancestors, and thus giving them more time to make
improvements to those technologies [@Sterelny:2012ea]. On this view,
human cultural evolution has been defined not by our ability to inherit the
skills of our ancestors, but by the ability to exceed and improve them.

```{r fig1, fig.pos="b", fig.align="center", fig.width=3.5, fig.cap="Transmission of problem solving knowledge. Participants were assigned to generations within chains. Each participant completed the same problem solving task for 25 minutes. Participants in generations after the first began the problem solving task with the solutions that were discovered by the previous generation."}
draw_graphviz("iterated-problem-solving", package = "totems")
```

We investigated the human propensity to exceed our ancestors in a problem
solving task using a transmission chain paradigm (Fig. 1). Previous research
using this paradigm has found that problem solving performance for simple tasks
like folding paper airplanes and building tours out of spaghetti can accumulate
over generations [@Caldwell:2008gf], although the mechanisms are unclear
[@Caldwell:2009fl]. @Morgan:2015hu more carefully manipulated the mechanisms of
transmission, and found that teaching, and especially verbal instruction, was
crucial for successful transmission of skilled behaviors along chains. Here we
controlled for successful transmission by allowing future generations to
inherit the discoveries made by the previous generation, and measured the
ability of participants to exceed the solutions they inherited.

# Methods

To understand how problem solving knowledge accumulates through vertical
transmission, we used a transmission chain paradigm where participants were
assigned to a single generation within a four-generation chain. Each
participant attempted the same problem solving task for 25 minutes. The
solutions that each participant had discovered by the end of the session were
passed on to be inherited by a participant in the next generation of the
chain. Thus, participants assigned to generations after the first began the
problem solving task with solutions inherited from the previous generation.

Participants played the "Totems" game adapted from @Derex:2015cbb. Their task
was to discover how to build tools with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. 2). More advanced tools produce larger and more
intricate totems, resulting in higher performance scores.

```{r fig2, fig.height=6, fig.cap="A sample of the solution landscape. The top row of 6 items were available to problem solvers at the start of the game. New items could be produced through the combination of different items (more than one arrow points to the item) or through the refinement of a single item (a single arrow points to the item). The axe is required to construct the first totem pole."}
draw_image("landscape-sample.gv", package = "totems")
```

Participants discovered new tools by combining existing items. Participants
could combine up to four items at a time (with replacement), meaning the
initial six items could form a total of `r exp1$n_unique_guesses_6`
combinations. Of all possible combinations, very few resulted in new items. For
example, of all the guesses that could be formed from the initial items, only
three (`r methods$n_unique_guesses_6_pct`%) yielded new tools.

In the game, as tools are accumulated, the number of possible combinations that
can be made with those tools increases exponentially such that the discovery of
later tools was less likely to happen by chance alone. This increase in
combinatorial complexity suggests that later tools might be harder to discover
than earlier tools. However, combinatorial complexity is not a perfect measure
of difficulty in this task because participants do not guess entirely at
random. In reality participants are far more likely to make some guesses than
others. For example, once discovering an axe, participants quickly discover
that they can use the axe to chop down a tree, regardless of the other tools
they may have at the time. At the same time, participants do not find all tools
equally intuitive, and the difference in combinatorial complexity should not be
ignored. In our results, we report performance based on both measures.

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record. Participants could review their past innovations and see the
recipes for their previous discoveries. **Participants assigned to generations
after the first inherited the innovation record of the previous generation
participant.** From the beginning of the experiment, these participants could
review the recipes for all the innovations that had been discovered by their
ancestor. Note that the participants inherited the recipes, but not the tools
themselves. In order use these tools in further combinations, the tools and all
of their constituent parts first had to be recreated.

## Participants

Participants were recruited from the UW-Madison student body and received
course credit in exchange for participation. Each participant was assigned to
a single generation of a four-generation chain. Data was collected
for a total of `r exp1$n_teams` complete chains (N=`r exp1$n_participants` participants).

# Results

## Performance by generation

We found that problem solving knowledge accumulated over generations such that
participants in later generations were able to solve more problems in the same
amount of time than their ancestors (Fig. 3). We tested this in multiple ways.
First, we used Page's trend test (a repeated measures test for monotonicity),
and found evidence that the number of tools discovered within each chain
increased over generations `r exp1$page_test`. Next, we tested whether the
extent to which participants were able to exceed their ancestors changed over
generation. We fit a hierarchical regression model to the number of tools
discovered in each generation with these effects nested by chain. On average,
second generation participants were able to discover `r exp1$gen0_slope` more
tools than first generation participants, `r exp1$gen0_slope_stats`. This
effect decreased by `r exp1$gen0sqr_slope` each generation for third and fourth
generation participants, `r exp1$gen0sqr_slope_stats`. This diminishing return
suggests that future generations have a harder time exceeding their ancestors.

```{r fig3, fig.cap="Performance by generation. Each of the thin green lines is a chain. The thick blue line is the model predictions with ±1 standard error."}
innovations_by_generation_plot
```

Although later generation participants made fewer new discoveries, the
discoveries they made were often more complex. We measured complexity as the
combinatorial complexity of the number of items in the participants' inventories
at the time a discovery was made. We found that when discoveries were weighted
by combinatorial complexity, participants solved more complex problems
with each generation, `r exp1$difficulty_by_generation_mod_slope_stats`,
and this benefit did not diminish for later generations,
`r exp1$difficulty_by_generation_quad_mod_slope_stats`.

## Performance by inheritance size

```{r fig4, fig.height=8, fig.cap="Performance by inheritance size. A. Number of tools created relative to those inherited. The dotted line is a reference with slope=1 such that points above the line indicate future generations exceeding their ancestors. B. Number of new tools relative to those inherited. The same reference line is now shown horizontally. C. Change in accumulated difficulty score. Analogous to B, but with innovations weighted by complexity. In B and C, the error ranges specify the model predictions with ±1 standard error."}
grid.arrange(
  innovations_created_and_inherited_plot + ggtitle("A"),
  new_innovations_plot + ggtitle("B"),
  delta_difficulty_plot + ggtitle("C"),
  ncol = 1
)
```

Because there is no difference between a second generation participant
who inherits 10 tools and a fourth generation participant who inherits
the same tools, we also looked at problem solving performance relative
to the number of tools that were inherited (Fig. 4). As the number of
inherited tools increased, the number of new tools discovered decreased,
`r exp1$inheritance_size_slope_stats` (Fig. 4B). That participants who
inherited more items had a harder time exceeding them is expected based on the
analysis of performance over generations reported above. However, when tools
are weighted by combinatorial complexity, the amount by which participants are
able to exceed their ancestors actually increases for larger inheritance sizes,
`r exp1$delta_diff_slope_stats` (Fig. 4C), though we are hesitant to draw strong
conclusions from this finding as it may be explanable by the characteristics of
the solution landscape, a point we return to in the Discussion.

We next asked whether inheriting more tools had an impact on the rate of new
tool discovery. To answer this question, we first needed to identify the length
of time spent recreating inherited tools as opposed to discovering new ones.
Participants took on average `r exp1$mean_learning_time_min` minutes of the 25
minute session (`r exp1$proportion_learning_time`%) to recreate the inherited
tools---a portion of the experiment we refer to as the learning period. The
length of the learning period correlated positively with the number of inherited
tools (Fig. 5).

```{r fig5, fig.cap="Learning rates. Correlation between the number of tools inherited and the time it took to recreate the inherited items. Outliers who were appear unwilling or unable to recreate the inherited items are shown as X's, but included in all analyses."}
learning_rates_plot
```

We then calculated discovery rates for each participant: the number of
novel discoveries relative to the amount of time out of the 25 minute
session spent discovering new tools (Fig. 6). The overall rate of
problem solving was `r exp1$minutes_per_new_innovation` minutes per new
innovation (`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based on
the number of inherited tools, as revealed by comparing a model predicting
unique innovations from playing time alone to one predicting unique
innovations from the interaction between playing time and inheritance size,
`r exp1$playing_time_modcomp`. This suggests that inheriting more tools did not
meaningfully diminish the rate of future problem solving.

```{r fig6, fig.cap="Problem solving rate in the discovery stage. Discovery time is the amount of time out of a 25 minute session dedicated to discovering new innovations that were not discovered by an ancestor. The line shows the predictions of the hierarchical regression model with ±1 standard error. The slope of this line did not significantly vary based on the size of the inheritance."}
playing_time_plot
```

## Performance by inheritance

Next we compared performance in the first generation, where participants did
not benefit from any previous ancestor, to performance in generations 2-4.
We found that participants who benefited from inheritance on average made
fewer guesses per item than when the same items were attempted in the first
generation, `r exp1$guesses_per_item_by_inheritance`. As expected, this effect
is especially pronounced for the items that were inherited. We did not find
any evidence that inheritance had an affect on the number of guesses per novel
item, `r exp1$guesses_per_new_item_by_inheritance_mod` (Fig. 7). This finding
suggests that although inheritance benefits participants in recreating inherited
solutions, it does not confer any benefits to future problem solving.

```{r fig7, fig.height=6, fig.cap="Guesses per tool by inheritance condition. Each line is the average number of guesses it took to discover a particular tool in either inheritance condition. Error bars show ±1 standard of the model predictions. A. Overall guesses per tool, including those inherited. B. Guesses per novel tool."}
grid.arrange(
  guesses_per_item_by_inheritance_plot + ggtitle("A"),
  guesses_per_new_item_by_inheritance_plot + ggtitle("B"),
  ncol = 1
)
```

Finally, we analyzed the overall guessing strategy employed by participants
who inherited from an ancestor versus those who did not. Guesses were
categorized into one of four types: redundant guesses, repeat items,
unique guesses, and unique items. The overall proportion of guesses
made by participants based on their inheritance condition is shown in
Fig. 8. We found that participants who inherited solutions at the start
tended to make fewer redundant guesses than first generation participants,
`r exp1$num_redundant_guesses`, a difference of 12%, `r exp1$prop_redundant_guesses`.
This result suggests that starting off the problem solving task with
solutions inherited from a previous generation influenced the way
in which future guesses are generated.

```{r fig8, fig.width=4, fig.cap="Guessing strategies by inheritance condition. Redundant guesses were incorrect guesses made once before. Repeat items were recreated tools. Unique guesses were incorrect guesses never made. Unique items were unique guesses that created a new tool."}
prop_guess_types_50min_plot
```

# Discussion

The ability to accumulate a body of socially learned problem solving knowledge
is visible in all human cultures [@Boyd:2017diff; @Fernandez:2001civ]. Although
the capacity for cumulative improvement is believed to be integral to human
cultural evolution, we do not fully understand how effectively problem solving
knowledge can be utilized by future generations of problem solvers. How
effective are human problem solvers at ratcheting up the tools and other
solutions to problems that were discovered by others? Here
we investigated the likelihood that problem solvers could exceed their ancestors
over generations of problem solving.

In contrast to previous research that has investigated how social learning
strategies such as verbal instruction play a role in high fidelity transmission
of problem solving knowledge, here we examined the likelihood that problem
solving knowledge would accumulate given that solutions could be
effectively transferred to the next generation. Participants in our study
inherited the solutions that had been discovered by their ancestor, and we
measured their ability to recreate and extend this body of problem solving
knowledge. We asked whether an individuals' ability to improve upon inherited
knowledge change across generations and as the amount of inherited information
increases.

We found that problem solvers were consistently able to solve more problems in a
single 25 minute session than their ancestors, and thus were able to
cumulatively improve upon the solutions they inherited. All participants were
expected to be able to recreate the tools they inherited, but whether they could
discover new tools, beyond those inherited, was unknown. Given the combinatorial
complexity of the solution landscape, participants were highly unlikely to
strike upon a beneficial combination by chance alone. Because of this, some
participants were unable to discover any new tools, but most did discover new
tools, even when inheriting an already large number of previously discovered
tools.

We also explored the impact of inherited solutions on future problem solving
performance. Here our results were mixed. We found that problem solvers who
inherited more tools tended to discover fewer new tools than their ancestors,
suggesting that inheritance has a diminishing return on future problem solving
performance. However, if the combinatorial complexity of the solutions is taken
into account, we found that these problem solvers were able to accumulate more
complex tools than they would have been likely to discover on their own. In
addition, we found that controlling for the amount of time each participant had
to discover new tools (as opposed to recreating inherited tools) failed to
reveal an effect of inheritance size on the rate of problem solving. Finally,
we investigated whether starting off a problem solving task influenced the
way in which problems were solved, and found that inheritance did not have
an influence on the number of guesses required for new items, but that
inheritance might benefit future generations by promoting a better problem
solving strategy.

Our conclusions are limited by the nature of the solution landscape in the
Totems game, and the restriction in our methods to a single problem solving
strategy. The sparsity of the solution landscape, where many thousands of
possible combinations can be made but far less than 1% yield new tools,
indicates that participants must use tacit knowledge to help form combinations
that are most likely to yield new tools. For example, once an axe was
discovered, its combination with the tree to create a totem was very likely.
This challenges the notion that the difficulty of a particular tool is directly
related to its combinatorial complexity. In addition, we believe the
accumulation of problem solving knowledge through vertical transmission must be
compared with the accumulation of problem solving knowledge through other forms
of problem solving that do not involve vertical transmission.

More than any other animal, humans are particularly skilled at inheriting and
improving tools and other solutions to problems, but whether the ability
to inherit from others has effects on problem solving beyond giving a head
start to individual learning is not known. Although much work is still
needed to fully understand the human propensity for cumulative cultural
evolution, we believe our research is a valuable contribution to ongoing
efforts to understand how and why human culture is so integrally cumulative.

\setlength{\parindent}{-0.1in}
\setlength{\leftskip}{0.125in}
\noindent
