---
title: "The impact of inheritance on problem solving"
author:
  - name: Pierce Edmiston
    affiliation_id: 1
  - name: Maxime Derex
    affiliation_id: 2
  - name: Gary Lupyan
    affiliation_id: 1
affiliation:
  - id: 1
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - id: 2
    department: Biosciences Department
    university: University of Exeter
email: pedmiston\@wisc.edu
address: 1202 W. Johnson St., Department of Psychology, University of Wisconsin, Madison, WI 53703.
abstract: |

bibliography: "references.bib"
csl: "templates/plos-one.csl"
output:
  bookdown::pdf_document2:
    template: templates/plos-one.tmpl
    keep_tex: yes
    number_sections: no
---
```{r plos-one, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  
  # Width: 789 – 2250 pixels (at 300 dpi). Height maximum: 2625 pixels (at 300 dpi).
  fig.width = 3,
  fig.height = 3,
  dpi = 300,

  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "plos-one.R"))

# Read all the diagrams made with graphviz in the totems package
chunks <- c(
  "team-structures",
  "team-structures-all",
  "team-structures-exp2",
  "team-structures-exp3",
  "team-structures-exp4"
)
sapply(chunks, read_graphviz_chunk, package = "totems")
```

```{r setup, include=FALSE, cache=FALSE}
```

# Introduction {-}

```{r intro, include=FALSE}
```

(ref:types-of-time) Types of time.

```{r types-of-time, fig.width=5, fig.height=2, fig.cap="(ref:types-of-time)"}
grid.arrange(
  read_graphviz("team-structures", package = "totems"),
  gg_time,
  gg_person,
  nrow = 1,
  widths = c(0.25, 0.375, 0.375)
)
```

# Methods {-}

```{r methods, include=FALSE}
```

(ref:team-structures-all) All experimental conditions.

```{r team-structures-all, out.width="50%", engine="dot", fig.cap='(ref:team-structures-all)'}
```

Participants played the "Totems" game adapted from @Derex:2015cbb. Their task
was to discover how to build tools with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). More advanced tools
produce larger and more intricate totems, resulting in higher performance
scores.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6
items were available to problem solvers at the start of the game. New items
could be produced through the combination of different items (more than one
arrow points to the item) or through the refinement of a single item (a single
arrow points to the item). The axe is required to construct the first totem
pole.

```{r landscape-sample, fig.height=4, fig.cap="(ref:landscape-sample)"}
draw_image("landscape-sample.gv", package = "totems")
```

Participants discovered new tools by combining existing items. Participants
could combine up to four items at a time (with replacement), meaning the
initial six items could form a total of `r methods$n_unique_guesses_6`
combinations. Of all possible combinations, very few resulted in new items. For
example, of all the guesses that could be formed from the initial items, only
three (`r methods$n_unique_guesses_6_pct`%) yielded new tools (Fig. 2). As
participants accumulate solutions, the combinatorial complexity of the problem
space increases exponentially such that the discovery of more complex tools is
less likely to happen by chance.

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record. Participants could review their past innovations and see the
recipes for their previous discoveries. **Participants assigned to generations
after the first inherited the innovation record of the previous generation
participant.** From the beginning of the experiment, these participants could
review the recipes for all the innovations that had been discovered by their
ancestor. Note that the participants inherited the recipes, but not the tools
themselves. In order use these tools in further combinations, the tools and all
of their constituent parts first had to be recreated.

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the Workshop and selecting the Try button. If the guess
created an item, it could be dragged into the Stock panel, and used again. Once
segments of a totem were made through cutting, carving, and painting, they could be
dragged into the Totem panel for scoring.

```{r interface, fig.cap="(ref:interface)"}
draw_image("gameplay", "totems")
```

## Participants {-}

Participants were students at the University of Wisconsin-Madison who completed
the experiment in exchange for course credit. Students received credit based on
the duration of each completed session, 25 minutes or 50 minutes. Participants
assigned to the Isolated, multi-session condition who completed all four
sessions were awarded additional credit.

Participants were recruited until there were at least 40 teams in each strategy.
Most strategies, other than four-person synchronic teams, ended up receiving
more than 40 teams. The reason for this is that participants were scheduled in
groups of 4 participants at a time, and frequently one or more participants
would fail to show up for their extra credit opportunity. In these cases, the
remaining participants were run as a combination of diachronic and
two-person synchronic participants, resulting in a higher-than-expected
number of participants in those conditions.

(ref:condition-counts) Total participant and team recruitment.

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "(ref:condition-counts)")
```

# Experiment 1 {-}

```{r exp1, include=FALSE}
```

In Experiment 1, we compared the simplest form of diachronic inheritance---a
two-person diachronic team---to two alternative strategies for allocating the
same 50 minutes of labor: isolated individuals working alone and two-person
sychronic teams (Fig. \@ref(fig:team-structures-exp2)).

(ref:team-structures-exp2) Problem solving strategies compared in Experiment 2.

```{r team-structures-exp2, engine="dot", out.width="50%", fig.cap="(ref:team-structures-exp2)"}
```

## Methods {-}

(ref:exp1-participants) Participants in Experiment 1.

```{r exp1-participants, results="asis"}
kable(Exp1N, caption = "(ref:exp1-participants)")
```

## Results {-}

Diachronic problem solvers with inheritance
significantly exceeded the number of innovations discovered in a single
25-minute session, `r exp1$DG2_v_DG1`. We found that diachronic
problem solvers with inheritance outperformed isolated individuals given 50
minutes, however this difference was not reliable, `r exp1$DG2_v_I50`.
Conversely, two-person synchronic teams outperformed second generation diachronic
teams, but again, this different was not reliable, `r exp1$S2_v_DG2`. These
results, although they do not demonstrate a clear benefit to inheritance in terms of
overall performance, are in line with our conclusion that inheritance does not
negatively impact problem solving relative to alternative strategies (Fig.
\@ref(fig:innovations-50min).

Comparing two-person teams (both diachronic and synchronic) to isolated
individuals revealed a benefit to teamwork in this task,
`r exp1$teamwork_stats`, with no residual difference between diachronic and
synchronic strategies, `r exp1$teamwork_residual`. This result suggests that
maximizing individual learning time is not the most productive way to improve
problem solving efficiency, instead demonstrating that teamwork is beneficial
even when controlling for labor hours, instead of calendar hours.

(ref:innovations-50min) Innovations discovered in 50 labor minutes by strategy. **A.**
Rate of innovation by labor hours. **B.** Final innovations by strategy. Bars with
errors show model predictions with ±1 standard error.

```{r innovations-50min, fig.width=5, fig.height=3, fig.cap="(ref:innovations-50min)"}
grid.arrange(
  innovation_rate_50min_plot + ggtitle("A"),
  num_innovations_50min_plot + ggtitle("B"),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

Next we investigated the impact of inheritance on a per-innovation basis.
We hypothesized that inheritance would reduce the overall number of guesses
expended in discovering each item over the alternative strategies.
To count the number of guesses for each innovation, we tallied all guesses
made from the moment in which a new item was adjacent to the current inventory,
until that item was discovered. This way we only included guesses that could
have created the item.

We found that diachronic inheritance did reduce the total number of guesses
made for each new item relative to the same items being guessed with an
isolated and synchronic strategy (Fig. \@ref(fig:guesses-per-item)A). However,
we found that this overall benefit to inheritance in terms of fewer guesses
per item did not extend to items that had not been discovered by the previous
generation. Comparing the number of guesses per item only for items that
were newly discovered by the second generation diachronic problem solvers
did not reveal a benefit to inheritance (Fig. \@ref(fig:guesses-per-item)B).

(ref:guesses-per-innovation) Average guesses per innovation.

```{r guesses-per-innovation, fig.width=5}
guesses_per_item_treatment_plot
```

Finally we investigated the guessing strategy employed in each strategy.
We were interested in whether inheritance caused second generation
diachronic problem solvers to guess differently than players in
the other strategies. Our results are shown in Fig. \@ref(fig:guess-types).

```{r guess-types, fig.width=5}
prop_guess_types_50min_plot
```

## Discussion {-}

# Experiment 2 {-}

(ref:team-structures-exp3) Problem solving strategies compared in Experiment 3.

```{r team-structures-exp3, engine="dot", out.width="70%", fig.cap="(ref:team-structures-exp3)"}
```

## Methods {-}

## Results {-}

## Discussion {-}

# Experiment 3 {-}

(ref:team-structures-exp4) Problem solving strategies compared in Experiment 4.

```{r team-structures-exp4, engine="dot", out.width="50%", fig.cap="(ref:team-structures-exp4)"}
```

## Methods {-}

## Results {-}

## Discussion {-}

# General Discussion {-}

# References {-}

\setlength{\parindent}{-0.1in}
\setlength{\leftskip}{0.15in}
\noindent
