---
title: "The impact of inheritance on problem solving"
author:
  - name: Pierce Edmiston
    affiliation: 1
  - name: Maxime Derex
    affiliation: 2
  - name: Gary Lupyan
    affiliation: 1
affiliation:
  - n: 1
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - n: 2
    department: Biosciences Department
    university: University of Exeter
email: pedmiston\@wisc.edu
address: 1202 W. Johnson Street, Department of Psychology, University of Wisconsin-Madison, Madison, WI 53703.
abstract: |
  Humans are able to accumulate tools and other solutions to problems like no
  other species of animal. A consequence of our system of cultural inheritance
  is that future generations rarely have to start from scratch when solving new
  problems, as demonstrated by the prevalence of reuse in the history of human
  technology. We investigated how inheriting solutions from a previous
  generation impacted problem solving by comparing the performance of teams who
  share solutions across generations to alternative strategies for solving the
  same set of problems but without inheritance. We found that inheritance was
  effective in that participants were reliably able to solve more problems than
  their ancestors, and that the benefits of inheritance were similar to
  individuals given the same total amount of time and teams working in parallel.
  We also found that inheritance caused participants to engage in a more
  efficient guessing strategy, and as a consequence were less likely to become
  fixated on incorrect solutions. These results suggest that human cultural
  inheritance does more than give future generations a head start on problem
  solving, and may actually change the way new problems are solved.
bibliography: "references.bib"
csl: "templates/plos-one.csl"
output:
  bookdown::pdf_document2:
    template: templates/plos_latex_template.tex
    keep_tex: yes
    number_sections: no
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  dpi = 300,
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)

# Read all the diagrams made with graphviz in the totems package
chunks <- c(
  "team-structures",
  "team-structures-all",
  "team-structures-exp1",
  "team-structures-exp2",
  "team-structures-exp3",
  "team-structures-exp4"
)
sapply(chunks, read_graphviz_chunk, package = "totems")
```

```{r setup, include=FALSE, cache=FALSE}
```

# Introduction

```{r intro, include=FALSE}
```

Humans are effective problem solvers, having solved a wide range of problems
related to foraging, hunting, and preparing food, while surviving predators,
each other, and a large range of terrestrial environments [@Boyd:2017diff;
@Fernandez:2001civ]. What has enabled our success in being able to solve such a
diverse set of problems? Some have suggested that the answer lies more in our
ability to inherit knowledge from others than our ability to make discoveries by
ourselves [@Richerson:2008not; @Henrich:2015srt; @Boyd:2017diff]. Humans possess
a number of advanced social learning abilities including teaching through verbal
instruction and imitation that provide reliable ways of transferring problem
solving knowledge across individuals [@Dean:2012gv]. If problem solving
knowledge can be acquired via social learning, then future generations can adapt
and improve it, allowing cultures to accumulate technological complexity over
generations.

However, the ability to learn socially from others is not sufficient to explain
cumulative cultural evolution. Although social learning was once thought to be
rare in the animal kingdom [e.g., @Thorndike:1898], it has now been documented
in a range of species from chimps [@Whiten:1999cic] to fish [@Laland:1997ts]
and even bees [@Alem:2016fu]. If cumulative cultural evolution depended simply
on social learning, we might expect these species to likewise show evidence of
cumulative cultural evolution, yet such evidence is notably lacking
[@Dean:2012gv; @Tennie:2009gaa]. There are a few exceptions: New Caledonian
crows have been argued to have improved the grub-skewers they craft from
pandanus leaves [@Hunt:2003dj], and at least one population of chimpanzees has
evolved a more advanced termite probe [@Sanz:2009ja], but for the most part
tool use in animals has remained largely unchanged [cf. @Mercader:2007fd].

Humans, in contrast, have demonstrated a remarkable ability to adapt
and improve the tools and other innovations discovered by others. The
history of human technology is argued to be better understood as a process
of gradual refinement and repurposing rather than punctuated advances brought by
the discoveries of rare geniuses [@Basalla:1989wq; @Sole:2013cm]. Rapid
refinement of inherited innovations has not always been the case over the course
of human history, as demonstrated by the long periods in the archaeological
record of slow or stagnant growth in stone tool complexity [@Torre:2011jl;
@Lycett:2008di]. As to what it was that allowed our ancestors to begin to
develop more complex tools, an answer that has been proposed is that humans
evolved more robust ways of transmitting cultural information to the next
generation, allowing future generations of problem solvers to more quickly learn
the skills honed by their ancestors, and thus giving them more time to make
improvements to those technologies [@Sterelny:2012ea]. On this view,
human cultural evolution has been defined not by our ability to inherit the
skills of our ancestors, but by the ability to exceed them.

We investigated the human propensity to exceed our ancestors in a problem
solving task using a transmission chain paradigm (Fig. \@ref(fig:team-structures-all)).
Previous research using this paradigm has found that problem solving performance
for simple tasks like folding paper airplanes and building tours out of
spaghetti can accumulate over generations [@Caldwell:2008gf], although the
mechanisms are unclear [@Caldwell:2009fl]. @Morgan:2015hu more carefully
manipulated the mechanisms of transmission, and found that teaching, and
especially verbal instruction, was crucial for successful transmission of
skilled behaviors along chains. Here we controlled for successful transmission
by allowing future generations to inherit the discoveries made by the previous
generation, and measured the ability of participants to exceed the solutions
they inherited.

# Methods

```{r methods, include=FALSE}
```

To understand how problem solving knowledge accumulates through vertical
transmission, we used a transmission chain paradigm where participants were
assigned to a single generation within a four-generation chain. Each
participant attempted the same problem solving task for 25 minutes. The
solutions that each participant had discovered by the end of the session were
passed on to be inherited by a participant in the next generation of the
chain. Thus, participants assigned to generations after the first began the
problem solving task with solutions inherited from the previous generation.

(ref:team-structures-all) All experimental conditions. The **Diachronic**
condition is the treatment condition: problem solving with inheritance. First
generation diachronic participants began without any previous solutions and
worked on the problem solving task for 25 minutes. Diachronic participants
assigned to generations 2-4 inherited the solutions discovered by the previous
generation at the start of the problem solving task. The two control conditions
were Isolated and Synchronic. **Isolated** participants either completed a
single 50 minute session, or returned for 4 subsequent 25 minute sessions. Upon
returning, these participants "inherited" the solutions they had created in the
previous session. **Synchronic** participants completed the problem solving task
in groups of 2 or 4 individuals, each participating for 25 minutes. Synchronic
participants collaborated in parallel, sharing solutions with all team members
as they were discovered.

```{r team-structures-all, engine="dot", out.width="50%", fig.cap="(ref:team-structures-all)"}
```

Participants played the "Totems" game adapted from @Derex:2015cbb. Their task
was to discover how to build tools with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). More advanced tools produce larger and more
intricate totems, resulting in higher performance scores.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6 items were available to problem solvers at the start of the game. New items could be produced through the combination of different items (more than one arrow points to the item) or through the refinement of a single item (a single arrow points to the item). The axe is required to construct the first totem pole.

```{r landscape-sample, fig.height=6, fig.cap="(ref:landscape-sample)"}
draw_image("landscape-sample.gv", package = "totems")
```

Participants discovered new tools by combining existing items. Participants
could combine up to four items at a time (with replacement), meaning the
initial six items could form a total of `r methods$n_unique_guesses_6`
combinations. Of all possible combinations, very few resulted in new items. For
example, of all the guesses that could be formed from the initial items, only
three (`r methods$n_unique_guesses_6_pct`%) yielded new tools (Fig. 2). As
participants accumulate solutions, the combinatorial complexity of the problem
space increases exponentially such that the discovery of more complex tools is
less likely to happen by chance.

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record. Participants could review their past innovations and see the
recipes for their previous discoveries. **Participants assigned to generations
after the first inherited the innovation record of the previous generation
participant.** From the beginning of the experiment, these participants could
review the recipes for all the innovations that had been discovered by their
ancestor. Note that the participants inherited the recipes, but not the tools
themselves. In order use these tools in further combinations, the tools and all
of their constituent parts first had to be recreated.

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the Workshop and selecting the Try button. If the guess
created an item, it could be dragged into the Stock panel, and used again. Once
segments of a totem were made through cutting, carving, and painting, they could be
dragged into the Totem panel for scoring.

```{r interface, fig.cap="(ref:interface)"}
draw_image("gameplay", "totems")
```

## Participants

Participants (N=`n_recruited`) were students at the University of
Wisconsin-Madison who completed the experiment in exchange for course credit.
Students received credit for each completed session based on the duration of the
session, 25 minutes or 50 minutes. Participants in the Isolated, multi-session
experiment who completed all four sessions were awarded additional credit.

Participants were recruited until there were at least 40 teams in each strategy.
Most strategies, other than four-person synchronic teams, ended up receiving
more than 40 teams. The reason for this is that participants were scheduled in
groups of 4 participants at a time, and frequently one or more participants
would fail to show up for their extra credit opportunity. In these cases, the
remaining participants were run as a combination of diachronic and
two-person synchronic participants, resulting in a higher-than-expected
number of participants in those conditions.

Data from individual participants (N=`n_excluded`) were excluded from the final
analyses for the following reasons. Most of these (N=`n_errors`) were excluded
because they experienced errors in the computer game. Others
(N=`n_noncomplaint`) were noncomplaint, giving up on the problem solving task
(N=`n_gave_up`).

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "Total participant and team recruitment.")
```

# Experiment 1

```{r exp1, include=FALSE}
```

In Experiment 1, we measured problem solving ability over four generations of
diachronic problem solving (Fig. \@ref(fig:team-structures-exp1)). We answered
the following questions about the impact of diachronic inheritance on problem
solving. First, we determined whether and by how much future generations of
problem solvers were able to recreate and exceed the solutions obtained by their
ancestors. Second, we explored the impact of inheritance on future problem
solving by examining whether the size of the inheritance impacted future
problem solving ability.

(ref:team-structures-exp1) Diachronic problem solving examined in
Experiment 1. Participants were assigned to four-person diachronic teams.
Each participant completed the same problem solving task for 25 minutes.
Participants in generations 2-4 began the problem solving task with the
solutions that were discovered by the previous generation.

```{r team-structures-exp1, engine="dot", out.width="25%", fig.cap="(ref:team-structures-exp1)"}
```

## Methods

Participants played the Totems game
for a single 25-minute session. Each participant was assigned to a single
generation within a single four-person team (($N_{participants}$=`r exp1$n_participants`) , $N_{teams}$=`r exp1$n_teams`).

Participants assigned to the first generation of a new team played to
Totems game for 25-minutes. Their solutions were inherited by one
(and only one) generation 2 player, and so on through 4 generations
of diachronic collaboration.

(ref:exp1-participants) Participants in Experiment 1.

```{r exp1-participants, results="asis"}
kable(Exp1Participants, caption = "(ref:exp1-participants)")
```

## Results

### Performance by generation

We found that problem solving knowledge accumulated over generations such that
participants in later generations were able to solve more problems in the same
amount of time than their ancestors (Fig. 3). We tested this in multiple ways.
First, we used Page's trend test (a repeated measures test for monotonicity),
and found evidence that the number of tools discovered within each chain
increased over generations `r exp1$page_test`. Next, we tested whether the
extent to which participants were able to exceed their ancestors changed over
generation. We fit a hierarchical regression model to the number of tools
discovered in each generation with these effects nested by chain. On average,
second generation participants were able to discover `r exp1$gen0_slope` more
tools than first generation participants, `r exp1$gen0_slope_stats`. This
effect decreased by `r exp1$gen0sqr_slope` each generation for third and fourth
generation participants, `r exp1$gen0sqr_slope_stats`. This diminishing return
suggests that future generations have a harder time exceeding their ancestors.

```{r fig3, fig.cap="Performance by generation. Each of the thin green lines is a chain. The thick blue line is the model predictions with ±1 standard error."}
innovations_by_generation_plot
```

Although later generation participants made fewer new discoveries, the
discoveries they made were often more complex. We measured complexity as the
combinatorial complexity of the number of items in the participants' inventories
at the time a discovery was made. We found that when discoveries were weighted
by combinatorial complexity, participants solved more complex problems
with each generation, `r exp1$difficulty_by_generation_mod_slope_stats`,
and this benefit did not diminish for later generations,
`r exp1$difficulty_by_generation_quad_mod_slope_stats`.

### Performance by inheritance size

```{r fig4, fig.height=8, fig.cap="Performance by inheritance size. A. Number of tools created relative to those inherited. The dotted line is a reference with slope=1 such that points above the line indicate future generations exceeding their ancestors. B. Number of new tools relative to those inherited. The same reference line is now shown horizontally. C. Change in accumulated difficulty score. Analogous to B, but with innovations weighted by complexity. In B and C, the error ranges specify the model predictions with ±1 standard error."}
grid.arrange(
  innovations_created_and_inherited_plot + ggtitle("A"),
  new_innovations_plot + ggtitle("B"),
  delta_difficulty_plot + ggtitle("C"),
  ncol = 1
)
```

Because there is no difference between a second generation participant
who inherits 10 tools and a fourth generation participant who inherits
the same tools, we also looked at problem solving performance relative
to the number of tools that were inherited (Fig. 4). As the number of
inherited tools increased, the number of new tools discovered decreased,
`r exp1$inheritance_size_slope_stats` (Fig. 4B). That participants who
inherited more items had a harder time exceeding them is expected based on the
analysis of performance over generations reported above. However, when tools
are weighted by combinatorial complexity, the amount by which participants are
able to exceed their ancestors actually increases for larger inheritance sizes,
`r exp1$delta_diff_slope_stats` (Fig. 4C), though we are hesitant to draw strong
conclusions from this finding as it may be explanable by the characteristics of
the solution landscape, a point we return to in the Discussion.

We next asked whether inheriting more tools had an impact on the rate of new
tool discovery. To answer this question, we first needed to identify the length
of time spent recreating inherited tools as opposed to discovering new ones.
Participants took on average `r exp1$mean_learning_time_min` minutes of the 25
minute session (`r exp1$proportion_learning_time`%) to recreate the inherited
tools---a portion of the experiment we refer to as the learning period. The
length of the learning period correlated positively with the number of inherited
tools (Fig. 5).

```{r fig5, fig.cap="Learning rates. Correlation between the number of tools inherited and the time it took to recreate the inherited items. Outliers who were appear unwilling or unable to recreate the inherited items are shown as X's, but included in all analyses."}
learning_rates_plot
```

We then calculated discovery rates for each participant: the number of
novel discoveries relative to the amount of time out of the 25 minute
session spent discovering new tools (Fig. 6). The overall rate of
problem solving was `r exp1$minutes_per_new_innovation` minutes per new
innovation (`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based on
the number of inherited tools, as revealed by comparing a model predicting
unique innovations from playing time alone to one predicting unique
innovations from the interaction between playing time and inheritance size,
`r exp1$playing_time_modcomp`. This suggests that inheriting more tools did not
meaningfully diminish the rate of future problem solving.

```{r fig6, fig.cap="Problem solving rate in the discovery stage. Discovery time is the amount of time out of a 25 minute session dedicated to discovering new innovations that were not discovered by an ancestor. The line shows the predictions of the hierarchical regression model with ±1 standard error. The slope of this line did not significantly vary based on the size of the inheritance."}
playing_time_plot
```

### Performance by inheritance

Next we compared performance in the first generation, where participants did
not benefit from any previous ancestor, to performance in generations 2-4.
We found that participants who benefited from inheritance on average made
fewer guesses per item than when the same items were attempted in the first
generation, `r exp1$guesses_per_item_by_inheritance`. As expected, this effect
is especially pronounced for the items that were inherited. We did not find
any evidence that inheritance had an affect on the number of guesses per novel
item, `r exp1$guesses_per_new_item_by_inheritance_mod` (Fig. 7). This finding
suggests that although inheritance benefits participants in recreating inherited
solutions, it does not confer any benefits to future problem solving.

```{r fig7, fig.height=6, fig.cap="Guesses per tool by inheritance condition. Each line is the average number of guesses it took to discover a particular tool in either inheritance condition. Error bars show ±1 standard of the model predictions. A. Overall guesses per tool, including those inherited. B. Guesses per novel tool."}
grid.arrange(
  guesses_per_item_by_inheritance_plot + ggtitle("A"),
  guesses_per_new_item_by_inheritance_plot + ggtitle("B"),
  ncol = 1
)
```

Finally, we analyzed the overall guessing strategy employed by participants
who inherited from an ancestor versus those who did not. Guesses were
categorized into one of four types: redundant guesses, repeat items,
unique guesses, and unique items. The overall proportion of guesses
made by participants based on their inheritance condition is shown in
Fig. 8. We found that participants who inherited solutions at the start
tended to make fewer redundant guesses than first generation participants,
`r exp1$num_redundant_guesses`, a difference of 12%, `r exp1$prop_redundant_guesses`.
This result suggests that starting off the problem solving task with
solutions inherited from a previous generation influenced the way
in which future guesses are generated.

```{r fig8, fig.width=4, fig.cap="Guessing strategies by inheritance condition. Redundant guesses were incorrect guesses made once before. Repeat items were recreated tools. Unique guesses were incorrect guesses never made. Unique items were unique guesses that created a new tool."}
prop_guess_types_50min_plot
```

## Discussion

In Experiment 1, we measured the problem solving ability of four-person diachronic teams, and found that diachronic problem solvers were consistently able to cumulatively improve upon the solutions they inherited. Although this finding is not surprising given that participants started a problem solving task with some of the problems already solved, it is important to highlight that solving new problems first required recreating all of the inherited solutions, which took time and effort. Once recreated, some participants were unable to solve any new problems. Others were unable to recreate all of their inherited solutions in the 25-minute session---although this behavior is likely attributable to non-compliant behavior or a misunderstanding of the task. Nonetheless, the presence of these outliers simply highlights that diachronic problem solving, as measured in this paradigm, is not guaranteed to improve cumulatively, and explains why it is important to first establish that
performance does indeed accumulate in diachronic teams.

We next asked whether the size of the inheritance had an impact on the likelihood of being improved, and found multiple, conflicting results. When all innovations are weighted equally, the effect of inheritance decreased as inheritance size increased, which could mean that that larger inheritances had a negative effect on problem solving. However, when the innovations are weighted by the difficulty of the solution in terms of the combinatorial complexity of the solution, the benefit of inheritance did not decrease, as problem solvers who inherited more solutions were able to solve more combinatorially difficult problems. Finally, when controlling for the amount of time spent solving new problems as opposed to recreating inherited ones, the rate of problem solving (solutions per minute) also did not decrease with larger inheritances. However, this conclusion is limited by the collinear relationship between inheritance size and time solving unique problems.

We believe the source of these conflicting interpretations is an underlying inability to understand the impact of inheritance on problem solving without comparison to a control condition. In Experiment 2, we attempt to understand the impact of inheritance on problem solving relative to the benefits that may accrue from individual learning over four separate 25-minute sessions.

# Experiment 2

```{r exp2, include=FALSE}
```

In Experiment 2, we compared the simplest form of diachronic inheritance---a
two-person diachronic team---to two alternative strategies for allocating the
same 50 minutes of labor: isolated individuals working alone and two-person
sychronic teams (Fig. \@ref(fig:team-structures-exp2)). In our analysis we focus
primarily on the comparison of the second generation diachronic problem solvers,
who began the problem solving task with the solutions inherited from a first
generation problem solver, to problem solvers utilizing the alternative
strategies to solve the same set of problems. In addition to measuring the
impact of inheritance on overall performance, we also investigate whether
diachronic collaboration causes problem solvers to approach the task differently
than the alternative strategies. To answer this question, we investigate the
average number of guesses expended to discover each new item as well as the
distribution of guess types made by problem solvers in each strategy.

(ref:team-structures-exp2) Problem solving strategies compared in Experiment 2.

```{r team-structures-exp2, engine="dot", fig.cap="(ref:team-structures-exp2)"}
```

## Methods

Participants assigned to the diachronic condition completed the Experiment as
reported above. Participants in the synchronic condition were randomly assigned
to two-person teams. Synchronic problem solvers each attempted the Totems game
in a single 25-minute session at separate computers. Synchronic team membership
was anonymized such that participants were not explicitly made aware of who
their teammate was, and were not allowed to interact or communicate with their
teammate other than through the computer interface. The only interaction allowed
through the computer interface was the sharing of unique innovations. Once
a synchronic problem solver discovered a new item, the recipe for its creation
could be viewed and copied by the other problem solver in the team. Isolated
individuals were recruited for a single 50-minute session.

(ref:exp2-participants) Participants in Experiment 2. Data for diachronic
participants was taken from the first and second generations of Experiment 1.

```{r exp2-participants, results="asis"}
kable(Exp2N, caption = "(ref:exp2-participants)")
```

## Results

As reported in Experiment 1, diachronic problem solvers with inheritance
significantly exceeded the number of innovations discovered in a single
25-minute session, `r exp2$DG2_v_DG1`. In Experiment 2, we found that diachronic
problem solvers with inheritance outperformed isolated individuals given 50
minutes, however this difference was not reliable, `r exp2$DG2_v_I50`.
Conversely, two-person synchronic teams outperformed second generation diachronic
teams, but again, this different was not reliable, `r exp2$S2_v_DG2`. These
results, although they do not demonstrate a clear benefit to inheritance in terms of
overall performance, are in line with our conclusion that inheritance does not
negatively impact problem solving relative to alternative strategies (Fig.
\@ref(fig:innovations-50min).

Comparing two-person teams (both diachronic and synchronic) to isolated
individuals revealed a benefit to teamwork in this task,
`r exp2$teamwork_stats`, with no residual difference between diachronic and
synchronic strategies, `r exp2$teamwork_residual`. This result suggests that
maximizing individual learning time is not the most productive way to improve
problem solving efficiency, instead demonstrating that teamwork is beneficial
even when controlling for labor hours, instead of calendar hours.

(ref:innovations-50min) Innovations discovered in 50 labor minutes by strategy. **A.**
Rate of innovation by labor hours. **B.** Final innovations by strategy. Bars with
errors show model predictions with ±1 standard error.

```{r innovations-50min, fig.width=6, fig.height=4, fig.cap="(ref:innovations-50min"}
grid.arrange(
  innovation_rate_50min_plot + ggtitle("A"),
  num_innovations_50min_plot + ggtitle("B"),
  nrow = 1,
  widths = c(0.6, 0.4)
)
```

Next we investigated the impact of inheritance on a per-innovation basis.
We hypothesized that inheritance would reduce the overall number of guesses
expended in discovering each item over the alternative strategies.
To count the number of guesses for each innovation, we tallied all guesses
made from the moment in which a new item was adjacent to the current inventory,
until that item was discovered. This way we only included guesses that could
have created the item.

We found that diachronic inheritance did reduce the total number of guesses
made for each new item relative to the same items being guessed with an
isolated and synchronic strategy (Fig. \@ref(fig:guesses-per-item)A). However,
we found that this overall benefit to inheritance in terms of fewer guesses
per item did not extend to items that had not been discovered by the previous
generation. Comparing the number of guesses per item only for items that
were newly discovered by the second generation diachronic problem solvers
did not reveal a benefit to inheritance (Fig. \@ref(fig:guesses-per-item)B).

```{r guesses-per-item, fig.width=6, fig.height=4}
grid.arrange(
  guesses_per_item_by_inheritance_plot + ggtitle("A"),
  guesses_per_new_item_by_inheritance_plot + ggtitle("B"),
  nrow = 1
)
```

Finally we investigated the guessing strategy employed in each strategy.
We were interested in whether inheritance caused second generation
diachronic problem solvers to guess differently than players in
the other strategies. Our results are shown in Fig. \@ref(fig:guess-types).

```{r guess-types, fig.width=6}
prop_guess_types_50min_plot
```

## Discussion

# Experiment 3

(ref:team-structures-exp3) Problem solving strategies compared in Experiment 3.

```{r team-structures-exp3, engine="dot", fig.cap="(ref:team-structures-exp3)"}
```

## Methods

## Results

## Discussion

# Experiment 4

(ref:team-structures-exp4) Problem solving strategies compared in Experiment 4.

```{r team-structures-exp4, engine="dot", fig.cap="(ref:team-structures-exp4)"}
```

## Methods

## Results

## Discussion

# Discussion

# References

\setlength{\parindent}{-0.1in}
\setlength{\leftskip}{0.15in}
\noindent
