---
title: "The impact of team structure on problem solving performance"
author:
  - { name: Pierce Edmiston, affiliation_id: 1 }
  - { name: Maxime Derex, affiliation_id: 2 }
  - { name: Gary Lupyan, affiliation_id: 1 }
affiliation:
  - id: 1
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - id: 2
    department: Biosciences Department
    university: University of Exeter
email: pedmiston\@wisc.edu
address: 1202 W. Johnson St., Department of Psychology, University of Wisconsin, Madison, WI 53703.
bibliography: "references.bib"
csl: "templates/plos-one.csl"
output:
  bookdown::pdf_document2:
    template: templates/plos-one.tmpl
    keep_tex: yes
    number_sections: no
---
```{r team-structures-config, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",

  # PLOS One figure requirements.
  # Width: 789 – 2250 pixels (at 300 dpi). Height maximum: 2625 pixels (at 300 dpi).
  fig.width = 3,
  fig.height = 3,
  dpi = 300
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "team-structures.R"))
```

```{r Setup}
```

# Introduction {-}

```{r Intro}
```

Cultural evolution involves both vertical and horizontal transmission
[@Boyd:1985cu]. Vertical transmission is the process of passing cultural
products from parent to child, whereas horizontal transmission involves passing
cultural products among peers in the same generation.


Some cultural products,
such as religious beliefs, are more likely to be learned from parents than from
peers [@Newcomb:1937vy]. But far more cultural products are shared among
unrelated individuals via horizontal transmission. As a result, horizontal
transmission has been argued to be the default process in cultural evolution
[@Kroeber:1948ant].

Whether or not individuals learn from parents or peers may have consequences
for the distribution of traits in a population. For example, @Bettinger:1999vv
argued that high correlations among projectile point designs recovered in the
Great Basin region indicated a culture of biased horizontal transmission, in
which members of a society tend to copy the most prestigious hunters.
Projectile points found in a different region over the same time period were
very dissimilar from one another, indicating that each family may have
innovated their own point designs through generations of vertical transmission.
This interpretation has been supported by experimental and computational
simulations showing that fully connected networks of problem solvers result in
greater convergence and smaller variation in design than independently evolving
lineages [@Mesoudi:2008vi; @Mesoudi:2008we].

The distinction between vertical and horizontal transmission is motivated by
analogous processes in genetic evolution. In genetic evolution, vertical
transmission occurs via reproduction, whereas horizontal transmission is the
transfer of genetic material by any means _other_ than reproduction
[@Soucy:2015gp]. Thus, the difference between vertical and horizontal
transmission of genetic material is motivated by an underlying difference in
mechanism, as the cellular mechanisms involved in reproduction (e.g., meiosis)
are different from those involved in horizontal transfer (e.g., conjugation).

In contrast, there is no such difference in mechanism between the vertical and
horizontal transmission of cultural products. All processes of cultural
transmission involve social learning, in which one person learns a behavior
from the observation or interaction with someone else. The cognitive mechanisms
that allow for social learning are functionally the same regardless of who is
being copied. For example, the ability to reason about other's mental states,
also known as theory of mind, is believed to be independent of who is being
reasoned about.

Although social learning is involved for both vertical and horizontal transmission,


When considering the consequences of vertical and horizontal transmission,
horizontal transmission is considered advantageous because it provides a level
of redundancy that combats the potential for information loss due to
transmission error. If every instance of cultural transmission has some
probability of transmission error, the ability to share innovations widely
within a given generation increases the likelihood that it will survive to be
used and improved by the next generation. As a case study of this phenomena,
@Henrich:2004ht modeled the apparent loss of technological complexity among
Tasmanian islanders over an 8,000 year period as the result of a decreasing
population size. With fewer replicators within each generation to learn complex
skills, eventually transmission errors resulted in the loss of some
technologies. This effect has been demonstrated in a behavioral experiment
[@Derex:2013kk]. To overcome the potential for information loss, human cultures
must increase in size, and excel at propagating innovations horizontally to as
many people in each generation as possible.

Another way in which horizontal transmission is believed to have different
consequences than purely vertical transmission is in providing multiple
behavioral models to learn from. A common assumption in models of the cultural
evolutionary process is that individuals learn behaviors from just one model.
However, computational models by @Enquist:2010gw have shown that when
individuals learn from just one parent (as in vertical transmission),
cumulative culture is unlikely to occur. The reason is that, along chains of
vertical transmission, any error in transmission is necessarily propagated to
future generations, increasing the likelihood of maladaptive traits becoming
established in populations. In contrast, when individuals can learn from
multiple cultural models and adaptively filter among the observed behaviors to
determine the best person to copy, then the cumulative cultural process can
proceed as expected [@Rendell:2011gk].

Horizontal transmission is beneficial in providing redundancy among replicators
and opportunities for adaptive filtering, but are there any advantages to
vertical transmission? To answer this question, we compared the effectiveness
of vertical and horizontal transmission in a problem solving task.

In this research, we distinguish between vertical and horizontal transmission
along two dimensions of time. Vertical transmission is defined as coordination
over generations, such that innovations discovered by one generation are passed
on to be used and improved by the next generation. Horizontal transmission is
defined as coordination in parallel and in the same generation. We refer to
labor that is coordinated via vertical transmission as "diachronic
collaboration" and labor that is coordinated via horizontal transmission as
"synchronic collaboration". We compared the effectiveness of diachronic
and synchronic collaboration to a problem solving strategy involving
isolated individuals working alone.

In our comparison of diachronic, synchronic, and isolated strategies,
we controlled for **labor time**. Labor time can be contrasted with
calendar time and learning time (Fig. \@ref(fig:types-of-time)). We
controlled for labor time so that our results can be interpreted in
terms of the effect each problem solving strategy has on individual
problem solving ability. Without controlling for labor time, any differences
between the three strategies are more difficult to interpret. For example,
if only calendar hours are controlled, an isolated individual given 30 minutes
could be compared to a synchronic team given 30 minutes _regardless of the
size of the synchronic team_. This fact makes it difficult to understand
the effect of teamwork on individual performance when only controlling for
calendar time [@Propp:2003kj; @Tindale;1992ub].

(ref:types-of-time) Team structures controlling for labor time. **Left.** Team
structures. Each node is a person. **Middle.** Relationship between labor time
and calendar time. Synchronic teams spend labor time more quickly than the
other strategies. **Right.** Learning times by strategy, showing that an
isolated strategy provides the most time for individual learning.

```{r types-of-time, fig.width=5, fig.height=2, fig.cap="(ref:types-of-time)"}
gridExtra::grid.arrange(
  crotchet::read_graphviz("team-structures", package = "totems"),
  gg_time,
  gg_person,
  nrow = 1,
  widths = c(0.25, 0.375, 0.375)
)
```

# Methods {-}

```{r Methods}
```

## Team structures {-}

Participants were randomly assigned to complete the same problem solving task in
different team structure conditions (Fig. \@ref(fig:team-structures-all)).
Participants in the diachronic condition worked on the problem solving task one
at a time. At the end of a diachronic participant's session, the solutions they
discovered were passed on to be inherited by a single participant assigned
to the next generation of the same team.

Participants assigned to the synchronic team structure worked independently and
in parallel in teams of two or four. Their labor was coordinated such that
solutions discovered by one team member were immediately shared with the other
members of the team.

Isolated participants completed the problem solving task as individuals, only
benefiting from the solutions they could discover on their own. Isolated participants
completed the experiment in a single 50 minute session or, in a different condition,
came back for four 25 minute sessions.

(ref:team-structures-all) Team structures. Diachronic participants worked one
at a time, each inheriting the solutions discovered by the previous generation,
labeled G1-G4. Isolated participants were recruited for a single 50 minute
session, or returned for four 25 minute sessions, labeled S1-S4. Synchronic
participants were assigned to teams of two or four individuals that all worked
at the same time.

```{r team-structures-all, out.width="80%", fig.cap='(ref:team-structures-all)'}
crotchet::draw_graphviz("team-structures-all", package = "totems")
```

## Problem solving task {-}

Participants played the "Totems game" adapted from @Derex:2015cbb. Their task
was to discover tools that could be used with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). More advanced tools
produce larger and more intricate totems, resulting in higher performance
scores.

Participants could combine up to four items at a time (with replacement),
meaning the initial six items could form a total of
`r methods$n_unique_combinations_6` combinations. Of all possible combinations,
very few resulted in new items. For example, of all the combinations that could
be formed from the initial items, only three
(`r methods$n_unique_combinations_6_pct`%) yielded new tools.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6
items were available to problem solvers at the start of the game. New items
could be produced through the combination of different items (more than one
arrow points to the item) or through the refinement of a single item (a single
arrow points to the item). The axe is required to construct the first totem
pole.

```{r landscape-sample, fig.width=5, fig.cap="(ref:landscape-sample)"}
draw_image("landscape-sample.gv", package = "totems")
```

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record (Fig. \@ref(fig:innovation-record)). Participants could
review their past discoveries by clicking on an item to see its recipe.

The innovation record was also used as the means for sharing solutions among
members of a team and across sessions. For diachronic participants, the
innovation record of the previous generation participant was available to them
from the start of the experiment. They could view their predecessors'
innovation record and see how to make each one. Participants assigned to the
synchronic condition could view the innovation records of each of their
teammates at any point during the session. Once someone on a team discovered
a tool, the recipe for its creation could be viewed by all members of the team.
Isolated participants returning for multiple sessions started with the
innovation record from their previous session.

The sharing of innovation records provided the recipes for creating
the tools, but not the tools themselves. In order use these tools in further
combinations, the tools, including all of the constituent parts, first had to
be created.

(ref:innovation-record) Example innovation record.

```{r innovation-record, fig.cap="(ref:innovation-record)", eval=FALSE}
draw_image("innovation-record", "totems")
```

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the **Workshop** and selecting the "Try" button. If the
guess created an item, it could be dragged into the **Stock** panel, and used
again in later combinations. Once segments of a totem were made through
cutting, carving, and painting, they could be dragged into the **Totem** panel
for scoring.

```{r interface, fig.cap="(ref:interface)", out.width="100%"}
draw_image("gameplay", "totems")
```

## Participants {-}

Participants were undergraduate students at the University of Wisconsin-Madison
who completed the experiment in exchange for course credit. Students received
credit based on the duration of each completed session. Sessions were either 25
minutes or 50 minutes. Participants assigned to the Isolated, multi-session
condition who completed all four sessions were awarded additional credit as
an incentive to prevent dropout.

Total participant and team recruitment is presented in Table
\@ref(tab:counts-table). Data from these participants is reported as Experiments
1-3. Data from some participants is reported in multiple experiments.
Specifically, data from first and second generation diachronic participants is
included in the analyses for all experiments. Data from third and fourth
generation diachronic participants is reported in Experiment 2 and Experiment 3.
Data from two-person synchronic teams is reported in Experiment 1 and Experiment
3.

(ref:condition-counts) Total participant and team recruitment across all experiments.

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "(ref:condition-counts)")
```

# Experiment 1: Simple diachronic inheritance {-}

```{r 50min}
```

In Experiment 1, we compared the simplest form of diachronic inheritance---a
two-person diachronic team where each person is given 25 minutes---to two
alternative strategies for allocating the same 50 minutes of labor: isolated
individuals working alone and two-person synchronic teams working in parallel
(Fig. \@ref(fig:team-structures-50min)).

(ref:team-structures-50min) Team structures compared in Experiment 1.

```{r team-structures-50min, out.width="50%", fig.cap="(ref:team-structures-50min)"}
crotchet::draw_graphviz("team-structures-50min", package = "totems")
```

## Results {-}

We compared the performance of participants assigned to use the diachronic,
synchronic, and isolated strategies in three ways. First we compared performance
based on the total number of innovations that were discovered. Second we report
guessing effectiveness on a per-item basis. Were some tools easier to discover
by some team structures than others? Finally, we compared the guessing
strategies used by participants in each condition. Our analyses focus on
comparing the effectiveness of diachronic collaboration to the alternative
strategies.

### Innovations {-}

The number of innovations discovered using each of the three team strategies is
shown in Fig. \@ref(fig:innovations-50min). Our primary hypothesis concerned
the effectiveness of a diachronic strategy as compared to the alternative
strategies for solving the same problems in the same amount of time. To test
this, we fit a hierarchical linear model predicting total innovations for each
participant based on team strategy, with strategy contrast-coded such that a
diachronic strategy was the reference group. The hierarchical model included
random intercepts for each team. The results showed that a diachronic strategy
was similarly effective to either alternative strategy. Diachronic teams
discovered more innovations on average than isolated individuals, but post hoc
tests revealed that this difference was not statistically significant,
`r exp1$DG2_v_I50`. Conversely, diachronic teams discovered fewer innovations on
average than two-person synchronic teams, but again this difference was not
statistically significant, `r exp1$S2_v_DG2`.

We did find an overall benefit to teamwork in this task. To test this,
we compared two-person teams (both diachronic and synchronic) to isolated
individuals. In the hierarchical linear model, team strategy was coded with
Helmert contrasts, and random intercepts for team were included.
Teams performed reliably better on this task than individuals,
`r exp1$teamwork_stats`, without a significant residual difference between the
diachronic and synchronic strategies, `r exp1$teamwork_residual`. Post hoc tests
revealed that this effect of teamwork was driven primarily by synchronic teams
outperforming isolated individuals, `r exp1$S2_v_I50`. This result, showing an
advantage of teamwork over individual labor, suggests that maximizing individual
learning time is not always the most effective way to improve problem solving
performance.

(ref:innovations-50min) Innovations discovered in 50 labor minutes by strategy.
**A.** Rate of innovation by labor hours. Each line is the average number of
innovations discovered by participants in each condition. **B.** Final
innovations by strategy. Each point is a single participant. Bars with errors
show model predictions with ±1 standard error.

```{r innovations-50min, fig.width=5, fig.height=3, fig.cap="(ref:innovations-50min)"}
grid.arrange(
  innovation_rate_50min_plot + ggtitle("A"),
  num_innovations_50min_plot + ggtitle("B"),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

### Guesses per tool {-}

In addition to comparing the three strategies in terms of overall number of
tools, we also performed a more fine-grained analysis of guessing behaviors. We
compared the three strategies in terms of how many guesses were required to
discover each tool. To count the number of guesses that were required for each
tool, we tallied all guesses made from the moment in which a new tool was
eligible for discovery until that tool was discovered. A new tool was eligible
for discovery once all of the items required to create the new tool had been
discovered. We refer to new tools that are eligible for discovery as "adjacent",
since they are adjacent to the current inventory in the solution landscape. The
minimum number of adjacent tools for any inventory was one, indicating that
there was a single tool that must be discovered before new tools can be
created. The median number of adjacent tools was four. When there was more than
one adjacent tools, each incorrect guess was counted toward the total number of
guesses required for each of the adjacent tools.

We found that second generation diachronic participants required significantly
fewer guesses to discover each tool relative to the same tools being
discovered by participants in conditions without diachronic inheritance,
`r exp1$guesses_per_item_inheritance` (Fig. \@ref(fig:guesses-per-innovation)A).
However, this overall benefit to inheritance in terms of fewer guesses per
tool did not extend to tools that had not been discovered by the previous
generation. Comparing the number of guesses per tool for only the tools that
were newly discovered by the second generation diachronic participants did not
reveal a benefit to inheritance that extended beyond the inherited solutions,
`r exp1$guesses_per_new_item_inheritance`
(Fig. \@ref(fig:guesses-per-innovation)B).

(ref:guesses-per-innovation) Guesses per tool by inheritance condition.
Each thin gray line shows the average number of guesses for a particular
innovation. The "No inheritance" condition includes guesses from synchronic
and isolated participants, as well as first generation diachronic players.
The inheritance condition includes second generation diachronic
participants only. **A.** Guesses per tool for all tools, including those inherited
from an ancestor. **B.** Guesses per tool for new tools, not discovered by an ancestor.

```{r guesses-per-innovation, fig.width=5, fig.cap="(ref:guesses-per-innovation)"}
gridExtra::grid.arrange(
  guesses_per_item_plot + ggtitle("A"),
  guesses_per_new_item_plot +
    ggtitle("B") +
    theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    ),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

### Guessing strategy {-}

The analyses of performance reported above treat all incorrect guesses as the
same, but in reality some guesses are more informative than others.
Specifically, in this task an optimal guessing strategy is to only attempt
unique guesses. However, many people make repeat guesses, thus employing a
suboptimal strategy. The nature of these repeat guesses varies, as there are
four types of guesses that could be made, shown in Table
\@ref(tab:guess-types). We next report our analysis of whether participants in
each team structure employed different guessing strategies.

(ref:guess-types) Guess types. Each guess can be categorized into one of four
types. Redundant guesses were incorrect guesses made (at least) once before.
Repeat items were guesses that recreated previously discovered tools.
Unique guesses were incorrect guesses never made. Unique items were unique
guesses that created a new tool.

```{r guess-types, results="asis"}
recode_guess_type() %>%
  select(`Guess type` = GuessTypeLabel,
         `Unique guess` = UniqueGuess,
         `Created item` = CreatedItem,
         `Unique result` = UniqueResult) %>%
  kable(caption = "(ref:guess-types)")
```

We calculated the proportion of guesses made by participants in each condition
falling into the four guess types. Our results are shown in Fig.
\@ref(fig:guess-props). For our analysis, we focused on the proportions of
redundant and unique guesses, corresponding to an inefficient and an efficient
guessing strategy, respectively.

Looking just at inefficient redundant guesses, we found that second generation
diachronic participants generated a lower proportion of redundant guesses than
participants in the other strategies, `r exp1$prop_redundant_inheritance`.
Post hoc comparisons revealed that second generation diachronic participants
had a lower proportion of redundant guesses than first generation
diachronic participants, `r exp1$prop_redundant_dg2`, and isolated
individuals, `r exp1$prop_redundant_i50`, but not synchronic participants,
`r exp1$prop_redundant_i50`.

Looking just at efficient unique guesses, we found no difference between
second generation diachronic participants and the other strategies,
`r exp1$prop_unique_inheritance`. Pairwise comparisons suggested
a difference between second generation diachronic participants
and isolated individuals, `r exp1$prop_unique_i50`, but neither
first generation participants, `r exp1$prop_unique_dg2` nor
synchronic participants, `r exp1$prop_unique_s2`.

(ref:guess-props) Guessing strategies by team structure. See Table
\@ref(tab:guess-types) for a description of the types of guesses. DG1 and DG2 refer
to first and second generation diachronic participants. I50 refers to isolated
participants given 50 minutes. S2 refers to two-person synchronic teams.

```{r guess-props, fig.width=5, fig.cap="(ref:guess-props)"}
prop_guess_types_50min_plot
```

## Discussion {-}

In Experiment 1, we found that the simplest form of vertical transmission, in
which one person inherits from a previous generation, is comparable to
alternative strategies for solving the same set of problems in the same amount
of labor time (see Fig. \@ref(fig:types-of-time)). Two generations of
diachronic problem solvers each working for 25 minutes results in similar final
performance to an isolated individual given 50 minutes and two problem solvers
working in parallel for 25 minutes each.

Although in Experiment 1, the three strategies were similar in overall
performance, we detected the impact of inheriting from a previous generation in
the guessing efficiency and guessing strategies employed by second generation
diachronic participants. Inheriting the solutions from a previous generation
improved the efficiency of future generations of problem solvers by allowing
them to recreate those same tools with fewer guesses overall. However, this
improvement in efficiency did not extend to future problem solving, as second
generation diachronic problem solvers were not any more efficient in solving
problems that had not been discovered by their ancestor.

We also found evidence that second generation diachronic participants employed a
different guessing strategy as reflected in the distribution of guesses by type.
In this experiment, redundant guesses are unhelpful in that they will never
yield new discoveries, and yet most participants generate a significant
proportion of redundant guesses. We found that the tendency to guess redundantly
may be lessened by diachronic collaboration, as second generation diachronic
problem solvers generated fewer redundant guesses than their ancestors and
participants in an isolated strategy. However, similar proportions of redundant
guesses were generated by participants in the synchronic strategy, suggesting
that diachronic inheritance is not the only strategy for reducing redundancy in
guessing.

The results of Experiment 1 also support and extend the findings reported by
@Derex:2015cbb that synchronic collaboration can result in an improvement in
problem solving performance relative to isolated individuals working alone. Here
we demonstrate that this advantage holds even when labor hours are controlled.

The results of Experiment 1 suggest that, rather than engaging in a problem
solving task as an isolated individual, it might in some cases be more efficient
to pass the solutions on to a future generation via diachronic collaboration.
In Experiment 2, we test this idea further by comparing the benefit of
inheriting from someone else, as in the case of diachronic teams, to the
benefit of inheriting from oneself, over multiple sessions.

# Experiment 2: Diachronic versus individual inheritance {-}

```{r SelfOther}
```

In Experiment 2, we compared the effectiveness of four generations of
diachronic inheritance---inheriting from someone else---to the effectiveness of
isolated individuals given four separate sessions (Fig.
\@ref(fig:team-structures-selfother)). Participants in the isolated condition
came back to the experiment for four separate sessions, each 25 minutes long.
At each session after the first, these participants inherited the recipes for
the tools they had created at the end of the previous session. Thus, in this
experiment, both diachronic and isolated individuals started each session
recreating inherited tools before extending these to discover
new tools.

Based on the results of Experiment 1 and previous work using this paradigm
[@Derex:2015cbb], we expected individual problem solvers to become fixated on
previous incorrect solutions. The purpose of Experiment 2 was to see if
diachronic collaboration could be used to remediate these costs of fixation.

(ref:team-structures-selfother) Problem solving strategies compared in Experiment 2.

```{r team-structures-selfother, out.width="70%", fig.cap="(ref:team-structures-selfother)"}
crotchet::draw_graphviz("team-structures-selfother", package = "totems")
```

## Results {-}

We start by reporting differences in the total number of tools discovered using
either a diachronic or isolated strategy. To foreshadow our results, we found
that isolated individuals given four sessions were able to accumulate more tools
than diachronic teams given the same total amount of time. The remaining analyses
test hypotheses about why an isolated strategy was more effective than a
diachronic strategy.

### Innovations {-}

The number of innovations discovered in each diachronic generation and isolated
session are shown in Fig. \@ref(fig:innovations-by-generation-selfother). After
a single 25 minute session, there was no difference between diachronic and
isolated participants, `r exp2$gen1_innovations`. This result was expected,
as after a single session, there was no difference between participants
randomly assigned to either strategy. For diachronic participants, each
new generation was able to discover `r exp2$tools_per_diachronic_gen` more
tools than they inherited, `r exp2$tools_per_diachronic_gen_stats`. However,
isolated individuals discovered `r exp2$tools_per_isolated_gen` more tools than
they inherited, `r exp2$tools_per_isolated_gen_stats`. That is, an isolated
strategy proved more effective than a diachronic strategy over generations,
`r exp2$innovations_by_inheritance_slope`. After four generations of diachronic
collaboration compared to four sessions of isolated labor, an isolated strategy
was more effective, `r exp2$gen4_innovations`.

(ref:innovations-by-generation-selfother) Innovations discovered by diachronic
and isolated strategies. Each thin line is a diachronic team (**left**) or an
isolated individual (**right**). Thick lines with error bars show model
predictions with ±1 standard error.

```{r innovations-by-generation-selfother, fig.width=5, fig.cap="(ref:innovations-by-generation-selfother)"}
innovations_by_generation_plot
```

### Guessing strategy {-}

In order to determine why an isolated strategy was more effective than a
diachronic strategy, we asked whether isolated individuals employed a more
efficient guessing strategy than diachronic participants. As in Experiment 1, we
measured both the proportion of redundant guesses and unique guesses made by
participants in either strategy, corresponding to an inefficient and efficient
guessing strategy, respectively.

Although participants in the isolated strategy discovered more tools after
four sessions than diachronic teams after four generations, they did not engage
in a more efficient guessing strategy. Unlike in Experiment 1, diachronic
participants did not guess a lower proportion of redundant guesses compared to
isolated individuals, `r exp2$prop_redundant_strategy`. For participants in both
strategies, the proportion of redundant guesses decreased over generations and
sessions, `r exp2$prop_redundant_gen`. However, this decrease was not affected
by strategy, `r exp2$prop_redundant_inter`.

For the proportion of unique guesses, there were no differences between
participants in either strategy, `r exp2$prop_unique_strategy`, or
over generations `r exp2$prop_unique_gen`. The change in proportion
of unique guesses over generations was not affected by strategy,
`r exp2$prop_unique_inter`.

(ref:guessing-strategy-by-generation) Proportion of guesses by participants in
diachronic and isolated conditions over generations.

```{r guessing-strategy-by-generation, fig.width=5, fig.cap="(ref:guessing-strategy-by-generation)"}
prop_guess_types_selfother_plot
```

If participants in the isolated strategy are not guessing more efficiently or
less redundantly than participants in the diachronic strategy, why then are they
able to accumulate more tools than participants in the diachronic condition? To
answer this question, we investigated the time spent recreating the inherited
tools versus time spent inventing new tools.

### Learning times {-}

Participants took on average `r exp2$mean_learning_time` minutes of the 25
minute session (`r exp2$prop_learning_time`%) to recreate the inherited
tools---a portion of the experiment we refer to as the learning period. We
found that isolated individuals returning for multiple sessions took significantly
less time to recreate the tools they had invented at the end of their previous
session as compared to diachronic participants who were playing
the Totems game for the first time, `r exp2$learning_times_inter` (Fig.
\@ref(fig:learning-times)).

This result suggests that what made an isolated strategy successful in
Experiment 2 is not that isolated individuals were better at guessing new tools,
but that they were able to more quickly recreate the tools they inherited, and
thus had more time to dedicate to discovering new tools. Diachronic
participants, who were exposed to the game for the first time, took longer to
recreate the inherited items, and had less time to discover new tools.

(ref:learning-times) Learning times by strategy. Learning time refers to
the number of minutes it took to recreate the inherited items. For
diachronic participants (**left**), learning time scaled linearly with
generations, but for isolated participants returning for multiple sessions
(**right**), learning times were stable across sessions.

```{r learning-times, fig.width=5, fig.cap="(ref:learning-times)"}
learning_times_plot
```

### Guesses for new discoveries {-}

Both diachronic and isolated participants first had to recreate the tools
discovered by the previous generation participant or in the previous session.
Given the tendency for individuals to get fixated in this task, we asked
whether a diachronic or isolated strategy was more effective at discovering new
tools. To test this, we counted the number of guesses it took each participant
to make the first new tool after recreating all inherited tools.

We found that diachronic participants were able to discover the first new tool
with fewer guesses than isolated participants (Fig.
\@ref(fig:cost-of-first-innovation)). On average, participants in both diachronic
and isolated strategies took `r exp2$first_discovery_mean` guesses to discover
a new item not inherited from a previous participant or previous session. Diachronic
participants discovered a novel item in `r exp2$first_discovery_beta` fewer guesses
than isolated individuals, `r exp2$first_discovery_stats`. This difference got larger
over generations, `r exp2$first_discovery_by_gen_stats`, indicating that an isolated
strategy was becoming more fixated on previous incorrect solutions.

(ref:cost-of-first-innovation) Number of guesses required to discover the first
new tool each session.

```{r cost-of-first-innovation, fig.width=5, fig.cap="(ref:cost-of-first-innovation)"}
first_discovery_by_generation_plot
```

## Discussion {-}

In Experiment 2, we found that isolated individuals given four sessions
were able to outperform diachronic individuals, each given only 25 minutes.
However, an isolated strategy did not result in more efficient problem
solving than a diachronic strategy. Isolated participants did not guess
less redundantly, or generate a higher proportion of unique guesses,
than diachronic participants. Isolated participants, given multiple sessions,
were able to more quickly recreate the tools they discovered in the previous
session, thus giving them more time to discover new tools. In contrast,
for diachronic participants, recreating the tools inherited from their
ancestor took longer and longer over generations, leaving less time for
new tool discovery.

We did obtain evidence of one way in which a diachronic strategy can
promote a more efficient problem solving. We found that isolated
individuals were more likely to be fixated on incorrect solutions, causing
them to guess inefficiently. These fixation effects extended over multiple
sessions, causing the performance of some isolated individuals to stagnate.
Diachronic participants, however, were less impacted by fixation, and were
more easily able to move beyond the solutions they inherited to discover
new solutions. These results reveal an important benefit to diachronic
collaboration and inheritance.

In Experiment 3, we investigated whether adding more participants to
a synchronic strategy increases problem solving performance to the same
extend as adding generations to a diachronic team.

# Experiment 3: Scalability of diachronic and synchronic collaboration {-}

```{r TeamSize}
```

In Experiment 3 we measured the scalability of diachronic and synchronic collaboration.
Does adding more people to the problem solving team improve the performance of
each strategy equally? We hypothesized that diachronic collaboration would be
more scalable than synchronic collaboration. To test this, we compared
diachronic and synchronic collaboration given 50 minutes to the same strategies
given 100 minutes.

## Results {-}

As reported in Experiment 1, two-person synchronic teams discovered on average
more innovations than two-generation diachronic teams, although this difference
was not statistically significant, `r exp3$DvS_2`. Comparing four-person
synchronic teams to four-generation diachronic teams revealed a different
pattern. Adding two people to a synchronic team did not result in a significant
improvement to team performance, `r exp3$S_2v4`. In contrast, adding
two people to a diachronic team did result in an improvement to team
performance, `r exp3$D_2v4`. As a result, there was a reliable interaction
between team strategy and team size, `r exp3$DvS_2v4`. These results suggest that a diachronic
strategy is more scalable than a synchronic strategy, meaning that adding more people to a
diachronic team may result in a bigger improvement in final performance compared to adding
people to a synchronic team.

(ref:max-innovations-by-teamsize) Innovations by team size. Performance of
diachronic teams after two or four generations is compared to performance of
synchronic teams with two or four individuals.

```{r max-innovations-by-teamsize, fig.width=5, fig.cap="(ref:max-innovations-by-teamsize)"}
max_innovations_by_teamsize_plot
```

## Discussion {-}

In Experiment 3, we demonstrated that diachronic and synchronic collaboration differ
in their ability to scale up to larger teams. Diachronic collaboration can add
more generations that each correspond to an increase in team performance, but the utility
of adding problem solvers to a synchronic strategy is smaller. Specifically we found
that the reason might be related to synchronic individuals each making the same
incorrect guesses in tandem, meaning that each individual is unable to only guess
unique combinations for the team.

# Discussion {-}
