---
title: "The impact of team structure on problem solving performance"
author:
  - { name: Pierce Edmiston, affiliation_id: 1 }
  - { name: Maxime Derex, affiliation_id: 2 }
  - { name: Gary Lupyan, affiliation_id: 1 }
affiliation:
  - id: 1
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - id: 2
    department: Biosciences Department
    university: University of Exeter
email: pedmiston\@wisc.edu
address: 1202 W. Johnson St., Department of Psychology, University of Wisconsin, Madison, WI 53703.
bibliography: "references.bib"
csl: "templates/plos-one.csl"
output:
  bookdown::pdf_document2:
    template: templates/plos-one.tmpl
    keep_tex: yes
    number_sections: no
---
```{r team-structures-config, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",

  # PLOS One figure requirements.
  # Width: 789 – 2250 pixels (at 300 dpi). Height maximum: 2625 pixels (at 300 dpi).
  fig.width = 3,
  fig.height = 3,
  dpi = 300
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "team-structures.R"))
```

```{r Setup}
```

# Introduction {-}

```{r Intro}
```

(ref:types-of-time) Team structures controlling for labor time. **Left.** Team structures. Each node is a person. **Middle.** Relationship between labor time and calendar time. Synchronic teams spend labor time more quickly than the other strategies. **Right.** Learning times by strategy, showing that an isolated strategy provides the most time for individual learning.

```{r types-of-time, fig.width=5, fig.height=2, fig.cap="(ref:types-of-time)"}
gridExtra::grid.arrange(
  crotchet::read_graphviz("team-structures", package = "totems"),
  gg_time,
  gg_person,
  nrow = 1,
  widths = c(0.25, 0.375, 0.375)
)
```

# Methods {-}

```{r Methods}
```

## Team structures {-}

Participants were randomly assigned to complete the same problem solving task in
different team structure conditions (Fig. \@ref(fig:team-structures-all)).
Participants in the diachronic condition worked on the problem solving task one
at a time. At the end of a diachronic participant's session, the solutions they
discovered were passed on to be inherited by a single participant assigned
to the next generation of the same team.

Participants assigned to the synchronic team structure worked independently and
in parallel in teams of two or four. Their labor was coordinated such that
solutions discovered by one team member were immediately shared with the other
members of the team.

Isolated participants completed the problem solving task as individuals, only
benefiting from the solutions they could discover on their own. Isolated participants
completed the experiment in a single 50 minute session or, in a different condition,
came back for four 25 minute sessions.

(ref:team-structures-all) Team structures. Diachronic participants worked one
at a time, each inheriting the solutions discovered by the previous generation,
labeled G1-G4. Isolated participants were recruited for a single 50 minute
session, or returned for four 25 minute sessions, labeled S1-S4. Synchronic
participants were assigned to teams of two or four individuals that all worked
at the same time.

```{r team-structures-all, out.width="80%", fig.cap='(ref:team-structures-all)'}
crotchet::draw_graphviz("team-structures-all", package = "totems")
```

## Problem solving task {-}

Participants played the "Totems game" adapted from @Derex:2015cbb. Their task
was to discover tools that could be used with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). More advanced tools
produce larger and more intricate totems, resulting in higher performance
scores.

Participants could combine up to four items at a time (with replacement),
meaning the initial six items could form a total of
`r methods$n_unique_combinations_6` combinations. Of all possible combinations,
very few resulted in new items. For example, of all the combinations that could
be formed from the initial items, only three
(`r methods$n_unique_combinations_6_pct`%) yielded new tools.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6
items were available to problem solvers at the start of the game. New items
could be produced through the combination of different items (more than one
arrow points to the item) or through the refinement of a single item (a single
arrow points to the item). The axe is required to construct the first totem
pole.

```{r landscape-sample, fig.width=5, fig.cap="(ref:landscape-sample)"}
draw_image("landscape-sample.gv", package = "totems")
```

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record (Fig. \@ref(fig:innovation-record)). Participants could
review their past discoveries by clicking on an item to see its recipe.

The innovation record was also used as the means for sharing solutions among
members of a team and across sessions. For diachronic participants, the
innovation record of the previous generation participant was available to them
from the start of the experiment. They could view their predecessors'
innovation record and see how to make each one. Participants assigned to the
synchronic condition could view the innovation records of each of their
teammates at any point in the session. Once someone on a team discovered
a tool, the recipe for its creation could be viewed by all members of the team.
Isolated participants returning for multiple sessions started with the
innovation record from their previous session.

Note that the sharing of innovation records provided the recipes for creating
the tools, but not the tools themselves. In order use these tools in further
combinations, the tools, including all of the constituent parts, first had to
be created.

(ref:innovation-record) Example innovation record.

```{r innovation-record, fig.cap="(ref:innovation-record)", eval=FALSE}
draw_image("innovation-record", "totems")
```

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the **Workshop** and selecting the "Try" button. If the
guess created an item, it could be dragged into the **Stock** panel, and used
again in later combinations. Once segments of a totem were made through
cutting, carving, and painting, they could be dragged into the **Totem** panel
for scoring.

```{r interface, fig.cap="(ref:interface)", out.width="100%"}
draw_image("gameplay", "totems")
```
## Participants {-}

Participants were undergraduate students at the University of Wisconsin-Madison
who completed the experiment in exchange for course credit. Students received
credit based on the duration of each completed session. Sessions were either 25
minutes or 50 minutes. Participants assigned to the Isolated, multi-session
condition who completed all four sessions were awarded additional credit as
an incentive to prevent dropout.

(ref:condition-counts) Total participant and team recruitment.

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "(ref:condition-counts)")
```

## Data {-}

The data collected and reported here is available in online repositories
and in an R package.

## Statistical analyses {-}

We used hierarchical linear regression models for our analyses.

# Experiment 1: Simple diachronic inheritance {-}

```{r 50min}
```

In Experiment 1, we compared the simplest form of diachronic inheritance---a
two-person diachronic team where each person is given 25 minutes---to two
alternative strategies for allocating the same 50 minutes of labor: isolated
individuals working alone and two-person synchronic teams working in parallel
(Fig. \@ref(fig:team-structures-50min)).

(ref:team-structures-50min) Team structures compared in Experiment 1.

```{r team-structures-50min, out.width="50%", fig.cap="(ref:team-structures-50min)"}
crotchet::draw_graphviz("team-structures-50min", package = "totems")
```

## Results {-}

We compared the performance of diachronic, synchronic, and isolated
participants in three ways. First we compared performance based on the number
of innovations discovered. Next we investigated guessing effectiveness on
a per-item basis. Were some items easier to discover in some team structures
than others? Finally, we compared the guessing strategy of participants in each
condition by analyzing the proportion of redundant versus unique
guesses made by each participant. In our analyses, we focus on comparing the
effectiveness of diachronic inheritance to alternative strategies for solving
the same problems.

### Innovations {-}

The number of innovations discovered using each of the three team strategies is
shown in Fig. \@ref(fig:innovations-50min). Our primary hypothesis concerned
the effectiveness of a diachronic strategy as compared to the alternative
strategies for solving the same problems in the same amount of time, but
without diachronic inheritance. The results show that when controlling for
labor hours, a diachronic strategy is as effective as either alternative
strategy. Diachronic teams discovered more innovations on average than isolated
individuals, but post hoc tests revealed that this difference was not
statistically significant, `r exp1$DG2_v_I50`. Conversely, diachronic teams
discovered fewer innovations on average than two-person synchronic teams, but
again this difference was not statistically significant, `r exp1$S2_v_DG2`.

We did find an overall benefit to teamwork in this task. To test this,
we compared two-person teams (both diachronic and synchronic) to isolated
individuals. Teams performed reliably better on this task than individuals,
`r exp1$teamwork_stats`, without a significant residual difference between the
diachronic and synchronic strategies, `r exp1$teamwork_residual`. Post hoc tests
revealed that this effect of teamwork was driven primarily by synchronic teams
outperforming isolated individuals, `r exp1$S2_v_I50`. This result, showing an
advantage of teamwork over individual labor, suggests that maximizing individual
learning time is not always the most effective way to improve problem solving
performance.

(ref:innovations-50min) Innovations discovered in 50 labor minutes by strategy.
**A.** Rate of innovation by labor hours. Each line is the average number of
innovations discovered by participants in each condition. **B.** Final
innovations by strategy. Each point is a single participant. Bars with errors
show model predictions with ±1 standard error.

```{r innovations-50min, fig.width=5, fig.height=3, fig.cap="(ref:innovations-50min)"}
grid.arrange(
  innovation_rate_50min_plot + ggtitle("A"),
  num_innovations_50min_plot + ggtitle("B"),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

### Guesses per tool {-}

In addition to comparing the three strategies in terms of overall performance,
we also performed a more fine-grained analysis of guessing behaviors. First we
compared the three strategies in terms of how many guesses were required to
discover each tool. To count the number of guesses that were required
for each tool, we tallied all guesses made from the moment in which a new
tool was eligible for discovery until that tool was discovered. A new tool was
eligible for discovery once all of the items required to create the new tool
had been discovered. We refer to new tools that are eligible for discovery
as adjacent, since they are adjacent to the current inventory in the
solution landscape. The minimum number of adjacent tools for any inventory
was one, indicating that there was a single tool that must be discovered
before new tools can be created. The median number of adjacent tools was four.
When there was more than one adjacent tools, each incorrect guess was counted
toward the total number of guesses required for each of the adjacent tools.

We found that diachronic inheritance reduced the average number of guesses
required to discover each tool relative to the same tools being discovered by
participants in either the isolated or synchronic conditions,
`r exp1$guesses_per_item_inheritance` (Fig.
\@ref(fig:guesses-per-innovation)A). However, we found that this overall benefit
to inheritance in terms of fewer guesses per tool did not extend to tools that
had not been discovered by the previous generation. Comparing the number of
guesses per tool only for tools that were newly discovered by the second
generation diachronic problem solvers did not reveal a benefit to inheritance
that extended beyond the inherited solutions,
`r exp1$guesses_per_new_item_inheritance` (Fig. \@ref(fig:guesses-per-innovation)B).

(ref:guesses-per-innovation) Guesses per tool by inheritance condition.
Each thin gray line shows the average number of guesses for a particular
innovation. The no inheritance condition includes guesses from synchronic
and isolated participants, as well as first generation diachronic players.
The inheritance condition includes second generation diachronic
participants only. **A.** Guesses per tool for all tools. **B.** Guesses per
tool for new tools, not discovered by an ancestor.

```{r guesses-per-innovation, fig.width=5, fig.cap="(ref:guesses-per-innovation)"}
gridExtra::grid.arrange(
  guesses_per_item_plot + ggtitle("A"),
  guesses_per_new_item_plot +
    ggtitle("B") +
    theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    ),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

### Guessing strategy {-}

The analyses of guesses reported above treat all incorrect guesses as the same,
but in reality some guesses are more informative than others. Specifically, in
this task an optimal guessing strategy is to only attempt unique guesses.
However, many people make repeat guesses, thus employing a suboptimal strategy.
The nature of these repeat guesses varies, as there are four types of guesses
that could be made, shown in Table \@ref(tab:guess-types). We next report our
analysis of whether participants in each team structure employed different
guessing strategies.

(ref:guess-types) Guess types. Each guess can be categorized into one of four
types. Redundant guesses were incorrect guesses made (at least) once before.
Repeat items were redundant guesses that recreated already discovered tools.
Unique guesses were incorrect guesses never made. Unique items were unique
guesses that created a new tool.

```{r guess-types, results="asis"}
recode_guess_type() %>%
  select(`Guess type` = GuessTypeLabel,
         `Unique guess` = UniqueGuess,
         `Created item` = CreatedItem,
         `Unique result` = UniqueResult) %>%
  kable(caption = "(ref:guess-types)")
```

We calculated the proportion of guesses made by participants in each condition
falling into the four guess types. Our results are shown in Fig.
\@ref(fig:guess-props).

For our analysis, we focused on the proportions of
redundant and unique guesses. Looking just at redundant guesses, we found
that inheritance caused participants to guess fewer redundant guesses
than the other strategies, `r exp1$prop_redundant_inheritance`.
Post hoc comparisons revealed that second generation diachronic
participants had a lower percentage of redundant guesses than first
generation participants, `r exp1$prop_redundant_dg2`, and isolated
individuals, `r exp1$prop_redundant_i50`, but not synchronic participants,
`r exp1$prop_redundant_i50`.

Looking at the proportion of unique guesses, we found no overall effect of
inheritance, `r exp1$prop_unique_inheritance`. Pairwise comparisons revealed
a significant difference between second generation diachronic participants
and isolated individuals, `r exp1$prop_unique_i50`, but neither
first generation participants, `r exp1$prop_unique_dg2` nor
synchronic participants, `r exp1$prop_unique_s2`.

(ref:guess-props) Guessing strategies by team structure. DG1 and DG2 refer to first
and second generation diachronic participants. I50 refers to isolated participants
given 50 minutes. S2 refers to two-person synchronic teams.

```{r guess-props, fig.width=5, fig.cap="(ref:guess-props)"}
prop_guess_types_50min_plot
```

## Discussion {-}

In Experiment 1, we found that the simplest form of inheritance, where one
person inherits from a previous generation, is comparable to alternative
strategies for solving the same set of problems. Two generations of diachronic
problem solvers each working for 25 minutes results in similar final
performance to an isolated individual given 50 minutes and two problem solvers
working in parallel for 25 minutes each.

We detected the effects of inheriting from a previous generation in the guessing efficiency
and guessing strategies employed by diachronic problem solvers. Inheriting the solutions
from a previous generation improved the efficiency of future generations of problem solvers
by allowing them to recreate those same tools with fewer guesses overall. However, this
improvement in efficiency did not extend to future problem solving, as second generation
diachronic problem solvers were not any more efficient in solving problems that had not
been discovered by their ancestor.

Second generation diachronic players did seem to employ a different guessing strategy
as reflected in the distribution of guesses by type. In the Totems game, redundant guesses
are unhelpful in that they will never yield new discoveries, and yet most participants
generate a significant proportion of redundant guesses. We found that the tendency to guess
redundantly may be lessened by diachronic collaboration, as second generation
diachronic problem solvers generated fewer redundant guesses than their
ancestors and participants in an isolated strategy. However, similar proportions
of redundant guesses were generated by participants in the synchronic strategy,
suggesting that diachronic inheritance is not the only effective strategy
for reducing redundancy.

The results of Experiment 1 also support and extend the original findings reported by
@Derex:2015cbb that synchronic collaboration can result in an improvement in
problem solving performance relative to isolated individuals working alone. Here
we demonstrate that this advantage holds even when labor hours are controlled.

In Experiment 1, we found evidence that isolated individuals might improve
problem solving performance by passing their solutions on to a future generation
rather than completing the same problems individually. In Experiment 2, we
test this idea further by comparing the benefit of inheriting from someone else,
as in the case of diachronic teams, to the benefit of inheriting from oneself,
over multiple sessions.

# Experiment 2: Inheriting from self or other {-}

```{r SelfOther}
```

In Experiment 2, we compared the effectiveness of four generations of
diachronic inheritance---inheriting from someone else--- to the effectiveness
of isolated individuals given four independent sessions. (Fig.
\@ref(fig:team-structures-selfother)). Participants in the isolated condition
came back to the experiment for four separate sessions, each 25 minutes long.
At each session after the first, these participants inherited the recipes for
the tools they had created at the end of the previous session.

(ref:team-structures-selfother) Problem solving strategies compared in Experiment 2.

```{r team-structures-selfother, out.width="70%", fig.cap="(ref:team-structures-selfother)"}
crotchet::draw_graphviz("team-structures-selfother", package = "totems")
```

## Results {-}

We compared the performance of four generations of diachronic problem solving
to four sessions of isolated problem solving in two main ways. First,
we investigated the different in overall performance. Second, we investigated
the cost of fixation in both strategies.

### Innovations {-}

```{r innovations-by-generation-selfother, fig.width=5}
innovations_by_generation_plot
```

### Fixation {-}

```{r cost-of-first-innovation, fig.width=5}
first_discovery_by_generation_plot
```

## Discussion {-}

In Experiment 2, we found that isolated individuals given four sessions
were able to outperform diachronic individuals, each given only 25 minutes.
However, isolated individuals were not always more efficient problem
solvers than diachronic individuals. Specifically, we found that isolated
individuals were more likely to be fixated on incorrect solutions, causing
them to guess inefficiently. These fixation effects extended over multiple
sessions, causing the performance of some isolated individuals to stagnate.
Diachronic participants, however, were less impacted by fixation, and were
more easily able to move beyond the solutions they inherited to discover
new solutions. These results reveal an important benefit to diachronic
collaboration and inheritance.

In Experiment 2, we compared teams given 100 labor minutes. In Experiment 3,
we report the performance of four person synchronic teams given the same
number of labor minutes.

# Experiment 3: Scalability of diachronic and synchronic collaboration {-}

```{r TeamSize}
```

In Experiment 3 we measured the scalability of diachronic and synchronic collaboration.
Does adding more people to the problem solving team improve the performance of
each strategy equally? We hypothesized that diachronic collaboration would be
more scalable than synchronic collaboration. To test this, we compared
diachronic and synchronic collaboration given 50 minutes to the same strategies
given 100 minutes.

## Results {-}

```{r max-innovations-by-teamsize, fig.width=5}
max_innovations_by_teamsize_plot
```

```{r bots-teamsize, fig.width=5}
bots_team_size_plot
```

## Discussion {-}

# General Discussion {-}
