---
title: "The impact of team structure on problem solving performance"
author:
  - { name: Pierce Edmiston, affiliation_id: 1 }
  - { name: Maxime Derex, affiliation_id: 2 }
  - { name: Gary Lupyan, affiliation_id: 1 }
affiliation:
  - id: 1
    department: Department of Psychology
    university: University of Wisconsin-Madison
  - id: 2
    department: Biosciences Department
    university: University of Exeter
email: pedmiston\@wisc.edu
address: 1202 W. Johnson St., Department of Psychology, University of Wisconsin, Madison, WI 53703.
bibliography: "references.bib"
csl: "templates/plos-one.csl"
output:
  bookdown::pdf_document2:
    template: templates/plos-one.tmpl
    keep_tex: yes
    number_sections: no
---
```{r team-structures-config, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",

  # PLOS One figure requirements.
  # Width: 789 – 2250 pixels (at 300 dpi). Height maximum: 2625 pixels (at 300 dpi).
  fig.width = 3,
  fig.height = 3,
  dpi = 300
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "team-structures.R"))
```

```{r Setup}
```

# Introduction {-}

```{r Intro}
```

Cultural evolution is a complex process involving both vertical and horizontal
transmission [@Boyd:1985cu]. Vertical transmission is the process of passing
cultural products from parent to child, whereas horizontal transmission
involves passing cultural products among peers in the same generation. Some
cultural products are more likely to be passed vertically than horizontally.
For example, religious beliefs are more likely to be learned from parents than
from peers [@Newcomb:1937vy]. But far more cultural products are believed to
be shared among unrelated individuals via horizontal transmission. As a result,
horizontal transmission has been argued to be the default process in cultural
evolution [@Kroeber:1948ant].

The distinction between vertical and horizontal transmission is motivated by
analogous processes in genetic evolution. In genetic evolution, vertical
transmission occurs via reproduction, whereas horizontal transmission is the
transfer of genetic material by any means _other_ than reproduction
[@Soucy:2015gp]. Thus, the distinction between vertical and horizontal
transmission of genetic material is motivated by an underlying difference in
mechanism, as the cellular mechanisms involved in reproduction (e.g., meiosis)
are different from those involved in horizontal transfer (e.g., conjugation).

In contrast, there is no such difference in mechanism between the vertical and
horizontal transmission of cultural products. All processes of cultural
transmission involve social learning, in which one person learns a behavior
from the observation or interaction with someone else. The cognitive mechanisms
that allow for social learning are functionally the same regardless of who is
being copied. If the same social learning abilities are involved in both
vertical and horizontal transmission, is there any difference in the
consequences of either vertical or horizontal transmission?

Whether individuals learn from parents or peers does have consequences on the
distribution of traits in a population. For example, @Bettinger:1999vv surveyed
variation in the design of projectile points found in the Great Basin region,
and found two types of points. In eastern California, design attributes of the
recovered projectile points were poorly coordinated with one another,
indicating greater variability in design in this region. They argued that this
variability among members of the same society indicates a culture of vertical
transmission, in which each family may have invented their own projectile point
designs and passed them down along family lines. In contrast, in central Nevada
over the same period, projectile point designs were highly correlated with one
another, indicating a process of biased horizontal transmission. Bettinger and
Eerkens argued that in these societies, new generations of hunters may have
copied the most prestige hunters in the group rather than their parents, which
had the result of more similarity among projectile point designs. This
hypothesis has been supported by experimental and computational simulations
showing that fully connected networks of problem solvers result in greater
convergence and smaller variation in design than independently evolving
lineages [@Mesoudi:2008vi; @Mesoudi:2008we].

When considering the consequences of vertical and horizontal transmission,
horizontal transmission is considered advantageous because it provides a level
of redundancy that combats the potential for information loss due to
transmission error. If every instance of cultural transmission has some
probability of transmission error, the ability to share innovations widely
within a given generation increases the likelihood that it will survive to be
used and improved by the next generation. As a case study of this phenomena,
@Henrich:2004ht modeled the apparent loss of technological complexity among
Tasmanian islanders over an 8,000 year period as the result of a decreasing
population size. With fewer replicators within each generation to learn complex
skills, eventually transmission errors resulted in the loss of some
technologies. This effect has been demonstrated in a behavioral experiment
[@Derex:2013kk]. To overcome the potential for information loss, human cultures
must increase in size, and excel at propagating innovations horizontally to as
many people in each generation as possible.

Another way in which horizontal transmission is believed to have different
consequences than purely vertical transmission is in providing multiple
behavioral models to learn from. A common assumption in models of the cultural
evolutionary process is that individuals learn behaviors from just one model.
However, computational models by @Enquist:2010gw have shown that when
individuals learn from just one parent (as in vertical transmission),
cumulative culture is unlikely to occur. The reason is that, along chains of
vertical transmission, any error in transmission is necessarily propagated to
future generations, increasing the likelihood of maladaptive traits becoming
established in populations. In contrast, when individuals can learn from
multiple cultural models and adaptively filter among the observed behaviors to
determine the best person to copy, then the cumulative cultural process can
proceed as expected [@Rendell:2011gk].

Horizontal transmission is beneficial in providing redundancy among replicators
and opportunities for adaptive filtering, but are there any advantages to
vertical transmission? To answer this question, we compared the effectiveness
of vertical and horizontal transmission in a problem solving task.

In this research, we distinguish between vertical and horizontal transmission
along two dimensions of time. Vertical transmission is defined as coordination
over generations, such that innovations discovered by one generation are passed
on to be used and improved by the next generation. Horizontal transmission is
defined as coordination in parallel and in the same generation. We refer to
labor that is coordinated via vertical transmission as "diachronic
collaboration" and labor that is coordinated via horizontal transmission as
"synchronic collaboration". We compared the effectiveness of diachronic
and synchronic collaboration to a problem solving strategy involving
isolated individuals working alone.

In our comparison of diachronic, synchronic, and isolated strategies,
we controlled for **labor time**. Labor time can be contrasted with
calendar time and learning time (Fig. \@ref(fig:types-of-time)). We
controlled for labor time so that our results can be interpreted in
terms of the effect each problem solving strategy has on individual
problem solving ability. Without controlling for labor time, any differences
between the three strategies are more difficult to interpret. For example,
if only calendar hours are controlled, an isolated individual given 30 minutes
could be compared to a synchronic team given 30 minutes _regardless of the
size of the synchronic team_. This fact makes it difficult to understand 
the effect of teamwork on individual performance when only controlling for
calendar time [@Propp:2003kj; @Tindale;1992ub].

(ref:types-of-time) Team structures controlling for labor time. **Left.** Team
structures. Each node is a person. **Middle.** Relationship between labor time
and calendar time. Synchronic teams spend labor time more quickly than the
other strategies. **Right.** Learning times by strategy, showing that an
isolated strategy provides the most time for individual learning.

```{r types-of-time, fig.width=5, fig.height=2, fig.cap="(ref:types-of-time)"}
gridExtra::grid.arrange(
  crotchet::read_graphviz("team-structures", package = "totems"),
  gg_time,
  gg_person,
  nrow = 1,
  widths = c(0.25, 0.375, 0.375)
)
```

# Methods {-}

```{r Methods}
```

## Team structures {-}

Participants were randomly assigned to complete the same problem solving task in
different team structure conditions (Fig. \@ref(fig:team-structures-all)).
Participants in the diachronic condition worked on the problem solving task one
at a time. At the end of a diachronic participant's session, the solutions they
discovered were passed on to be inherited by a single participant assigned
to the next generation of the same team.

Participants assigned to the synchronic team structure worked independently and
in parallel in teams of two or four. Their labor was coordinated such that
solutions discovered by one team member were immediately shared with the other
members of the team.

Isolated participants completed the problem solving task as individuals, only
benefiting from the solutions they could discover on their own. Isolated participants
completed the experiment in a single 50 minute session or, in a different condition,
came back for four 25 minute sessions.

(ref:team-structures-all) Team structures. Diachronic participants worked one
at a time, each inheriting the solutions discovered by the previous generation,
labeled G1-G4. Isolated participants were recruited for a single 50 minute
session, or returned for four 25 minute sessions, labeled S1-S4. Synchronic
participants were assigned to teams of two or four individuals that all worked
at the same time.

```{r team-structures-all, out.width="80%", fig.cap='(ref:team-structures-all)'}
crotchet::draw_graphviz("team-structures-all", package = "totems")
```

## Problem solving task {-}

Participants played the "Totems game" adapted from @Derex:2015cbb. Their task
was to discover tools that could be used with the ultimate goal of creating "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered tools: a refined stick
used as a handle, a sharpened rock for the blade, and a string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). More advanced tools
produce larger and more intricate totems, resulting in higher performance
scores.

Participants could combine up to four items at a time (with replacement),
meaning the initial six items could form a total of
`r methods$n_unique_combinations_6` combinations. Of all possible combinations,
very few resulted in new items. For example, of all the combinations that could
be formed from the initial items, only three
(`r methods$n_unique_combinations_6_pct`%) yielded new tools.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6
items were available to problem solvers at the start of the game. New items
could be produced through the combination of different items (more than one
arrow points to the item) or through the refinement of a single item (a single
arrow points to the item). The axe is required to construct the first totem
pole.

```{r landscape-sample, fig.width=5, fig.cap="(ref:landscape-sample)"}
draw_image("landscape-sample.gv", package = "totems")
```

Once a tool was discovered, the recipe for its creation---a list of the items
that had to be combined in order to create the tool---was recorded in an
innovation record (Fig. \@ref(fig:innovation-record)). Participants could
review their past discoveries by clicking on an item to see its recipe.

The innovation record was also used as the means for sharing solutions among
members of a team and across sessions. For diachronic participants, the
innovation record of the previous generation participant was available to them
from the start of the experiment. They could view their predecessors'
innovation record and see how to make each one. Participants assigned to the
synchronic condition could view the innovation records of each of their
teammates at any point in the session. Once someone on a team discovered
a tool, the recipe for its creation could be viewed by all members of the team.
Isolated participants returning for multiple sessions started with the
innovation record from their previous session.

The sharing of innovation records provided the recipes for creating
the tools, but not the tools themselves. In order use these tools in further
combinations, the tools, including all of the constituent parts, first had to
be created.

(ref:innovation-record) Example innovation record.

```{r innovation-record, fig.cap="(ref:innovation-record)", eval=FALSE}
draw_image("innovation-record", "totems")
```

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the **Workshop** and selecting the "Try" button. If the
guess created an item, it could be dragged into the **Stock** panel, and used
again in later combinations. Once segments of a totem were made through
cutting, carving, and painting, they could be dragged into the **Totem** panel
for scoring.

```{r interface, fig.cap="(ref:interface)", out.width="100%"}
draw_image("gameplay", "totems")
```
## Participants {-}

Participants were undergraduate students at the University of Wisconsin-Madison
who completed the experiment in exchange for course credit. Students received
credit based on the duration of each completed session. Sessions were either 25
minutes or 50 minutes. Participants assigned to the Isolated, multi-session
condition who completed all four sessions were awarded additional credit as
an incentive to prevent dropout.

(ref:condition-counts) Total participant and team recruitment.

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "(ref:condition-counts)")
```

# Experiment 1: Simple diachronic inheritance {-}

```{r 50min}
```

In Experiment 1, we compared the simplest form of diachronic inheritance---a
two-person diachronic team where each person is given 25 minutes---to two
alternative strategies for allocating the same 50 minutes of labor: isolated
individuals working alone and two-person synchronic teams working in parallel
(Fig. \@ref(fig:team-structures-50min)).

(ref:team-structures-50min) Team structures compared in Experiment 1.

```{r team-structures-50min, out.width="50%", fig.cap="(ref:team-structures-50min)"}
crotchet::draw_graphviz("team-structures-50min", package = "totems")
```

## Results {-}

We compared the performance of participants assigned to the diachronic,
synchronic, and isolated strategies in three ways. First we compared
performance based on the total number of innovations that were discovered.
Second we report guessing effectiveness on a per-item basis. Were some items
easier to discover by some team structures than others? Finally, we compared
the guessing strategies used by participants in each condition In our analyses,
we focus on comparing the effectiveness of diachronic inheritance to
alternative strategies for solving the same problems.

### Innovations {-}

The number of innovations discovered using each of the three team strategies is
shown in Fig. \@ref(fig:innovations-50min). Our primary hypothesis concerned
the effectiveness of a diachronic strategy as compared to the alternative
strategies for solving the same problems in the same amount of time, but
without vertical inheritance. To test this, we fit a hierarchical linear
model predicting total innovations for each participant based on team strategy,
with strategy contrast-coded such that a diachronic strategy was the reference
group. The results suggest that when controlling for labor hours, a diachronic
strategy is similarly effective to either alternative strategy. Diachronic
teams discovered more innovations on average than isolated individuals, but
post hoc tests revealed that this difference was not statistically significant,
`r exp1$DG2_v_I50`. Conversely, diachronic teams discovered fewer innovations
on average than two-person synchronic teams, but again this difference was not
statistically significant, `r exp1$S2_v_DG2`.

We did find an overall benefit to teamwork in this task. To test this,
we compared two-person teams (both diachronic and synchronic) to isolated
individuals. In the hierarchical linear model, strategy was coded with
Helmert contrasts. Teams performed reliably better on this task than individuals,
`r exp1$teamwork_stats`, without a significant residual difference between the
diachronic and synchronic strategies, `r exp1$teamwork_residual`. Post hoc tests
revealed that this effect of teamwork was driven primarily by synchronic teams
outperforming isolated individuals, `r exp1$S2_v_I50`. This result, showing an
advantage of teamwork over individual labor, suggests that maximizing individual
learning time is not always the most effective way to improve problem solving
performance.

(ref:innovations-50min) Innovations discovered in 50 labor minutes by strategy.
**A.** Rate of innovation by labor hours. Each line is the average number of
innovations discovered by participants in each condition. **B.** Final
innovations by strategy. Each point is a single participant. Bars with errors
show model predictions with ±1 standard error.

```{r innovations-50min, fig.width=5, fig.height=3, fig.cap="(ref:innovations-50min)"}
grid.arrange(
  innovation_rate_50min_plot + ggtitle("A"),
  num_innovations_50min_plot + ggtitle("B"),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

### Guesses per tool {-}

In addition to comparing the three strategies in terms of overall number of
tools, we also performed a more fine-grained analysis of guessing behaviors. We
compared the three strategies in terms of how many guesses were required to
discover each tool. To count the number of guesses that were required for each
tool, we tallied all guesses made from the moment in which a new tool was
eligible for discovery until that tool was discovered. A new tool was eligible
for discovery once all of the items required to create the new tool had been
discovered. We refer to new tools that are eligible for discovery as "adjacent",
since they are adjacent to the current inventory in the solution landscape. The
minimum number of adjacent tools for any inventory was one, indicating that
there was a single tool that must be discovered before new tools can be
created. The median number of adjacent tools was four. When there was more than
one adjacent tools, each incorrect guess was counted toward the total number of
guesses required for each of the adjacent tools.

We found that diachronic inheritance reduced the average number of guesses
required to discover each tool relative to the same tools being discovered by
participants in conditions without diachronic inheritance,
`r exp1$guesses_per_item_inheritance` (Fig.
\@ref(fig:guesses-per-innovation)A). However, this overall benefit
to inheritance in terms of fewer guesses per tool did not extend to tools that
had not been discovered by the previous generation. Comparing the number of
guesses per tool for only the tools that were newly discovered by the second
generation diachronic participants did not reveal a benefit to inheritance
that extended beyond the inherited solutions,
`r exp1$guesses_per_new_item_inheritance` (Fig. \@ref(fig:guesses-per-innovation)B).

(ref:guesses-per-innovation) Guesses per tool by inheritance condition.
Each thin gray line shows the average number of guesses for a particular
innovation. The no inheritance condition includes guesses from synchronic
and isolated participants, as well as first generation diachronic players.
The inheritance condition includes second generation diachronic
participants only. **A.** Guesses per tool for all tools. **B.** Guesses per
tool for new tools, not discovered by an ancestor.

```{r guesses-per-innovation, fig.width=5, fig.cap="(ref:guesses-per-innovation)"}
gridExtra::grid.arrange(
  guesses_per_item_plot + ggtitle("A"),
  guesses_per_new_item_plot +
    ggtitle("B") +
    theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank()
    ),
  nrow = 1,
  widths = c(0.55, 0.45)
)
```

### Guessing strategy {-}

The analyses of performance reported above treat all incorrect guesses as the
same, but in reality some guesses are more informative than others.
Specifically, in this task an optimal guessing strategy is to only attempt
unique guesses. However, many people make repeat guesses, thus employing a
suboptimal strategy. The nature of these repeat guesses varies, as there are
four types of guesses that could be made, shown in Table
\@ref(tab:guess-types). We next report our analysis of whether participants in
each team structure employed different guessing strategies.

(ref:guess-types) Guess types. Each guess can be categorized into one of four
types. Redundant guesses were incorrect guesses made (at least) once before.
Repeat items were guesses that recreated previously discovered tools.
Unique guesses were incorrect guesses never made. Unique items were unique
guesses that created a new tool.

```{r guess-types, results="asis"}
recode_guess_type() %>%
  select(`Guess type` = GuessTypeLabel,
         `Unique guess` = UniqueGuess,
         `Created item` = CreatedItem,
         `Unique result` = UniqueResult) %>%
  kable(caption = "(ref:guess-types)")
```

We calculated the proportion of guesses made by participants in each condition
falling into the four guess types. Our results are shown in Fig.
\@ref(fig:guess-props). For our analysis, we focused on the proportions of
redundant and unique guesses. Looking just at redundant guesses, we found
that inheritance caused participants to guess fewer redundant guesses
than participants in the other strategies, `r exp1$prop_redundant_inheritance`.
Post hoc comparisons revealed that second generation diachronic
participants had a lower proportion of redundant guesses than first
generation diachronic participants, `r exp1$prop_redundant_dg2`, and isolated
individuals, `r exp1$prop_redundant_i50`, but not synchronic participants,
`r exp1$prop_redundant_i50`.

Looking at the proportion of unique guesses, we found no overall effect of
inheritance, `r exp1$prop_unique_inheritance`. Pairwise comparisons revealed
a significant difference between second generation diachronic participants
and isolated individuals, `r exp1$prop_unique_i50`, but neither
first generation participants, `r exp1$prop_unique_dg2` nor
synchronic participants, `r exp1$prop_unique_s2`.

(ref:guess-props) Guessing strategies by team structure. See Table
\@ref(tab:guess-types) for a description of the types of guesses. DG1 and DG2 refer
to first and second generation diachronic participants. I50 refers to isolated
participants given 50 minutes. S2 refers to two-person synchronic teams.

```{r guess-props, fig.width=5, fig.cap="(ref:guess-props)"}
prop_guess_types_50min_plot
```

## Discussion {-}

In Experiment 1, we found that the simplest form of vertical transmission, in
which one person inherits from a previous generation, is comparable to
alternative strategies for solving the same set of problems in the same amount
of labor time (see Fig. \@ref(fig:types-of-type)). Two generations of
diachronic problem solvers each working for 25 minutes results in similar final
performance to an isolated individual given 50 minutes and two problem solvers
working in parallel for 25 minutes each.

We detected the impact of inheriting from a previous generation in the guessing
efficiency and guessing strategies employed by diachronic problem solvers.
Inheriting the solutions from a previous generation improved the efficiency of
future generations of problem solvers by allowing them to recreate those same
tools with fewer guesses overall. However, this improvement in efficiency did
not extend to future problem solving, as second generation diachronic problem
solvers were not any more efficient in solving problems that had not been
discovered by their ancestor.

In addition, second generation diachronic participants employed a different
guessing strategy as reflected in the distribution of guesses by type. In the
Totems game, redundant guesses are unhelpful in that they will never yield new
discoveries, and yet most participants generate a significant proportion of
redundant guesses. We found that the tendency to guess redundantly may be
lessened by diachronic collaboration, as second generation diachronic problem
solvers generated fewer redundant guesses than their ancestors and participants
in an isolated strategy. However, similar proportions of redundant guesses were
generated by participants in the synchronic strategy, suggesting that
diachronic inheritance is not the only strategy for reducing redundancy.

The results of Experiment 1 also support and extend the findings reported by
@Derex:2015cbb that synchronic collaboration can result in an improvement in
problem solving performance relative to isolated individuals working alone. Here
we demonstrate that this advantage holds even when labor hours are controlled.

The results of Experiment 1 suggest that, rather than engaging in a problem
solving task as an isolated individual, it might in some cases be more efficient 
to pass the solutions on to a future generation via diachronic collaboration.
In Experiment 2, we test this idea further by comparing the benefit of
inheriting from someone else, as in the case of diachronic teams, to the
benefit of inheriting from oneself, over multiple sessions.

# Experiment 2: Diachronic versus individual inheritance {-}

```{r SelfOther}
```

In Experiment 2, we compared the effectiveness of four generations of
diachronic inheritance---inheriting from someone else---to the effectiveness of
isolated individuals given four separate sessions (Fig.
\@ref(fig:team-structures-selfother)). Participants in the isolated condition
came back to the experiment for four separate sessions, each 25 minutes long.
At each session after the first, these participants inherited the recipes for
the tools they had created at the end of the previous session. Thus, in this
experiment, both diachronic and isolated individuals started each session
recreating inherited items before extending the inherited solutions to create
new tools.

(ref:team-structures-selfother) Problem solving strategies compared in Experiment 2.

```{r team-structures-selfother, out.width="70%", fig.cap="(ref:team-structures-selfother)"}
crotchet::draw_graphviz("team-structures-selfother", package = "totems")
```

## Results {-}

We compared the performance of four generations of diachronic problem solving
to four sessions of isolated problem solving in two main ways. First, we report
differences in total number of tools discovered. Second, we investigated the
cost of exceeding previously inherited solutions.

Given the high proportion of redundant guesses among isolated individuals
observed in Experiment 1, we were particularly interested in if isolated
individuals, returning for multiple sessions, became fixated on previous and
incorrect solutions, and whether diachronic teams were able to avoid these
inefficiencies of redundant guessing strategies.

### Innovations {-}

The number of innovations discovered in each diachronic generation and
isolated session are shown in Fig.
\@ref(fig:innovations-by-generation-selfother). After a single 25 minute
session, there was no difference between diachronic and isolated participants,
`r exp2$gen1_innovations`. This result was expected, as after a single
session, there was no difference between participants randomly assigned to
either strategy. For diachronic participants, each new generation was able to
discover `r exp2$tools_per_diachronic_gen` more tools than their ancestor,
`r exp2$tools_per_diachronic_gen_stats`. However, each new session of isolated
participants was able to discover `r exp2$tools_per_isolated_gen` more tools
than they did in the previous session, `r exp2$tools_per_isolated_gen_stats`.
That is, an isolated strategy proved more effective than a diachronic strategy
over generations, `r exp2$innovations_by_inheritance_slope`. After four
generations of diachronic collaboration compared to four sessions of isolated
labor, an isolated strategy was more effective, `r exp2$gen4_innovations`.

(ref:innovations-by-generation-selfother) Innovations discovered by diachronic and isolated strategies.
Each thin line is a diachronic team (**left**) or an isolated individual (**right**). Thick lines with
error bars show model predictions with ±1 standard error.

```{r innovations-by-generation-selfother, fig.width=5, fig.cap="(ref:innovations-by-generation-selfother)"}
innovations_by_generation_plot
```

### Guessing strategy {-}

Although participants in the isolated strategy discovered more tools
after four sessions than diachronic teams after four generations,
they did not engage in a more efficient guessing strategy. As in Experiment 1,
participants in the diachronic condition had the most effective guessing
strategy in that they generated a lower proportion
of redundant guesses than participants in the isolated condition,
but in Experiment 2, this different was not statistically significant,
`r exp2$prop_redundant_strategy`. For participants in both strategies,
the proportion of redundant guesses decreased over generations and sessions,
`r exp2$prop_redundant_gen`. However, this decrease was not affected
by strategy, `r exp2$prop_redundant_inter`.

For the proportion of unique guesses, there were no differences between
participants in either strategy, `r exp2$prop_unique_strategy`, or
over generations `r exp2$prop_unique_gen`. Any change in proportion
of unique guesses over generations was not affected by strategy,
`r exp2$prop_unique_inter`.

(ref:guessing-strategy-by-generation) Proportion of guesses by participants in
diachronic and isolated conditions over generations.

```{r guessing-strategy-by-generation, fig.width=5, fig.cap="(ref:guessing-strategy-by-generation)"}
prop_guess_types_selfother_plot
```

If participants in the isolated strategy are not guessing more efficiently
than participants in the diachronic strategy, why then are they able
to accumulate more tools than participants in the diachronic condition?
To answer this question, we investigated the time spent recreating
the inherited tools versus time spent inventing new tools.

### Learning times {-}

Participants took on average `r exp2$mean_learning_time` minutes of the 25
minute session (`r exp2$prop_learning_time`%) to recreate the inherited
tools---a portion of the experiment we refer to as the learning period. We
found that isolated individuals returning for multiple sessions took significantly
less time to recreate the tools they had invented at the end of their previous
session as compared to diachronic participants who were recreating these tools
in the Totems game for the first time, `r exp2$learning_times_inter` (Fig.
\@ref(fig:learning-times)). This result suggests that what made an isolated
strategy successful in Experiment 2 is not that isolated individuals were better
at guessing new tools, but that they were able to more quickly recreate the
tools they inherited, and thus had more time to dedicate to discovering new tools.
Diachronic participants, who were exposed to the game for the first time,
took longer to recreate the inherited items, and had less time to discover
new tools.

(ref:learning-times) Learning times by strategy. Learning time refers to
the number of minutes it took to recreate the inherited items. For
diachronic participants (**left**), learning time scaled linearly with
generations, but for isolated participants returning for multiple sessions,
(**right**), learning times were stable across sessions.

```{r learning-times, fig.width=5, fig.cap="(ref:learning-times)"}
learning_times_plot
```

### Fixation {-}

Both diachronic and isolated participants first had to recreate the tools
discovered by the previous generation participant or in the previous session.
Given the tendency for individuals to get fixated in this task, we asked
whether a diachronic or isolated strategy was more effective at discovering new
tools. To test this, we counted the number of guesses it took each participant
to make the first new tool after recreating all inherited tools. We found that
diachronic participants were able to discover the first new tool with fewer
guesses than isolated participants (Fig. \@ref(fig:cost-of-first-innovation)).

(ref:cost-of-first-innovation) Number of guesses required to discover the first
new tool each session.

```{r cost-of-first-innovation, fig.width=5, fig.cap="(ref:cost-of-first-innovation)"}
first_discovery_by_generation_plot
```

## Discussion {-}

In Experiment 2, we found that isolated individuals given four sessions
were able to outperform diachronic individuals, each given only 25 minutes.
However, isolated individuals were not always more efficient problem
solvers than diachronic individuals. Specifically, we found that isolated
individuals were more likely to be fixated on incorrect solutions, causing
them to guess inefficiently. These fixation effects extended over multiple
sessions, causing the performance of some isolated individuals to stagnate.
Diachronic participants, however, were less impacted by fixation, and were
more easily able to move beyond the solutions they inherited to discover
new solutions. These results reveal an important benefit to diachronic
collaboration and inheritance.

In Experiment 2, we compared teams given 100 labor minutes. In Experiment 3,
we report the performance of four person synchronic teams given the same
number of labor minutes.

# Experiment 3: Scalability of diachronic and synchronic collaboration {-}

```{r TeamSize}
```

In Experiment 3 we measured the scalability of diachronic and synchronic collaboration.
Does adding more people to the problem solving team improve the performance of
each strategy equally? We hypothesized that diachronic collaboration would be
more scalable than synchronic collaboration. To test this, we compared
diachronic and synchronic collaboration given 50 minutes to the same strategies
given 100 minutes.

## Results {-}

### Innovations {-}

As reported in Experiment 1, two-person synchronic teams discovered on average more innovations than two-generation
diachronic teams, although this difference was not significant. Comparing four-person synchronic
teams to four-generation diachronic teams revealed a different pattern. Four-person diachronic
teams discovered on average more innovations than four-person synchronic teams, resulting in
a marginal interaction between team strategy and team size. These results suggest that a diachronic
strategy is more scalable than a synchronic strategy, meaning that adding more people to a
diachronic team may result in a bigger improvement in final performance compared to adding
people to a synchronic team.

(ref:max-innovations-by-teamsize) Innovations by team size. Performance of
diachronic teams after two or four generations is compared to performance of
synchronic teams with two or four individuals.

```{r max-innovations-by-teamsize, fig.width=5, fig.cap="(ref:max-innovations-by-teamsize)"}
max_innovations_by_teamsize_plot
```

### Simulations {-}

We used an agent-based simulation to model the scalability of diachronic and
synchronic strategies. In the simulations, problem solvers guessed combinations
of between 1 and 4 items at random. Diachronic and synchronic strategies were
provided the same number of guesses. Diachronic problem solvers guessed one at
a time, while synchronic problem solvers guessed in parallel. If a guess
yielded an item, it was added to the team inventory. For synchronic problem
solvers, guesses were added to each participant's inventory immediately.

The results show that as problem solvers are added to each team, a synchronic
strategy becomes progressively less effective as compared to a diachronic
strategy (Fig. \@ref(fig:bots-teamsize)). The reason is that synchronic players
increasingly make the same incorrect guesses because each player is generating
guesses from the same inventory. Diachronic players in contrast are not as
likely to generate the same incorrect guesses as a previous generation because
they are generating guesses from different sets of items.

(ref:bots-teamsize) Result of simulations of diachronic and synchronic problem solvers.

```{r bots-teamsize, fig.width=5, fig.cap="(ref:bots-teamsize)"}
bots_team_size_plot
```

## Discussion {-}

In Experiment 3, we demonstrated that diachronic and synchronic collaboration differ
in their ability to scale up to larger teams. Diachronic collaboration can add
more generations that each correspond to an increase in team performance, but the utility
of adding problem solvers to a synchronic strategy is smaller. Specifically we found
that the reason might be related to synchronic individuals each making the same
incorrect guesses in tandem, meaning that each individual is unable to only guess
unique combinations for the team.

# Discussion {-}
