---
title: "What is the impact of inheritance on problem solving ability?"
author: "Pierce Edmiston"
abstract: |
  The cognitive abilities believed to enable the accumulation of cultural
  products by humans and no other animal include the ability to pass on an
  innovation, such as an improvement in some tool, from one generation to the
  next. As a result, problem solvers rarely have to start from scratch when
  solving even new problems, as demonstrated by the prevalence of reuse in the
  evolution of human technology. But what is the impact of inheritance---
  starting with a solution that was created by someone else---on problem solving
  ability? Does inheriting an existing solution to a problem simply save time in
  future problem solving, or does it change the way problems are solved for
  future generations? The experiments reported here measure the impact of
  inheritance on problem solving ability. In the inheritance condition, one
  person works on a solution to a problem before passing it on to be inherited
  by another person to continue. Because the people this exchange are working
  together, I refer to them as Diachronic teams. The problem solving abilities
  of a diachronic strategy organized around inheritance is compared to
  alternative strategies for allocating the same number of labor hours:
  Synchronic teams comprising individuals working in parallel, and Isolated
  individuals working for the same total number of labor hours.
csl: "theme/apa.csl"
bibliography: "refs/diachronic-collaboration.bib"
output:
  html_document:
    theme: flatly
    css: theme/style.css
    toc: true
    toc_depth: 2
    toc_float: true
---
```{r config, include = FALSE}
library(knitr)
library(printr)
library(magrittr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 4,
  fig.height = 4,
  cache = TRUE
)

list.files("R", "*.R", full.names = TRUE) %>% sapply(read_chunk)
read_graphviz_chunk("team-structures", "totems")
read_graphviz_chunk("diachronic-collaboration", "totems")
```

```{r setup, cache = FALSE}
```

# Introduction

On November 1, 2015, the English Wikipedia published its 5 millionth article.
The article, on a shrub native to eastern Australia, was only 9 words long.
Since then, the article has been edited 419 times by 86 different authors, and
now includes 8 different sections on the taxonomy, habitat, and ecology of this
Australian plant. The collaborative effort of writing a Wikipedia article
exhibits a type of teamwork that is foundational to cumulative cultural
evolution, and may be unique to our species. Humans are particularly adept at
using, changing, and in some cases improving cultural products that were first
created by other people. The ability to inherit and improve a tool or some other
solution to a problem created by a previous generation of problem solvers is
necessary in order to observe the ratcheting of technological complexity in a
culture [@Tomasello:1993vh]. I refer to the type of collaboration that enables
ratcheting as **diachronic collaboration** (Fig. 1). In diachronic
collaboration, the products of labor, such as a tool to solve some
problem, are shared with or inherited by future generations of problem solvers.
The purpose of this research is to understand why diachronic collaboration may
be crucial for cumulative cultural evolution by assessing its effectiveness as a
problem solving strategy.

```{r diachronic-collaboration, engine="dot", fig.width=6, fig.pos="h", out.extra="", fig.cap="In diachronic collaboration, the products of one generation of labor are inherited to be used and improved by the next generation. An example of diachronic collaboration is Wikipedia article writing, where articles are effectively passed from editor to editor in a chain of edits. The ability to engage in diachronic collaboration---to improve something that was created by someone else---is required in order to observe the ratcheting of technological complexity in a culture."}
```

Humans are effective problem solvers. Part of this effectiveness comes from the
ability to reuse and otherwise adapt solutions to problems that were first
created by other people. Inventors rarely create inventions with no precursors,
and commonly use and adapt inventions that were first created by others
[@Basalla:1989wq]. Part of this reuse is the result of inventions breeding
further inventions [@Valverde:2007fs; @Borner:2004dm]. But what do innovators
and other problem solvers gain from inheriting solutions created by previous
generations? Do diachronic problem solvers simply get a head start on future
problem solving, or does inheriting an existing solution to a problem actually
change the way future problems are solved?

The purpose of this research is to identify the conditions under which
inheritance can be predicted to improve one's problem solving effectiveness
relative to alternative strategies for solving the same set of problems. I
compare the effectiveness of a diachronic strategy to two alternative strategies
for allocating the same number of labor hours: synchronic collaboration, where
team members begin with nothing and collaborate in parallel, and isolated
individuals, given the most time for individual learning (Fig. 2). The purpose
of this comparison is to understand the impact of inheritance on problem solving
by comparing problem solving ability when starting with an inherited
solution---defined as diachronic problem solvers after the first generation---to
alternative strategies for solving the same problems without diachronic
inheritance. Discovering whether and under what conditions diachronic
collaboration is more effective than alternative strategies will provide for the
first time empirical evidence of the ratcheting of human problem solving ability
via inheritance.

```{r team-structures, engine="dot", fig.width=2, fig.pos="h", fig.cap="Problem solving strategies. Each node is a person. Edge lengths represent labor hours spent solving a problem. In diachronic collaboration, first generation problem solvers work on solutions to a problem before passing them on to be inherited by the next generation. The impact of inheritance on problem solving is measured as the problem solving ability of diachronic problem solvers after the first generation. Isolated individuals do not inherit solutions from other players, but are given the most time for individual learning. In synchronic collaboration, a team of problem solvers work in parallel, sharing their progress, but synchronic teams do not benefit from direct diachronic inheritance because all problem solvers start at the same point."}
```

In the experiments below I test two main hypotheses regarding the conditions
under which inheritance might be found to improve problem solving ability
relative to the alternative strategies. The first hypothesis relates to the
kind of problem being solved. Diachronic inheritance is predicted to be
effective in overcoming problems in which fixation effects are common, causing
individuals to return to, or otherwise be mislead by previous and incorrect
solutions to problems. The plausibility of this hypothesis is demonstrated
via artificial simulations of problem solvers, described in detail below.

The second hypothesis is that diachronic inheritance will prove effective
as a function of the skills of problem solvers in subsequent generations. Specifically, the benefits of diachronic inheritance are predicted to be
amplified as the relative skills of the generations of problems solvers
diverge. Inheriting a solution that was created by someone with very different skills effectively starts problem solving off in a unique part of the solution
landscape that would be unlikely to be reached as an isolated individual
or a synchronic team.

The effectiveness of diachronic collaboration likely depends on the shape of the
solution landscape. A solution landscape is a visualization of the relationship
between variation in possible solutions and solution success, analogous to
fitness landscapes in an evolutionary sense. As the abilities of diachronic
problem solvers diverge, subsequent generations effectively start off in parts
of the solution landscape that would have been unreachable as an isolated
individual. For at least some solution landscapes, diachronic teams with
divergent skills are predicted to outperform isolated individuals, as
is demonstrated via artificial simulations below.

## Measuring problem solving ability

I measure problem solving ability as efficiency in completing a set of problems
in a specific amount of time. Other research on problem solving aims to determine
the cognitive abilities that enable the generation of problem solutions. Here
I focus less on whether solutions can be generated and more on how
efficiently progress is achieved. The goal of this research is to measure the
impact of inheritance on problem solving efficiency. Assuming a constant
problem landscape and diachronic problem solvers of equalivant abilities, the impact of inheritance on problem solving is expected to be better fit
by an interactive model than an additive model.

\begin{equation}
problem\ solution = ability \times time = inheritance\\
problem\ solution_{additive} = inheritance + ability \times time\\
problem\ solution_{interactive} = inheritance \times ability \times time\\
\end{equation}

```{r types-of-time, fig.width=6, fig.height=3, fig.cap="**Left.** Relationship between labor hours and calendar hours. Labor hours are being controlled across the strategies, which forces diachronic and synchronic teams to differ in calendar hours. The rate at which labor hours are spent (per calendar hour) is the same for diachronic teams and isolated individuals. **Right.** Differences between teams in terms of time for individual learning. Isolated individuals have the most time for individual learning. The benefits of individual learning are balanced against the potential costs of fatigue and fixation."}
```

# Methods

## Materials and Procedure

Participants solved a set of problems where solutions accumulate and must be
recombined to solve new problems. To solve the problems, participants played a
computer game, first reported in @Derex:2015cbb, that involves building
incrementally more complex tools in pursuit of building "a sacred totem to
appease the gods." To build a totem, participants first needed to construct an
axe out of three independently invented items: a refined stick used as a handle,
a sharpened rock for the blade, and string wound from bark fibers for binding
(Fig. 3). In the game, more advanced tools produce larger and more intricate
totems, resulting in higher performance scores.

```{r gameplay, fig.width=6, fig.height=4, fig.cap="The Totems game. **Left.** A sample of the problem landscape. The top row of 6 items are the resources available to problem solvers at the start of the experiment. Resources are combined to build the first generation of innovations, including a refined rock, a club derived from an antler, and a small branch of a tree. The first 6 generations of innovations are shown. The axe is required to construct the first totem pole. Although the full solution landscape contains more generations of innovations, many participants do not proceed past these 10 innovations in a single 25 minute session. **Right.** Gameplay interface. "}
grid.arrange(
  read_graphviz_with_images("landscape-sample", "totems"),
  read_image("gameplay", "totems"),
  nrow = 1,
  widths = c(0.35, 0.65)
)
```

## Participants

Participants are recruited until the final sample comprises 40 teams in each of the three strategy conditions for a total of 200 participants.

# 50 Labor Minutes

```{r team-structures-exp1}
draw_graphviz("team-structures-exp1", "totems")
```

## Number of innovations

```{r num-innovations-50}
```

```{r num-innovations-50-fig, fig.height=4, fig.width=8, fig.cap="**Number of innovations by strategy**. Left: The accumulation of innovations by labor time. Right: Final number of innovations discovered by each team or individual."}
grid.arrange(
  num_innovations_over_time_50 + theme(legend.position = c(0.2, 0.8)),
  final_num_innovations_50,
  widths = c(0.6, 0.4)
)
```

```{r num-innovations-by-generation-50-fig}
num_innovations_by_generation_50_plot
```

```{r performance-by-generation-50-mod, echo = 1, results = "asis"}
num_innovations_by_generation_50_mod <- lmer(
  NumInnovations ~ Diachronic_v_Isolated * Generation + (1|TeamID),
  data = NumInnovationsByGeneration50
)
tidy(num_innovations_by_generation_50_mod, effects = "fixed") %>%
  kable()
```

## Likelihood of creating a totem

```{r likelihood-50}
```

```{r likelihood-50-fig}
likelihood50_plot
```

## Guesses

### Total guesses

```{r total-guesses-50}
```

```{r total-guesses-50-fig}
total_guesses_50_plot
```

```{r total-guesses-per-generation-50-fig}
total_guesses_per_generation_50_plot
```

```{r performance-by-total-guesses-50-fig}
performance_by_total_guesses_50_plot
```

```{r performance-by-total-guesses-50-mod, echo = 1, results = "asis"}
performance_by_total_guesses_50_mod <- lm(
  NumInnovations ~ NumGuesses * (DvS + DSvI),
  data = TeamPerformance50
)
tidy(performance_by_total_guesses_50_mod) %>%
  kable()
```

### Unique guesses

```{r unique-guesses-50}
```

```{r unique-guesses-50-fig}
unique_guesses_50_plot
```

```{r unique-guesses-by-total-guesses-50-fig}
unique_guesses_by_total_guesses_50_plot
```

```{r unique-guesses-per-generation-50-fig}
unique_guesses_per_generation_50_plot
```

```{r num-innovations-by-unique-guesses-50-fig}
num_innovations_by_unique_guesses_50_plot
```

### Guess types

```{r types-of-guesses, results = "asis"}
recode_guess_type() %>%
  select(GuessTypeLabel, UniqueGuess, CreatedItem, UniqueResult) %>%
  kable()
```

```{r guess-type-50}
```

```{r prop-redundant-guesses-50-fig}
prop_redundant_guesses_50_plot
```

```{r prop-repeat-items-50-fig}
prop_repeat_items_50_plot
```

```{r prop-unique-guesses-50-fig}
prop_unique_guesses_50_plot
```

```{r prop-guess-types-50-fig}
prop_guess_types_plot
```

```{r prop-guess-types-by-generation-50-fig}
prop_guess_types_by_generation_50_plot
```

## Performance per inventory

```{r performance-per-inventory-50}
```

```{r total-guesses-per-problem-50-fig}
total_guesses_per_problem_50_plot
```

```{r total-guesses-per-problem-50-mod-fig}
total_guesses_per_problem_50_mod_plot
```

```{r total-labor-time-per-problem-50-fig}
total_labor_time_per_problem_50_plot
```

# 100 Labor Minutes

```{r team-structures-exp2}
draw_graphviz("team-structures-exp2", "totems")
```

## Performance

```{r performance-100}
```

```{r performance-100-fig, fig.height=4, fig.width=8, fig.cap="**Problem solving performance by strategy**. Left: The accumulation of inventions over labor time. Right: Final team performance."}
grid.arrange(
  performance_over_time_100 + theme(legend.position = c(0.2, 0.8)),
  final_performance_100,
  widths = c(0.6, 0.4)
)
```

# Discussion

# References
