---
title: "The impact of inheritance on problem solving ability"
author: "Pierce Edmiston, Max Derex, & Gary Lupyan"
abstract: |
  Human accumulate cultural products including tools and other technology like
  no other species of animal. A consequence of accumulating generations of
  solutions to problems is that new generations of problem solvers rarely have
  to start from scratch when solving new problems, as demonstrated by the
  widespread prevalence of reuse in the history of human technology. But what
  is the impact of inheritance---starting off a problem solving task with
  a partial solution that was created by someone else---on future problem
  solving ability? Does inheriting solutions discovered by a previous
  generation just save time in future problem solving, or does it affect
  the rate at which new problems are solved? To answer this question, we
  compared problem solving with inheritance to two alternative strategies for
  solving the same set of problems without inheritance. Critically, we compare
  these strategies while controlling for total labor time so that the
  consequences of inheritance can be interpreted relative to both the cost of
  creating it and the cost of not using an alternative problem solving strategy.
csl: "theme/apa.csl"
bibliography: "references.bib"
output:
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    toc_float: yes
    css: theme/style.css
    keep_md: yes
    number_sections: no
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)

# Read all the diagrams made with graphviz in the totems package
chunks <- c(
  "team-structures",
  "team-structures-all",
  "team-structures-exp1",
  "team-structures-exp2",
  "team-structures-exp3",
  "team-structures-exp4"
)
sapply(chunks, read_graphviz_chunk, package = "totems")
```

```{r setup, include=FALSE, cache=FALSE}
```

# Introduction

```{r intro, include=FALSE}
```

Humans are effective problem solvers. Part of this effectiveness comes from the
ability to reuse and adapt solutions to problems that were first created by
other people. Inventors rarely create inventions with no precursors, and
commonly use and adapt inventions that were first created by other people
[@Basalla:1989wq; @Valverde:2007fs]. But what do innovators and other problem
solvers gain from inheriting solutions created by others? Do problem solvers
gain a head start on future problem solving, or does inheriting an existing
solution to a problem affect the way in which future problems are solved? We
attempt to answer this question by studying the problem solving ability of
individuals who inherit solutions created by a previous generation at the start
of a problem solving task. This type of collaboration, organized around
inheritance, is referred to as **diachronic collaboration** 
(Fig. \@ref(fig:team-structures)). To understand how inheritance impacts problem
solving ability, we compared the effectiveness of diachronic collaboration to
two alternative strategies for solving the same set of problems in the same
amount of time but without diachronic inheritance: **synchronic collaboration**,
where team members begin with nothing and collaborate in parallel, and
**isolated individuals**, given the most time for individual learning
(Fig. \@ref(fig:team-structures)).

(ref:team-structures) Problem solving strategies. Each node is a single problem
solver. Edge lengths represent labor hours spent solving a set of problems. In
diachronic collaboration, first generation problem solvers work on solutions to
problems before passing them on to be inherited by the next generation. The
impact of inheritance on problem solving is measured as the problem solving
ability of diachronic problem solvers after the first generation. Isolated
individuals do not inherit solutions from other problem solvers, but are given
the most time for individual learning. In synchronic collaboration, multiple
problem solvers work in parallel, sharing their progress, but synchronic teams
do not benefit from diachronic inheritance because all problem solvers
begin at the same starting point.

```{r team-structures, engine="dot", fig.width=1.5, fig.cap="(ref:team-structures)"}
```

The focus in this research is on diachronic collaboration and the impact of
inheritance on problem solving ability among diachronic collaborators.
Diachronic collaboration exemplifies an important process in human cultural
evolution: the iterative improvement of tools and other solutions to problems
over generations of problem solving generally known as ratcheting
[@Tennie:2009gaa; @Tomasello:1993vh]. Diachronic collaboration is enabled by
social learning mechanisms that allow problem solvers to learn from previous
generations of problem solvers rather than attempting to solve the same problems
through independent and individual learning. Models of the cultural evolutionary
process demonstrate that social learning is adaptive relative to individual
learning when the environment is stable and individual learning is costly
[@Boyd:1988cu]. A population of smart individual learners who are unable to
learn from each other are limited in what they can accomplish in a single
lifetime, as their accomplishments cannot be inherited by any future problem
solvers. Thus, social learning is believed to provide a valuable shortcut to
individual learning. The goal of this research is to quantify this shortcut. We
measure how much time is saved by inheriting existing solutions to problems, 
and how inheritance impacts future problem solving.

Previous efforts to understand diachronic collaboration in a problem solving
context have focused on the capacity to inherit solutions from others
rather than the effectiveness of diachronic collaboration as a problem solving
strategy. For example, @Tennie:2009gaa argue that humans as compared to nonhuman
primates are uniquely able to inherit solutions to problems that they would be
unable to discover on their own. @Caldwell:2008gf demonstrated the capacity
of diachronic teams to cumulatively improve performance in tasks such as
paper airplane flying and spaghetti tower building.

The synchronic and isolated strategies serve as control conditions in these
experiments: ways of solving the same set of problems without diachronic
inheritance. Synchronic and isolated problem solvers begin the problem solving
task at the same starting point, i.e. with no existing solutions. Only
diachronic problem solvers after the first generation inherit solutions from
the previous generation to learn from at the start of the experiment.

The diachronic condition will be referred to as the inheritance condition, but
other types of inheritance that do not influence the starting point of a problem
solver beginning a problem solving task are included in the control conditions.
Synchronic problem solvers share solutions with one another throughout the
problem solving task. What distinguishes diachronic and synchronic inheritance
is the time at which the solutions are shared. For diachronic problem solvers,
solutions are inherited once, at the start of the problem solving task. For
synchronic problem solvers, solutions are inherited intermittently, throughout
the task. A comparison of diachronic and synchronic problem solving (Exps. 2,4)
will reveal whether diachronic-generational or synchronic-intermittent
inheritance mechanisms result in more effective problem solving.

A third type of inheritance, individual inheritance, is introduced in Experiment
3. Individual inheritance is when the products of one's own problem solving are
resumed in independent problem solving sessions. Individual inheritance
involves inheriting from oneself, as opposed to inheriting from someone else.
Experiment 3 investigates how diachronic inheritance, the inheritance of partial
solutions at the start of a problem solving task, impacts problem solving
ability differently than an isolated individual, resuming problem solving in
subsequent sessions.

In order to fully understand the impact of inheritance on problem solving
ability, it is necessary to account for both the cost of creating the
inheritance as well as the consequences of inheritance on future problem
solving. In this study the cost and consequence of inheritance is controlled by
allotting a fixed number of labor hours to diachronic teams. The other
strategies investigated in this research are compared given the same number of
labor hours. Labor hours can be contrasted by two other measures of time that
are not controlled in these experiments: calendar hours and individual learning
hours (Fig. \@ref(fig:types-of-time)).

(ref:types-of-time) Types of time. **Left.** A key showing the problem solving
strategies for reference. See Fig. \@ref(fig:team-structures). **Middle.**
Relationship between calendar hours and labor hours. Synchronic collaboration
consumes a fixed amount of labor time more rapidly (per calendar hour) than
either the diachronic or isolated strategies. **Right.** Amount of individual
learning time in each strategy. An isolated strategy maximizes the amount of
time for individual learning relative to both diachronic and synchronic
strategies.

```{r types-of-time, fig.width=8, fig.height=3, fig.cap="(ref:types-of-time)"}
grid.arrange(
  read_graphviz("team-structures", "totems"),
  gg_time,
  gg_person,
  nrow = 1,
  widths = c(0.2, 0.4, 0.4)
)
```

# Methods

```{r methods, include=FALSE}
```

Participants solved as many problems as they could in a fixed amount of time. To
solve each problem, participants had to guess combinations of items that would
result in new items being created. The new items, created from combinations of
previous items, are referred to as innovations, and each innovation is the
solution to a single problem amid the larger set of problems in the problem
solving task. The context of the problem solving task is that participants are
accumulating innovations like axes, stone knives, and paintbrushes in service of
creating "a sacred totem pole to appease the gods." As participants accumulate
more complex innovations, they can then create more intricate and elaborate
totem poles, and receive higher performance scores as a result.

Participants were randomly assigned to either the diachronic condition (the
treatment condition) or one of the control conditions: isolated or synchronic.
Participants in the diachronic condition either participated as a first
generation problem solver, or began the problem solving task with the solutions
discovered by a previous generation. Each participant was assigned to a single
generation in a single team; the innovations of each diachronic problem solver
were only inherited by, at most, one future generation problem solver.

Participants in the synchronic condition collaborated in parallel via separate
computers. Team membership was not revealed to the participants either before or
after the experiment, and no interaction, other than through the computer
interface, was allowed. The only interaction that was allowed was the sharing of
innovations with all members of a synchronic team. Once an innovation was
discovered by a synchronic participant, the recipe for the creation of that
innovation became available for all members of the synchronic team.

Participants in the isolated condition participated in a single 50 minute
session, or alternately returned for four 25 minute sessions, spaced at least
one day apart. When isolated participants returned for later sessions, they
inherited the tools they had discovered in the previous session.

The data generated using these methods are reported in four experiments.
Experiment 1 investigates the result of accumulated inheritance over four
generations of diachronic problem solving. The analyses explore the differences
between first generation problem solvers and problem solvers who benefit from
inheritance, defined as diachronic problem solvers after the first generation.
In Experiment 2, we compare the simplest form of diachronic inheritance, a
two-person diachronic team, to a two-person synchronic strategy, and an isolated
individual given 50 minutes. In Experiment 3, diachronic inheritance from
someone else is compared to inheriting from oneself: isolated participants
returning for multiple sessions. Experiment 4 compares the three main strategies
(diachronic, isolated, and synchronic) given 100 minutes, including a
four-person synchronic strategy.

(ref:team-structures-all) All experimental conditions. The **Diachronic**
condition is the treatment condition: problem solving with inheritance. First
generation diachronic participants began without any previous solutions and
worked on the problem solving task for 25 minutes. Diachronic participants
assigned to generations 2-4 inherited the solutions discovered by the previous
generation at the start of the problem solving task. The two control conditions
were Isolated and Synchronic. **Isolated** participants either completed a
single 50 minute session, or returned for 4 subsequent 25 minute sessions. Upon
returning, these participants "inherited" the solutions they had created in the
previous session. **Synchronic** participants completed the problem solving task
in groups of 2 or 4 individuals, each participating for 25 minutes. Synchronic
participants collaborated in parallel, sharing solutions with all team members
as they were discovered.

```{r team-structures-all, engine="dot", fig.cap="(ref:team-structures-all)"}
```

## Materials and Procedure

**The Totems game**. Participants played a computer game, first reported in
@Derex:2015cbb, requiring players to iteratively solve problems related to
creating new tools out of existing items for the purposes of building "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered items: a refined stick
used as a handle, a sharpened rock for the blade, and string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). In the game, more
advanced tools produce larger and more intricate totems, resulting in higher
performance scores.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6
items are the initial resources available to problem solvers at the start of the
session. Resources must be combined to build the first generation of
innovations, including a refined rock, a club derived from an antler, and a
small branch of a tree. The first 6 generations of possible innovations are
shown. The axe is required to construct the first totem pole.

```{r landscape-sample, fig.cap="(ref:landscape-sample)"}
# draw_graphviz_with_images("landscape-sample", "totems")
draw_image("landscape-sample.gv", "totems")
```

**Guesses.** Participants solved problems by trial-and-error guessing of different
combinations of items they believed would yield new tools. To generate new
guesses, participants dragged items into a workshop panel (Fig.
\@ref(fig:interface)).

Of all the valid combinations of items that could be made, only a small
percentage successfully yielded new items. The 6 initial resources can be
combined with between 1 and 4 other items with replacement for a total of
`r count_unique_guesses(6)` unique guesses, only three of which
yield any new tools: a branch can be broken off a tree, an antler refined into
a club, and two rocks can be combined to yield a shaped stone. As the number
of items available to each problem solver increased, so did the combinatorial
complexity of the problem space. Thus, problems became more difficult
over time.

**Stock.** If a guess yielded an item, it could be stored in a stock panel. A maximum of 12
items could be stored in the stock panel at any given time. Items could be
discarded if space was needed in the stock panel for new items, or if an item
was not needed for any reason.

**Recipes.** Once an item was discovered, the recipe for its creation was recorded in the
history panel. Participants could review their item history and see the recipes
for their previous innovations, even items they had discarded.

In addition to an individual history, the history pane also had other tabs,
depending on which strategy each participant was assigned. Diachronic
participants after the first generation began with a tab that could be selected
to show all item recipes discovered by the previous generation. This was the way
in which solutions to previous problems were shared with future generations.
Isolated participants, if returning for multiple sessions, had a tab with the
solutions they had come up with during their previous session. Synchronic
participants had tabs for other synchronic problem solvers on their team.
The tab displayed the current score of each player, so synchronic were notified
that a discovery had been made by seeing a change in their teammates' score.

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the Workshop and selecting the Try button. If the guess
created an item, it could be dragged into the Stock panel, and used again. Once
segments of a totem were made through cutting, carving, and painting, they could be
dragged into the Totem panel for scoring.

```{r interface, fig.cap="(ref:interface)"}
draw_image("gameplay", "totems")
```

**Post experiment survey.** After the experiment was completed, participants
completed a brief post experiment survey assessing their understanding of the
task, their experience with games like the Totems game, and whether they used
any particular strategy for guessing.

## Participants

Participants (N=`n_recruited`) were students at the University of
Wisconsin-Madison who completed the experiment in exchange for course credit.
Students received credit for each completed session based on the duration of the
session, 25 minutes or 50 minutes. Participants in the Isolated, multi-session
experiment who completed all four sessions were awarded additional credit.

Participants were recruited until there were at least 40 teams in each strategy.
Most strategies, other than four-person synchronic teams, ended up receiving
more than 40 teams. The reason for this is that participants were scheduled in
groups of 4 participants at a time, and frequently one or more participants
would fail to show up for their extra credit opportunity. In these cases, the
remaining participants were run as a combination of diachronic and
two-person synchronic participants, resulting in a higher-than-expected
number of participants in those conditions.

Data from individual participants (N=`n_excluded`) were excluded from the final
analyses for the following reasons. Most of these (N=`n_errors`) were excluded
because they experienced errors in the computer game. Others
(N=`n_noncomplaint`) were noncomplaint, giving up on the problem solving task
(N=`n_gave_up`).

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "Total participant and team recruitment.")
```

# Experiment 1

```{r exp1, include=FALSE}
```

In Experiment 1, we measured the impact of inheritance on problem
solving ability over four generations of diachronic problem solving (Fig.
\@ref(fig:team-structures-exp1)). We answered the following questions about the
impact of inheritance on problem solving. First, we determine whether and by how
much future generations of problem solvers were able to recreate and exceed the
solutions obtained by their ancestors. Second, we explore the impact of inheritance
on future problem solving by testing whether larger inheritances are harder or
easier to extend to new problems.

(ref:team-structures-exp1) Diachronic problem solving examined in
Experiment 1. Participants completed a problem solving task for 25 minutes.
Participants in generations 2-4 began the problem solving task with the
solutions that were discovered by the previous generation.

```{r team-structures-exp1, engine="dot", fig.cap="(ref:team-structures-exp1)"}
```

## Methods

Participants played the Totems game for a single 25-minute session. Each
participant was assigned to a single generation in a single team. Participants
in generations 2-4 began the problem solving task with the solutions that were
discovered by the previous generation teammate. The unique innovations of participants
in generations 1-3 were inherited by one (and only one) future generation.

(ref:exp1-participants) Participants in Experiment 1.

```{r exp1-participants, results="asis"}
kable(Exp1Participants, caption = "(ref:exp1-participants)")
```

## Results

Diachronic problem solvers were consistently able to exceed the number of
innovations discovered by their predecessors
(Fig. \@ref(innovations-by-generation)).
This result was obtained using Page's trend test, which is a
repeated measure test for monotonicity. We used this test to measure whether
the number of innovations increased within each team and over generations
against the null hypothesis that the number of innovations did not change
from generation to generation. We found evidence for cumulative
increases in problem solving ability such that later generations of problem
solvers were able to solve more problems in a single 25 minute session than
their predecessors, `r exp1$page_test`.

We next measured the linear increase in innovations over generations by fitting
a hierarchical regression model to the total innovations achieved in each
generation with random effects (intercepts and slopes) for teams. On average,
second generation participants were able to discover `r exp1$gen0_slope` more
innovations than first generation participants, `r exp1$gen0_slope_stats`. This
linear effect decreased by `r exp1$gen0sqr_slope` each generation for third and
fourth generation participants, `r exp1$gen0sqr_slope_stats`. A model comparison
of hierarchical regression models fit with linear and quadratic components
revealed that a quadratic fit was significantly better than a linear fit alone,
`r exp1$quad_modcomp`. This result suggests that accumulated inheritance had a
diminishing return on problem solving.

(ref:innovations-by-generation) Number of innovations discovered by each
generation. Each of the thin green lines is a team. Means in each generation are
shown as green X's. The thick blue line is the model predictions with ±1
standard error.

```{r innovations-by-generation, fig.cap="(ref:innovations-by-generation)"}
innovations_by_generation_plot
```

We next consider different explanations for the diminishing return of inheritance on
problem solving. First, we tested whether a larger inheritance had a detrimental
effect on future problem solving. We found that an increase in the number of
innovations inherited is negatively associated with a decrease in the number of
new innovations discovered by future generations of problem solvers,
`r exp1$inheritance_size_slope_stats` (Fig. \@ref(fig:size-of-inheritance)).
This finding is in line with the current proposal that inheritance has an
influence on future problem solving. In this case, the impact is negative, such
that inheriting more innovations appears to _decrease_ problem solving ability.

(ref:size-of-inheritance) Innovations created relative to those inherited. **The
left plot** shows the relationship between created and inherited innovations. The
dotted line is a reference line with slope=1 such that points above the line
indicate future generations exceeding their ancestors. **The right plot** shows the
relationship between inherited innovations and only the novel innovations. The
same reference line is now shown horizontally. The line shows the predictions of
the hierarchical regression model with ±1 standard error.

```{r size-of-inheritance, fig.width=5, fig.height=3, fig.cap="(ref:size-of-inheritance)"}
grid.arrange(
  innovations_created_and_inherited_plot,
  new_innovations_plot,
  nrow = 1
)
```

However, an alternative explanation, not having to do with inheritance affecting
problem solving ability, is that as more innovations are accumulated, the task
gets more difficult, such that problem solving might slow down irregardless of
inheritance. In the Totems game, as innovations are accumulated, the total
number of possible guesses than can be made from the accumulated items increases
exponentially, so that later generations of problems are less likely to stumble
upon a correct solution by chance. To control for combinatorial complexity, we
created an alternative outcome measure that, rather than counting all
innovations equally, instead weighted each innovation relative to the size of
the inventory at the time at which the innovation was discovered. By this
measure, innovations from later generations are more difficult because they
require combinatorially more guesses in order to find by chance alone.

Fig. \@ref(fig:delta-difficulty-fig) shows the
change in accumulated difficulty score based on the number of innovations
that were inherited. In contrast to the analysis that treated all
innovations equally, when innovations are weighted by combinatorial complexity,
the effects of inheritance and diachronic collaboration appear to be beneficial.
Diachronic problem solvers who inherit more innovations are better
able to exceed the accumulated difficulty score of their ancestors than
first generation diachronic problem solvers, `r exp1$delta_diff_slope_stats`.

(ref:delta-difficulty-fig): Change in accumulated difficulty score by
inheritance. Accumulated difficulty scores are sums of innovations weighted by
the combinatorial complexity of the possible choices available to the
participant when the innovation was discovered. Shown on the y-axis is the
difference between the accumulated difficulty scores of subsequent generations.
The line shows the predictions of a hierarchical regression model with ±1 standard
error.

```{r delta-difficulty-fig, fig.cap="(ref:delta-difficulty-fig)"}
delta_difficulty_plot
```

The final explanation to consider regarding whether inheritance impacts
problem solving performance is that larger inheritances take more time
to recreate, thus leaving less time for future problem solving, but
without influencing the rate at which future problems are solved.
Diachronic problem solvers inherit the recipes for creating
innovations in the Totems game, not the innovations themselves. Thus, diachronic problem
solvers who inherit solutions must start the experiment by recreating the
innovations already discovered by their ancestors. On average, diachronic
problem solvers after the first generation spent `r exp1$mean_learning_time_min`
minutes of a 25 minute session (`r exp1$proportion_learning_time`%) recreating
the inherited innovations in this **learning stage**
(Fig. \@ref(fig:learning-times)A).

The length of the learning stage depended on the number of innovations that were
inherited. On average diachronic problem solvers inherited
`r exp1$mean_inheritance_size` innovations (SD=`r exp1$sd_inheritance_size`). The
length of the learning stage scaled linearly with the size of the inheritance,
`r report_lm_mod(learning_rates_mod, "InheritanceSize")` (Fig. \@ref(fig:learning-times)B),
but there were some exceptions; some participants took disproportionately
long to recreate the items discovered by their ancestor.

(ref:learning-times) Learning times. **A.** The distribution of learning stage times, with outliers highlighted. **B.** Linear relationship between inheritance sizes and learning times. Outliers excluded from the regression are shown as X's.

```{r learning-times, fig.width=5, fig.height=3, fig.cap="(fig:learning-times)"}
grid.arrange(
  learning_times_plot + ggtitle("A"),
  learning_rates_plot + ggtitle("B"),
  nrow = 1
)
```

Problem solvers who spend more time in the learning stage conversely have less
time in the **playing stage**, where new innovations are tried and discovered.
Fig. \@ref(fig:playing-time-plot) shows the problem solving
rates for diachronic problem solvers based on the length of the playing stage.
The overall rate of problem solving in the playing stage was
`r exp1$minutes_per_new_innovation` minutes per new innovation
(`r exp1$innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based
on the size of the inheritance, as revealed by a model comparison comparing
a model predicting unique innovations from playing time alone to one predicting
unique innovations from both playing time and inheritance size,
`r exp1$playing_time_modcomp`.

(ref:playing-time-plot) Rate of discovery in the playing stage. Playing time
is the amount of time out of a 25 minute session dedicated to discovering new
innovations that were not discovered by an ancestor. The line shows the predictions
of the hierarchical regression model with ±1 standard error. The slope of this
line did not significantly vary based on the size of the inheritance.

```{r playing-time-plot, fig.cap="(ref:playing-time-plot)"}
playing_time_plot
```

## Discussion

In Experiment 1, we found that diachronic problem solvers were able to accumulate
innovations over four generations of problem solving. More than just demonstrating
the capacity to accumulate incremental improvements, we also sought to investigate
the impact of inheritance on future problem solving. To this end, we reported
multiple conflicting accounts. Looking just at the number of innovations discovered,
it appeared that inheritance had a diminishing return on problem solving, such
that later generation participants, by way of inheriting more solutions, had
a harder time exceeding their ancestors than earlier generations. However, if the
difficulty of the innovations is taken into account, a different result is reached,
namely that later generations of diachronic problem solvers were able to solve
more difficult problems than their ancestors. A third conclusion was reached
when looking just at the rate of problem solving in the playing stage, namely
that inheritance had no effect on the rate at which new problems were solved.

We interpret the conflicting nature of these results as evidence that the answer
to the question of whether inheritance impacts problem solving cannot be answered
by investigating diachronic problem solving in isolation. Instead, we use these
results to motivate the approaches in Experiments 2-4 that seek to understand
the impact of diachronic inheritance on problem solving by directly comparing
these results to the results obtained by engaging in a different form of
problem solving. Only by comparing the results of diachronic collaboration
to the result of alternative forms of collaboration can the impact of inheritance
on problem solving truly be discovered.

# Experiment 2

# Experiment 3

# Experiment 4

# Discussion

# References
