---
title: "The impact of inheritance on problem solving ability"
author: "Pierce Edmiston, Max Derex, & Gary Lupyan"
abstract: |
  Human are able to use and improve solutions to problems including tools and
  other technology that were created by other people like no other species of
  animal. What is the impact of inheritance---starting off a problem solving
  task with some of the problems already solved---on future problem solving
  ability? Does inheritance provide a shortcut to individual learning, or
  does it affect the rate or way in which future problems are solved?
  In Experiment 1, we demonstrate that future generations of problem
  solvers are able to cumulatively improve upon the solutions they inherited,
  but were unable to determine whether inheritance had an impact on problem solving
  ability. We resolve this issue in Experiments 2-4 by comparing diachronic
  problem solving to alternative strategies for solving the same set of
  problems but without diachronic inheritance. In Experiment 2 we demonstrate
  that diachronic inheritance results in fewer guesses per solution than
  problem solving without inheritance, but this benefit does not extend to
  solutions that were not inherited. In Experiment 3, we find that inheriting
  from oneself over repeated sessions is more effective than diachronic
  inheritance from someone else. However, we also find that diachronic
  inheritance can remediate the potential costs of fixation on incorrect
  solutions, which for isolated individuals can extend over multiple sessions.
  Experiment 4 includes four-person synchronic teams and finds that the
  inefficiencies of two-person synchronic teams are amplified for four-person
  teams such that diachronic colloration becomes more effective relative to
  synchronic collaboration as labor time increases. Together our results
  demonstrate cumulative gains though diachronic collaboration in a lab
  experiment, and identify some of the ways in which diachronic collaboration
  affords unique kinds of problem solving relative to alternative strategies
  for solving the same set of problems, namely that diachronic problem solving
  can remediate fixation effects and promote a more efficient guessing strategy.
csl: "theme/apa.csl"
bibliography: "references.bib"
output:
  bookdown::html_document2:
    toc: yes
    toc_depth: 2
    toc_float: yes
    css: theme/style.css
    keep_md: yes
    number_sections: no
---
```{r totems, include=FALSE}
library(knitr)
library(crotchet)

opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "hide",
  fig.width = 3,
  fig.height = 3,
  dpi = 96,
  cache = TRUE
)

# Read all knitr chunks in .R files
chunks_dir <- "R"
sapply(list.files(chunks_dir, "*.R", full.names = TRUE), read_chunk)

# Read all the diagrams made with graphviz in the totems package
chunks <- c(
  "team-structures",
  "team-structures-all",
  "team-structures-exp1",
  "team-structures-exp2",
  "team-structures-exp3",
  "team-structures-exp4"
)
sapply(chunks, read_graphviz_chunk, package = "totems")
```

```{r setup, include=FALSE, cache=FALSE}
```

# Introduction

```{r intro, include=FALSE}
```

Humans are effective problem solvers. Part of this effectiveness comes from the
ability to reuse and adapt solutions to problems that were first created by
other people. Inventors rarely create inventions with no precursors, and
commonly use and adapt inventions that were first created by other people
[@Basalla:1989wq; @Valverde:2007fs]. But what do innovators and other problem
solvers gain from inheriting solutions created by others? Do problem solvers
gain a head start on future problem solving, or does inheriting an existing
solution to a problem affect the way in which future problems are solved? We
attempt to answer this question by studying the problem solving ability of
individuals who inherit solutions created by a previous generation of
problem solver at the start of a problem solving task. We refer to this
type of collaboration, organized around inheritance, as **diachronic collaboration**
(Fig. \@ref(fig:team-structures)). We compared the effectiveness of diachronic
collaboration to two alternative strategies for solving the same set of problems in the same
amount of time but without diachronic inheritance: **synchronic collaboration**,
where team members begin with nothing and collaborate in parallel, and
**isolated individuals**, given the most time for individual learning
(Fig. \@ref(fig:team-structures)).

(ref:team-structures) Problem solving strategies. Each node is a single problem
solver. Edge lengths represent labor hours spent solving a set of problems. In
diachronic collaboration, first generation problem solvers work on solutions to
problems before passing them on to be inherited by the next generation. The
impact of inheritance on problem solving is measured as the problem solving
ability of diachronic problem solvers after the first generation. Isolated
individuals do not inherit solutions from other problem solvers, but are given
the most time for individual learning. In synchronic collaboration, multiple
problem solvers work in parallel, sharing their progress, but synchronic teams
do not benefit from diachronic inheritance because all problem solvers
begin at the same starting point.

```{r team-structures, engine="dot", fig.width=1.5, fig.cap="(ref:team-structures)"}
```

This research focuses on diachronic collaboration and the impact of
inheritance on problem solving ability among diachronic collaborators.
Diachronic collaboration exemplifies an important process in human cultural
evolution: the iterative improvement of tools and other solutions to problems
over generations of problem solving generally known as ratcheting
[@Tennie:2009gaa; @Tomasello:1993vh]. Diachronic collaboration is enabled by
social learning mechanisms that allow problem solvers to learn from previous
generations of problem solvers rather than attempting to solve the same problems
through independent and individual learning. Models of the cultural evolutionary
process demonstrate that social learning is adaptive relative to individual
learning when the environment is stable and individual learning is costly
[@Boyd:1988cu]. A population of smart individuals who are unable to
learn from each other are limited to what they can accomplish in a single
lifetime, as their accomplishments cannot be inherited by any future problem
solvers. Thus, social learning is believed to provide a valuable shortcut to
individual learning that causes improvements to accumulate over generations.
However, few attempts have been made to quantify the shortcut gained from
inheritance.

Part of the reason is that models of the cultural evolution process assume that
inheritance is an all-or-none process, meaning that if an innovation is
successful inherited, a shortcut to individual learning is guaranteed. This of
course does not guarantee that all innovations are inherited. If
the likelihood of successfully inheriting an innovation is low, for example,
a complex innovation, such as how to create a fishing net, the population of
effective replicators must be sufficiently large enough or else the loss of
technological complexity may ensue [@Henrich:2004ht]. In addition,
cumulative gains in technological complexity require that
individuals know how to filter the better innovations from maladaptive ones
[@Enquist:2010gw]. In contrast, others have argued that inheritance is
not an all-or-none process. inheritance often involves the transformation
of socially learned information [@Sperber:1996exc], and the successful
transmission of skilled behaviors can take years before becoming efficient
[@Sterelny:2012ea]. As a result, the capacity for inheritance does not
guarantee that diachronic collaboration will result in cumulative improvements.
Our study accounts for this problem by measuring the effect of inheritance
relative to the cost of creating it, and relative to the effectiveness of
altnerative forms of problem solving.

Previous efforts to understand diachronic collaboration in a problem solving
context have focused on our capacity to inherit solutions from others
rather than the effectiveness of diachronic collaboration as a problem solving
strategy. For example, @Tennie:2009gaa argued that humans as compared to nonhuman
primates are uniquely able to inherit solutions to problems that they would be
unable to discover on their own. @Caldwell:2008gf demonstrated the capacity
of diachronic teams to cumulatively improve performance in tasks such as
paper airplane flying and spaghetti tower building, but did not include
any control conditions that could have demonstrated the effects of diachronic
collaboration compared to other strategies, such as an individual given the
same total amount of time for individual learning. Such an experiment
has been performed on pigeons, where diachronic interactions between pairs
of pigeons were found to improve the efficiency of flight paths more than
individual birds and pairs of birds given the same total number of flights.
However, control conditions for transmission chain design experiments are
rarely employed [@Mesoudi:2008kr].

Previous efforts to understanding the impact of inheritance on problem solving
have compared problem solving in groups, where individuals can learn socially
from others, to problem solving as individuals. Although the benefit to
problem solving in groups is known to vary based on the problem being solved,
for problems that require cumulative increases in the complexity of the solutions,
groups of problem solvers are able to discover solutions they
would have been unlikely to discover on their own [@Derex:2015cbb]. However,
this approach of comparing group problem solving to individual problem solving
does not directly test the impact of diachronic inheritance on problem
solving because all problem solvers, even in a group, begin the problem
solving task at the start starting point and with no existing solutions. Here
we extend Derex et. al.'s paradigm in two ways. First, we include a diachronic
condition, where problem solvers start off the problem solving task by inheriting
the solutions discovered by a previous generation, and second, we compare a
diachronic strategy to alternative strategies that do not involve diachronic
inheritance while controlling for labor time, as opposed to calendar time.

The synchronic and isolated strategies serve as control conditions in these
experiments: ways of solving the same set of problems without diachronic
inheritance. Synchronic and isolated problem solvers begin the problem solving
task at the same starting point, i.e. with no existing solutions. Only
diachronic problem solvers after the first generation inherit solutions from
the previous generation to learn from at the start of the experiment.

The diachronic condition will be referred to as the inheritance condition, but
other types of inheritance that do not influence the starting point of a problem
solver beginning a problem solving task are included in the control conditions.
Synchronic problem solvers share solutions with one another throughout the
problem solving task. What distinguishes diachronic and synchronic inheritance
is the time at which the solutions are shared. For diachronic problem solvers,
solutions are inherited once, at the start of the problem solving task. For
synchronic problem solvers, solutions are inherited intermittently, throughout
the task. A comparison of diachronic and synchronic problem solving (Exps. 2,4)
will reveal whether diachronic-generational or synchronic-intermittent
inheritance mechanisms result in more effective problem solving.

A third type of inheritance, individual inheritance, is introduced in Experiment
3. Individual inheritance is when the products of one's own problem solving are
resumed in independent problem solving sessions. Individual inheritance
involves inheriting from oneself, as opposed to inheriting from someone else.
Experiment 3 investigates how diachronic inheritance, the inheritance of partial
solutions at the start of a problem solving task, impacts problem solving
ability differently than an isolated individual, resuming problem solving in
subsequent sessions.

In order to fully understand the impact of inheritance on problem solving
ability, it is necessary to account for both the cost of creating the
inheritance as well as the consequences of inheritance on future problem
solving. In this study the cost and consequence of inheritance is controlled by
allotting a fixed number of labor hours to diachronic teams. The other
strategies investigated in this research are compared given the same number of
labor hours. Labor hours can be contrasted by two other measures of time that
are not controlled in these experiments: calendar hours and individual learning
hours (Fig. \@ref(fig:types-of-time)).

(ref:types-of-time) Types of time. **Left.** A key showing the problem solving
strategies for reference. See Fig. \@ref(fig:team-structures). **Middle.**
Relationship between calendar hours and labor hours. Synchronic collaboration
consumes a fixed amount of labor time more rapidly (per calendar hour) than
either the diachronic or isolated strategies. **Right.** Amount of individual
learning time in each strategy. An isolated strategy maximizes the amount of
time for individual learning relative to both diachronic and synchronic
strategies.

```{r types-of-time, fig.width=8, fig.height=3, fig.cap="(ref:types-of-time)"}
grid.arrange(
  read_graphviz("team-structures", "totems"),
  gg_time,
  gg_person,
  nrow = 1,
  widths = c(0.2, 0.4, 0.4)
)
```

# Methods

```{r methods, include=FALSE}
```

Participants solved as many problems as they could in a fixed amount of time. To
solve each problem, participants had to guess combinations of items that would
result in new items being created. The new items, created from combinations of
previous items, are referred to as innovations, and each innovation is the
solution to a single problem amid the larger set of problems in the problem
solving task. The context of the problem solving task is that participants are
accumulating innovations like axes, stone knives, and paintbrushes in service of
creating "a sacred totem pole to appease the gods." As participants accumulate
more complex innovations, they can then create more intricate and elaborate
totem poles, and receive higher performance scores as a result.

Participants were randomly assigned to either the diachronic condition (the
treatment condition) or one of the control conditions: isolated or synchronic.
Participants in the diachronic condition either participated as a first
generation problem solver, or began the problem solving task with the solutions
discovered by a previous generation. Each participant was assigned to a single
generation in a single team; the innovations of each diachronic problem solver
were only inherited by, at most, one future generation problem solver.

Participants in the synchronic condition collaborated in parallel via separate
computers. Team membership was not revealed to the participants either before or
after the experiment, and no interaction, other than through the computer
interface, was allowed. The only interaction that was allowed was the sharing of
innovations with all members of a synchronic team. Once an innovation was
discovered by a synchronic participant, the recipe for the creation of that
innovation became available for all members of the synchronic team.

Participants in the isolated condition participated in a single 50 minute
session, or alternately returned for four 25 minute sessions, spaced at least
one day apart. When isolated participants returned for later sessions, they
inherited the tools they had discovered in the previous session.

The data generated using these methods are reported in four experiments.
Experiment 1 investigates the result of accumulated inheritance over four
generations of diachronic problem solving. The analyses explore the differences
between first generation problem solvers and problem solvers who benefit from
inheritance, defined as diachronic problem solvers after the first generation.
In Experiment 2, we compare the simplest form of diachronic inheritance, a
two-person diachronic team, to a two-person synchronic strategy, and an isolated
individual given 50 minutes. In Experiment 3, diachronic inheritance from
someone else is compared to inheriting from oneself: isolated participants
returning for multiple sessions. Experiment 4 compares the three main strategies
(diachronic, isolated, and synchronic) given 100 minutes, including a
four-person synchronic strategy.

(ref:team-structures-all) All experimental conditions. The **Diachronic**
condition is the treatment condition: problem solving with inheritance. First
generation diachronic participants began without any previous solutions and
worked on the problem solving task for 25 minutes. Diachronic participants
assigned to generations 2-4 inherited the solutions discovered by the previous
generation at the start of the problem solving task. The two control conditions
were Isolated and Synchronic. **Isolated** participants either completed a
single 50 minute session, or returned for 4 subsequent 25 minute sessions. Upon
returning, these participants "inherited" the solutions they had created in the
previous session. **Synchronic** participants completed the problem solving task
in groups of 2 or 4 individuals, each participating for 25 minutes. Synchronic
participants collaborated in parallel, sharing solutions with all team members
as they were discovered.

```{r team-structures-all, engine="dot", fig.cap="(ref:team-structures-all)", fig.width=1}
```

## Materials and Procedure

**The Totems game**. Participants played a computer game, first reported in
@Derex:2015cbb, requiring players to iteratively solve problems related to
creating new tools out of existing items for the purposes of building "a sacred
totem to appease the gods." To build a totem, participants first needed to
construct an axe out of three independently discovered items: a refined stick
used as a handle, a sharpened rock for the blade, and string wound from bark
fibers for binding (Fig. \@ref(fig:landscape-sample)). In the game, more
advanced tools produce larger and more intricate totems, resulting in higher
performance scores.

(ref:landscape-sample) A sample of the solution landscape. The top row of 6
items are the initial resources available to problem solvers at the start of the
session. Resources must be combined to build the first generation of
innovations, including a refined rock, a club derived from an antler, and a
small branch of a tree. The first 6 generations of possible innovations are
shown. The axe is required to construct the first totem pole.

```{r landscape-sample, fig.cap="(ref:landscape-sample)"}
# draw_graphviz_with_images("landscape-sample", "totems")
draw_image("landscape-sample.gv", "totems")
```

**Guesses.** Participants solved problems by trial-and-error guessing of different
combinations of items they believed would yield new tools. To generate new
guesses, participants dragged items into a workshop panel (Fig.
\@ref(fig:interface)).

Of all the valid combinations of items that could be made, only a small
percentage successfully yielded new items. The 6 initial resources can be
combined with between 1 and 4 other items with replacement for a total of
`r count_unique_guesses(6)` unique guesses, only three of which
yield any new tools: a branch can be broken off a tree, an antler refined into
a club, and two rocks can be combined to yield a shaped stone. As the number
of items available to each problem solver increased, so did the combinatorial
complexity of the problem space. Thus, problems became more difficult
over time.

**Stock.** If a guess yielded an item, it could be stored in a stock panel. A maximum of 12
items could be stored in the stock panel at any given time. Items could be
discarded if space was needed in the stock panel for new items, or if an item
was not needed for any reason.

**Recipes.** Once an item was discovered, the recipe for its creation was recorded in the
history panel. Participants could review their item history and see the recipes
for their previous innovations, even items they had discarded.

In addition to an individual history, the history pane also had other tabs,
depending on which strategy each participant was assigned. Diachronic
participants after the first generation began with a tab that could be selected
to show all item recipes discovered by the previous generation. This was the way
in which solutions to previous problems were shared with future generations.
Isolated participants, if returning for multiple sessions, had a tab with the
solutions they had come up with during their previous session. Synchronic
participants had tabs for other synchronic problem solvers on their team.
The tab displayed the current score of each player, so synchronic were notified
that a discovery had been made by seeing a change in their teammates' score.

(ref:interface) The Totems gameplay interface. Participants generated guesses
by dragging items into the Workshop and selecting the Try button. If the guess
created an item, it could be dragged into the Stock panel, and used again. Once
segments of a totem were made through cutting, carving, and painting, they could be
dragged into the Totem panel for scoring.

```{r interface, fig.cap="(ref:interface)"}
draw_image("gameplay", "totems")
```

**Post experiment survey.** After the experiment was completed, participants
completed a brief post experiment survey assessing their understanding of the
task, their experience with games like the Totems game, and whether they used
any particular strategy for guessing.

## Participants

Participants (N=`n_recruited`) were students at the University of
Wisconsin-Madison who completed the experiment in exchange for course credit.
Students received credit for each completed session based on the duration of the
session, 25 minutes or 50 minutes. Participants in the Isolated, multi-session
experiment who completed all four sessions were awarded additional credit.

Participants were recruited until there were at least 40 teams in each strategy.
Most strategies, other than four-person synchronic teams, ended up receiving
more than 40 teams. The reason for this is that participants were scheduled in
groups of 4 participants at a time, and frequently one or more participants
would fail to show up for their extra credit opportunity. In these cases, the
remaining participants were run as a combination of diachronic and
two-person synchronic participants, resulting in a higher-than-expected
number of participants in those conditions.

Data from individual participants (N=`n_excluded`) were excluded from the final
analyses for the following reasons. Most of these (N=`n_errors`) were excluded
because they experienced errors in the computer game. Others
(N=`n_noncomplaint`) were noncomplaint, giving up on the problem solving task
(N=`n_gave_up`).

```{r counts-table, results="asis"}
kable(ConditionCounts, caption = "Total participant and team recruitment.")
```

# Experiment 1

```{r exp1, include=FALSE}
```

In Experiment 1, we measured problem solving ability over four generations of
diachronic problem solving (Fig. \@ref(fig:team-structures-exp1)). We answered
the following questions about the impact of diachronic inheritance on problem
solving. First, we determined whether and by how much future generations of
problem solvers were able to recreate and exceed the solutions obtained by their
ancestors. Second, we explored the impact of inheritance on future problem
solving by examining whether the size of the inheritance impacted future
problem solving ability.

(ref:team-structures-exp1) Diachronic problem solving examined in
Experiment 1. Participants were assigned to four-person diachronic teams.
Each participant completed the same problem solving task for 25 minutes.
Participants in generations 2-4 began the problem solving task with the
solutions that were discovered by the previous generation.

```{r team-structures-exp1, engine="dot", fig.cap="(ref:team-structures-exp1)"}
```

## Methods

Participants played the Totems game
for a single 25-minute session. Each participant was assigned to a single
generation within a single four-person team (($N_{participants}$=`r exp1$n_participants`) , $N_{teams}$=`r exp1$n_teams`).

Participants assigned to the first generation of a new team played to
Totems game for 25-minutes. Their solutions were inherited by one
(and only one) generation 2 player, and so on through 4 generations
of diachronic collaboration.

(ref:exp1-participants) Participants in Experiment 1.

```{r exp1-participants, results="asis"}
kable(Exp1Participants, caption = "(ref:exp1-participants)")
```

## Results

Diachronic problem solvers were consistently able to exceed the number of
innovations discovered by their predecessors (Fig. \@ref(fig:innovations-by-generation)).
This result was obtained using Page's trend test, which is a
repeated measure test for monotonicity. We used this test to measure whether
the number of innovations increased within each team and over generations
against the null hypothesis that the number of innovations did not change
from generation to generation. We found evidence for cumulative
increases in problem solving ability such that later generations of problem
solvers were able to discover more innovations in a single 25 minute session
than their predecessors, `r exp1$page_test`.

We next measured the linear increase in innovations over generations
within each team by fitting a hierarchical regression model to the total
innovations achieved in each generation with nested effects (intercepts
and slopes) for teams. On average, second generation participants were
able to discover `r exp1$gen0_slope` more innovations than first generation
participants, `r exp1$gen0_slope_stats`. This linear effect decreased by
`r exp1$gen0sqr_slope` each generation for third and fourth generation
participants, `r exp1$gen0sqr_slope_stats`. A model comparison of hierarchical
regression models fit with linear and quadratic components revealed
that a quadratic fit was significantly better than a linear fit alone,
`r exp1$quad_modcomp`. This result suggests that accumulated inheritance had a
diminishing return on future problem solving.

(ref:innovations-by-generation) Number of innovations discovered by each
generation. Each of the thin green lines is a team. Means in each generation are
shown as green X's. The thick blue line is the model predictions with ±1
standard error.

```{r innovations-by-generation, fig.cap="(ref:innovations-by-generation)"}
innovations_by_generation_plot
```

We next consider different explanations for the diminishing return of inheritance on
future problem solving. First, we tested whether a larger inheritance had a detrimental
effect on future problem solving. We found that an increase in the number of
inherited innovations is negatively associated with a decrease in the number of
new innovations discovered by future generations of problem solvers,
`r exp1$inheritance_size_slope_stats` (Fig. \@ref(fig:size-of-inheritance)).
This finding is in line with the current proposal that inheritance has an
influence on future problem solving beyond simply providing a shortcut
to individual learning. In this case, the impact is negative, such
that inheriting more innovations appears to _decrease_ problem solving ability.

(ref:size-of-inheritance) Number of innovations created relative to those
inherited. **A.** The relationship between created and inherited innovations.
The dotted line is a reference with slope=1 such that points above the line
indicate future generations exceeding their ancestors. **B.** The relationship
between created innovations controlling for those inherited. The same reference
line as in A is now horizontal. The line shows the predictions of the
hierarchical regression model with ±1 standard error.

```{r size-of-inheritance, fig.width=6, fig.height=4, fig.cap="(ref:size-of-inheritance)"}
grid.arrange(
  innovations_created_and_inherited_plot + ggtitle("A"),
  new_innovations_plot + ggtitle("B"),
  nrow = 1
)
```

However, an alternative explanation, not having to do with inheritance affecting
problem solving ability, is that as more innovations are accumulated, the task
gets more difficult, such that problem solving might slow down irregardless of
what was inherited. In the Totems game, as innovations are accumulated, the total
number of possible guesses than can be made from the accumulated items increases
exponentially, so that later generations of problems are less likely to stumble
upon a correct solution by chance. To control for combinatorial complexity, we
created an alternative outcome measure that, rather than counting all
innovations equally, instead weighted each innovation relative to the size of
the inventory at the time at which the innovation was discovered. By this
measure, innovations from later generations are expected to be more difficult
because they require combinatorially more guesses in order to find by chance
alone.

Fig. \@ref(fig:delta-difficulty-fig) shows the
change in accumulated difficulty score based on the number of innovations
that were inherited. In contrast to the analysis that treated all
innovations equally, when innovations are weighted by combinatorial complexity,
the effects of inheritance and diachronic collaboration appear to be beneficial.
Diachronic problem solvers who inherit more innovations are better
able to exceed the accumulated difficulty score of their ancestors than
first generation diachronic problem solvers, `r exp1$delta_diff_slope_stats`.

(ref:delta-difficulty-fig) Change in accumulated difficulty score by
inheritance. Accumulated difficulty scores are sums of discovered innovations
weighted by the combinatorial complexity of the possible choices available to
the participant when the innovation was discovered. The outcome measure is the
difference between the accumulated difficulty scores of subsequent generations.
The line shows the predictions of a hierarchical regression model with ±1
standard error.

```{r delta-difficulty-fig, fig.cap="(ref:delta-difficulty-fig)"}
delta_difficulty_plot
```

The final explanation we consider regarding whether inheritance affected problem
solving ability is that larger inheritances take more time to recreate, thus
leaving less time for future problem solving. Even if inheritance did not affect
problem solving ability, participants who inherit more innovations may not
discover as many future innovations simply because they used up most of their
time recreating the items that were inherted. To investigate, we looked at
problem solving rates controlling for the amount of time spent recreating the
inherited items.

Diachronic problem solvers after the first generation inherit the recipes
for creating innovations in the Totems game, not the innovations themselves.
Thus, diachronic problem solvers who inherit solutions must start the
experiment by recreating the innovations already discovered by their
ancestors. On average, diachronic problem solvers after the first generation
spent `r exp1$mean_learning_time_min` minutes of a 25 minute session
(`r exp1$proportion_learning_time`%) recreating the innovations that were
inherted (Fig. \@ref(fig:learning-times)A). We refer to to this stage
as the **learning stage**.

The length of the learning stage depended on the number of innovations that were
inherited. On average diachronic problem solvers inherited
`r exp1$mean_inheritance_size` innovations (SD=`r exp1$sd_inheritance_size`). The
length of the learning stage scaled linearly with the size of the inheritance,
`r report_lm_mod(learning_rates_mod, "InheritanceSize")` (Fig. \@ref(fig:learning-times)B),
but there were some exceptions; some participants took disproportionately
long to recreate the items discovered by their ancestors. Participants
who spent over 80% of the session in the learning stage were considered outliers,
and excluded from analysis.

(ref:learning-times) Learning times. **A.** The distribution of learning times,
with outliers highlighted. **B.** Linear relationship between inheritance sizes
and learning times. Outliers excluded from the regression are shown as X's.

```{r learning-times, fig.width=6, fig.height=4, fig.cap="(ref:learning-times)"}
grid.arrange(
  learning_times_plot + ggtitle("A"),
  learning_rates_plot + ggtitle("B"),
  nrow = 1
)
```

Problem solvers who spend more time in the learning stage conversely have less
time in the **playing stage**, where new innovations are tried and discovered.
Fig. \@ref(fig:playing-time-plot) shows the problem solving
rates for diachronic problem solvers based on the length of the playing stage.
The overall rate of problem solving in the playing stage was
`r exp1$minutes_per_new_innovation` minutes per new innovation
(`r exp1$new_innovations_per_minute` innovations per minute),
`r exp1$playing_time_slope_stats`. This rate was not found to vary based
on the size of the inheritance, as revealed by a model comparison comparing
a model predicting unique innovations from playing time alone to one predicting
unique innovations from the interaction between playing time and inheritance size,
`r exp1$playing_time_modcomp`. This result suggests that inheriting more
items did not have an effect on the rate at which new problems are solved
when playing time is controlled.

(ref:playing-time-plot) Problem solving rate in the playing stage. Playing time
is the amount of time out of a 25 minute session dedicated to discovering new
innovations that were not discovered by an ancestor. The line shows the predictions
of the hierarchical regression model with ±1 standard error. The slope of this
line did not significantly vary based on the size of the inheritance.

```{r playing-time-plot, fig.cap="(ref:playing-time-plot)"}
playing_time_plot
```

## Discussion

In Experiment 1, we measured the problem solving ability of four-person diachronic teams, and found that diachronic problem solvers were consistently able to cumulatively improve upon the solutions they inherited. Although this finding is not surprising given that participants started a problem solving task with some of the problems already solved, it is important to highlight that solving new problems first required recreating all of the inherited solutions, which took time and effort. Once recreated, some participants were unable to solve any new problems. Others were unable to recreate all of their inherited solutions in the 25-minute session---although this behavior is likely attributable to non-compliant behavior or a misunderstanding of the task. Nonetheless, the presence of these outliers simply highlights that diachronic problem solving, as measured in this paradigm, is not guaranteed to improve cumulatively, and explains why it is important to first establish that
performance does indeed accumulate in diachronic teams.

We next asked whether the size of the inheritance had an impact on the likelihood of being improved, and found multiple, conflicting results. When all innovations are weighted equally, the effect of inheritance decreased as inheritance size increased, which could mean that that larger inheritances had a negative effect on problem solving. However, when the innovations are weighted by the difficulty of the solution in terms of the combinatorial complexity of the solution, the benefit of inheritance did not decrease, as problem solvers who inherited more solutions were able to solve more combinatorially difficult problems. Finally, when controlling for the amount of time spent solving new problems as opposed to recreating inherited ones, the rate of problem solving (solutions per minute) also did not decrease with larger inheritances. However, this conclusion is limited by the collinear relationship between inheritance size and time solving unique problems.

We believe the source of these conflicting interpretations is an underlying inability to understand the impact of inheritance on problem solving without comparison to a control condition. In Experiment 2, we attempt to understand the impact of inheritance on problem solving relative to the benefits that may accrue from individual learning over four separate 25-minute sessions.

# Experiment 2

```{r exp2, include=FALSE}
```

In Experiment 2, we compared the simplest form of diachronic inheritance---a
two-person diachronic team---to two alternative strategies for allocating the
same 50 minutes of labor: isolated individuals working alone and two-person
sychronic teams (Fig. \@ref(fig:team-structures-exp2)). In our analysis we focus
primarily on the comparison of the second generation diachronic problem solvers,
who began the problem solving task with the solutions inherited from a first
generation problem solver, to problem solvers utilizing the alternative
strategies to solve the same set of problems. In addition to measuring the
impact of inheritance on overall performance, we also investigate whether
diachronic collaboration causes problem solvers to approach the task differently
than the alternative strategies. To answer this question, we investigate the
average number of guesses expended to discover each new item as well as the
distribution of guess types made by problem solvers in each strategy.

(ref:team-structures-exp2) Problem solving strategies compared in Experiment 2.

```{r team-structures-exp2, engine="dot", fig.cap="(ref:team-structures-exp2)"}
```

## Methods

Participants assigned to the diachronic condition completed the Experiment as
reported above. Participants in the synchronic condition were randomly assigned
to two-person teams. Synchronic problem solvers each attempted the Totems game
in a single 25-minute session at separate computers. Synchronic team membership
was anonymized such that participants were not explicitly made aware of who
their teammate was, and were not allowed to interact or communicate with their
teammate other than through the computer interface. The only interaction allowed
through the computer interface was the sharing of unique innovations. Once
a synchronic problem solver discovered a new item, the recipe for its creation
could be viewed and copied by the other problem solver in the team. Isolated
individuals were recruited for a single 50-minute session.

(ref:exp2-participants) Participants in Experiment 2. Data for diachronic
participants was taken from the first and second generations of Experiment 1.

```{r exp2-participants, results="asis"}
kable(Exp2N, caption = "(ref:exp2-participants)")
```

## Results

As reported in Experiment 1, diachronic problem solvers with inheritance
significantly exceeded the number of innovations discovered in a single
25-minute session, `r exp2$DG2_v_DG1`. In Experiment 2, we found that diachronic
problem solvers with inheritance outperformed isolated individuals given 50
minutes, however this difference was not reliable, `r exp2$DG2_v_I50`.
Conversely, two-person synchronic teams outperformed second generation diachronic
teams, but again, this different was not reliable, `r exp2$S2_v_DG2`. These
results, although they do not demonstrate a clear benefit to inheritance in terms of
overall performance, are in line with our conclusion that inheritance does not
negatively impact problem solving relative to alternative strategies (Fig.
\@ref(fig:innovations-50min).

Comparing two-person teams (both diachronic and synchronic) to isolated
individuals revealed a benefit to teamwork in this task,
`r exp2$teamwork_stats`, with no residual difference between diachronic and
synchronic strategies, `r exp2$teamwork_residual`. This result suggests that
maximizing individual learning time is not the most productive way to improve
problem solving efficiency, instead demonstrating that teamwork is beneficial
even when controlling for labor hours, instead of calendar hours.

(ref:innovations-50min) Innovations discovered in 50 labor minutes by strategy. **A.**
Rate of innovation by labor hours. **B.** Final innovations by strategy. Bars with
errors show model predictions with ±1 standard error.

```{r innovations-50min, fig.width=6, fig.height=4, fig.cap="(ref:innovations-50min"}
grid.arrange(
  innovation_rate_50min_plot + ggtitle("A"),
  num_innovations_50min_plot + ggtitle("B"),
  nrow = 1,
  widths = c(0.6, 0.4)
)
```

Next we investigated the impact of inheritance on a per-innovation basis.
We hypothesized that inheritance would reduce the overall number of guesses
expended in discovering each item over the alternative strategies.
To count the number of guesses for each innovation, we tallied all guesses
made from the moment in which a new item was adjacent to the current inventory,
until that item was discovered. This way we only included guesses that could
have created the item.

We found that diachronic inheritance did reduce the total number of guesses
made for each new item relative to the same items being guessed with an
isolated and synchronic strategy (Fig. \@ref(fig:guesses-per-item)A). However,
we found that this overall benefit to inheritance in terms of fewer guesses
per item did not extend to items that had not been discovered by the previous
generation. Comparing the number of guesses per item only for items that
were newly discovered by the second generation diachronic problem solvers
did not reveal a benefit to inheritance (Fig. \@ref(fig:guesses-per-item)B).

```{r guesses-per-item, fig.width=6, fig.height=4}
grid.arrange(
  guesses_per_item_by_inheritance_plot + ggtitle("A"),
  guesses_per_new_item_by_inheritance_plot + ggtitle("B"),
  nrow = 1
)
```

Finally we investigated the guessing strategy employed in each strategy.
We were interested in whether inheritance caused second generation
diachronic problem solvers to guess differently than players in
the other strategies. Our results are shown in Fig. \@ref(fig:guess-types).

```{r guess-types, fig.width=6}
prop_guess_types_50min_plot
```

## Discussion

# Experiment 3

(ref:team-structures-exp3) Problem solving strategies compared in Experiment 3.

```{r team-structures-exp3, engine="dot", fig.cap="(ref:team-structures-exp3)"}
```

## Methods

## Results

## Discussion

# Experiment 4

(ref:team-structures-exp4) Problem solving strategies compared in Experiment 4.

```{r team-structures-exp4, engine="dot", fig.cap="(ref:team-structures-exp4)"}
```

## Methods

## Results

## Discussion

# Discussion

# References
